### 2020-07-11 Tex/MSc Data Science Project Proposal.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mdframed}
\usepackage[margin=3cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{textcomp}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage[colorlinks=true]{hyperref}
\usepackage{cleveref}
\crefdefaultlabelformat{[#2#1#3]}
\begin{document}
\renewcommand\citeform[1]{[#1]}
%
\title{MSc Data Science Proposal: Can a Convolutional Neural Network judge a book by its cover?}
\author{\\Ryan Hill MSci\\\\
Supervisor: Dr Hubie Chen\\\\
Birkbeck, University Of London\\\\
\texttt{rhill06@mail.bbk.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
\graphicspath{{images/}}
\begin{abstract}
	In this work we make a proposal for training convolutional neural networks to predict the genre of a book, based on its cover. We evaluate the prior work done in this area and the proposed improvements of the existing dataset known as \emph{BookCover30}. We break the work into 3 stages and provide detail on the proposed hardware and software usage within the project.
\end{abstract}
\begin{figure}[!b]
	\centering
	\includegraphics[scale=0.4]{bbk_logo.jpg}
\end{figure}

\clearpage
%
{\hypersetup{linkcolor=black}
\tableofcontents}
\thispagestyle{empty}
\clearpage
%
\setcounter{page}{1}
\section{Introduction} % (fold)
\label{sec:intro}
\subsection{Background} % (fold)
\label{sub:background}
\emph{Don't judge a book by its cover} is a common English idiom, with its origin somewhere around the mid 19\textsuperscript{th} century, often cited as being from the newspaper \emph{Piqua Democrat} as follows:
\begin{quote}
	``Don’t judge a book by its cover, see a man by his cloth, as there is often a good deal of solid worth and superior skill underneath a jacket and yaller pants.''
\end{quote}
The phrase itself refers to the idea that one should not judge an item, or person, by how it first appears, but instead you should take the time to understand it better. That being said, a huge amount of cost and effort can go into the design and choice of a cover for a book; costs for cover design alone can range from a few dollars to a few thousand \cite{AdriasG}, and revenue from publishing in 2018 was around \$122bn \cite{Watson}. This value, combined with the fact that humans have been shown to make judgments within 100ms of exposure to another person \cite{Willis2006}, means it is reasonable to assume that this advice is not often taken literally and that as humans we do, in fact, judge a book by its cover.

Beyond the focus of the judgement placed on book covers, they are designed to help the potential reader understand the type of story, or contents, a book may contain. As an example, consider the covers shown in \cref{fig:2bookcovers} that should, even if they had no words on them, help the reader identify immediately the type of content contained within. The type of this content is what we refer to as the \emph{genre} of the book. Even within this example it is clear that there could be some subjectivity, as the reader you may see a dragon and assume it is a fantasy novel, but the Teen \& Young Adult genre often encompasses many genres just due to the age of the intended audience. 
\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=.8\linewidth]{200px-Eragon_book_cover.png}  
	  \caption{Cover of Eragon, Teen \& Young Adult Genre}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=.8\linewidth]{cookbook_cover.jpg} 
	  \caption{Cover of Good Food Cookbook, Cookbooks, Food \& Wine Genre}
    \end{subfigure}
	\caption{Two book covers with distinctly different designs and genres}
	\label{fig:2bookcovers}
\end{figure*}

% subsection background (end)
\subsection{Problem Statement} % (fold)
\label{sub:problem_statement}
In this work we will try to answer one seemingly simple question; \emph{can a Convolutional Neural Network predict the genre of a book based solely upon an image of the front of the book?} Or put another way, \emph{can a Convolutional Neural Network judge a book by its cover?} This may seem like a fairly simple question, but in many cases it might already be quite difficult for a human to distinguish which genre a given book belongs to as just discussed. Beyond answering this question we will also attempt to take the easiest and hardest to identify genres (as determined by the accuracy of the classifier) and use feature visualisation to generate ideal covers for these genres, as the Convolutional Neural Network \emph{sees} them.

The rest of this work is structed as follows; first we review the existing literature and history of computer vision, predicting genre from book covers, and feature visualisation in \cref{sec:lit_rev}. We then discuss in more detail the plans for how to tackle this problem statement, including the hardware and software we plan to use, in \cref{sec:proposal}. Finally, we summarise this work in \cref{sec:conclusion}.
% subsection problem_statement (end)
% section intro (end)
\section{Literature Review} % (fold)
\label{sec:lit_rev}

\subsection{History of Computer Vision}
\label{sec:computer_vision}
The work in this section is mostly informed by the comprehensive article by R Demush \cite{Demush}. Computer vision was originally, like most theoretical work in artificial intelligence (AI) and machine learning, conceived in the 1960s. A 1966 summer project by MIT was intending to solve the computer vision problem within a few months, simply put by attaching a camera to a computer it would then `describe what it saw' \cite{Papert1966}. Clearly, that was a little over-ambitious.

One of the most influential papers in computer vision actually came in 1959 when two neurophysiologists, Hubel and Wiesel, published their work studying the response of visual neurons of a cat's brain, and how these experiences shape the neural architecture \cite{Hubel1959}. Paraphrasing some of their results, they discovered that there are both simple and complex neurons, with visual processing beginning with simple structures such as edges and then going to more complex structures. This is essentially the basis behind deep learning today, especially in computer vision. However, even with this knowledge, it wasn't until 1982 when neuroscientist David Marr published the next piece of the puzzle in the form of a book entitled \emph{Vision: A Computational Investigation into the Human Representation and Processing of Visual Information}\cite{Marr:1982:VCI:1095712}. While his representational framework was abstract and high-level, it was still ground-breaking work and helped us understand how we create more complex images from the simple building blocks, and gave us a path to understanding how we could build computer models to do this.

Around the same time \emph{Neocognitron}, a multilayer artificial neural net, was proposed by Kunihiko Fukushima which contained several convolution layers, the types of layers still used today. In 1989 Yann LeCun took this network and developed LeNet-5 from it, a Convolutional Neural Network (CNN) that contains many elements used by modern CNNs. This model had the goal of classifying images, as has become the main use of CNNs in the modern era. This work \cite{Cun1989} also included the production of the now famous MNIST dataset of handwritten digits.

For a few years, work continued to try and construct 3D models from 2D images, before instead moving towards object recognition around the turn of the new millennia. Development continued over the coming years, including a new focus on having a set of communal benchmark datasets to enable easy comparison. These included the original MNIST, before the development of Pascal VOC \cite{Everingham10}, and finally ImageNet \cite{imagenet_cvpr09}in 2010. ImageNet has now become the de-facto benchmark dataset for image classification with over a million manually cleaned images covering 1000 object classes. At this point, development started coming specifically within the architecture of CNNs, which is discussed briefly in \cref{sub:cnn_architecture}.

\subsection{Prior Work} % (fold)
\label{sub:prior_work}
Design has been explored in many areas before with regards to its history and the various styles, however it is a relatively new area for machine learning in comparison to other fields. CNNs have been used to identify the artists of works \cite{Karayev2013}, they have been suggested for use in forgery detection \cite{Gideon2018}, and with the advances in Generative Adversarial Networks (GANs) we can also produce realistic examples or enhance images of a given classification \cite{Ledig2016}.

CNNs have also been used to help classify artwork \cite{Zujovic2009}, clothing types\cite{Xiao2017}, and music \cite{Tzanetakis2002}; but only Iwana et al. \cite{KenjiIwana} have applied this work to book covers to attempt to classify the genre so far as we can tell. It is their work that we propose to build on. They produced a cleaned dataset, discussed further in \cref{sub:existing_datasets}, of 57,000 images spanning 30 equally represented genres and presented the classification results of an unseen test set of 10\% of the data after training AlexNet, a specific CNN architecture. They report a Top 1 accuracy of 24.7\% and a Top 3 accuracy of 40.3\%  across the 30 classes (i.e. a random chance of 3.3\% and 9.9\% respectively). Class specific Top 1 accuracies reach up to 68.9\% for the Test Preparation genre, but are as low as 6.8\% for the Politics \& Social Sciences genre. We will explore the results of their work further, and discuss the potential reasoning for this range of performance by class if we see similar results, in our main work.
% subsection prior_work (end)
\subsection{CNN Architectures and Transfer Learning}
\label{sub:cnn_architecture}
For all of the development of CNNs discussed in this section, they continue to maintain the same overall structure they started with, that of having a clear delineation on the network where they go from feature extraction to classification as shown in \cref{fig:cnn_struct}. The first part of the network consists of various convolution/filter and pooling layers, used to extract image features (e.g. edges) and to reduce the image dimension respectively. The second part of the network is a traditional densely connected artificial neural network used for the actual classification; a few fully connected layers followed by the output/target layer with as many neurons are there classes in the problem. Much of the development has gone into the feature extraction part of the network over the last 8 years or so; starting in 2012 with AlexNet \cite{Krizhevsky2017}, with the paper hailed as one of the most influential papers in applied computer vision. A few years later these were joined by what have been come to be known as the VGG CNNs (named after the Visual Geometry Group at Oxford)\cite{Simonyan2015}. The following year the family of Inception models (also known as GoogLeNet) start to appear \cite{Szegedy}; around the exact same time ResNets were proposed by He et al. at Microsoft \cite{He2015} which later led to the combination of these in 2017. We do not detail here the specifics of CNN layers, including the types of convolution or pooling, nor the differences in all these architectures as we will detail these in the work itself, but those chosen to be used within this work are mentioned in \cref{sub:part_1_genre_predictor}.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.3]{cnn_struct.jpeg}
	\caption{An simplified example of a CNN architecture}
	\label{fig:cnn_struct}
\end{figure}

We will be using transfer learning to reduce the effort involved in training our network. Transfer learning was first proposed in 1993 by Lorien Pratt \cite{Pratt1993} and has been said by Andrew Ng to be the next driver for commercial machine learning success. A comprehensive survey of transfer learning has been completed by Zhuang et al. \cite{Zhuang2019} so we do not detail here all of the developments and differences in various transfer learning algorithms, especially as we plan to take the most basic approach. Whilst many have worked to develop detailed algorithms that take into account the overlap of the source data domain and the target data domain and adjust a network accordingly, we are taking a more simple approach that is discussed in \cref{sub:part_1_genre_predictor} and reusing all but the (dense) classification layers of a network pretrained on ImageNet as the starting point for our training.
% subsection cnn_architecture (end)
\subsection{Existing Datasets} % (fold)
\label{sub:existing_datasets}
The dataset collected by Iwana et al.\cite{KenjiIwana} is the largest, and as far as we can tell only, clean dataset to collate together the book covers of over tens of thousands of books. The dataset exists in two distinct forms; the \emph{BookCover30} dataset and the \emph{Book32} dataset that contain 57,000 and 207,572 records respectively, split across 30 and 32 classes again respectively. Both datasets have the same columns; \emph{Amazon Index, Filename, Image URL, Title, Author, Category Id, Category}. For \emph{BookCover30} there is also a reduced dataset containing just the filename and category ID that can be used for training and testing after image download. A comparison of the volumes, as well as the split between train and test sets, and a key between the category ID and category label can be seen in \cref{tab:book_data_comparison}. In the case that a book belonged to multiple genres, the reported genre was chosen at random.

\begin{table}[]
	\centering
	\begin{tabular}{|l|l|l|l|l|}
	\hline
	\textbf{Label} & \textbf{Category Name}        & \textbf{\begin{tabular}[c]{@{}l@{}}Size\\ Book32\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Size \\ BookCover30\\ Training\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Size \\ BookCover30\\ Test\end{tabular}} \\ \hline
	0              & Arts \& Photography           & 6,460                                                          & 1710                                                                            & 190                                                                         \\ \hline
	1              & Biographies \& Memoirs        & 4,261                                                          & 1710                                                                            & 190                                                                         \\ \hline
	2              & Business \& Money             & 9,965                                                          & 1710                                                                            & 190                                                                         \\ \hline
	3              & Calendars                     & 2,636                                                          & 1710                                                                            & 190                                                                         \\ \hline
	4              & Children's Books              & 13,605                                                         & 1710                                                                            & 190                                                                         \\ \hline
	5              & Comics \& Graphic Novels      & 3,026                                                          & 1710                                                                            & 190                                                                         \\ \hline
	6              & Computers \& Technology       & 7,979                                                          & 1710                                                                            & 190                                                                         \\ \hline
	7              & Cookbooks, Food \& Wine       & 8,802                                                          & 1710                                                                            & 190                                                                         \\ \hline
	8              & Crafts, Hobbies \& Home       & 9,934                                                          & 1710                                                                            & 190                                                                         \\ \hline
	9              & Christian Books \& Bibles     & 9,139                                                          & 1710                                                                            & 190                                                                         \\ \hline
	10             & Engineering \& Transportation & 2,672                                                          & 1710                                                                            & 190                                                                         \\ \hline
	11             & Health, Fitness \& Dieting    & 11,886                                                         & 1710                                                                            & 190                                                                         \\ \hline
	12             & History                       & 6,807                                                          & 1710                                                                            & 190                                                                         \\ \hline
	13             & Humor \& Entertainment        & 6,896                                                          & 1710                                                                            & 190                                                                         \\ \hline
	14             & Law                           & 7,314                                                          & 1710                                                                            & 190                                                                         \\ \hline
	15             & Literature \& Fiction         & 7,580                                                          & 1710                                                                            & 190                                                                         \\ \hline
	16             & Medical Books                 & 12,089                                                         & 1710                                                                            & 190                                                                         \\ \hline
	17             & Mystery, Thriller \& Suspense & 1,998                                                          & 1710                                                                            & 190                                                                         \\ \hline
	18             & Parenting \& Relationships    & 2,523                                                          & 1710                                                                            & 190                                                                         \\ \hline
	19             & Politics \& Social Sciences   & 3,402                                                          & 1710                                                                            & 190                                                                         \\ \hline
	20             & Reference                     & 3,268                                                          & 1710                                                                            & 190                                                                         \\ \hline
	21             & Religion \& Spirituality      & 7,559                                                          & 1710                                                                            & 190                                                                         \\ \hline
	22             & Romance                       & 4,291                                                          & 1710                                                                            & 190                                                                         \\ \hline
	23             & Science \& Math               & 9,276                                                          & 1710                                                                            & 190                                                                         \\ \hline
	24             & Science Fiction \& Fantasy    & 3,800                                                          & 1710                                                                            & 190                                                                         \\ \hline
	25             & Self-Help                     & 2,703                                                          & 1710                                                                            & 190                                                                         \\ \hline
	26             & Sports \& Outdoors            & 5,968                                                          & 1710                                                                            & 190                                                                         \\ \hline
	27             & Teen \& Young Adult           & 7,489                                                          & 1710                                                                            & 190                                                                         \\ \hline
	28             & Test Preparation              & 2,906                                                          & 1710                                                                            & 190                                                                         \\ \hline
	29             & Travel                        & 18,338                                                         & 1710                                                                            & 190                                                                         \\ \hline
	30             & Gay \& Lesbian                & 1,339                                                          & 0                                                                               & 0                                                                           \\ \hline
	31             & Education \& Teaching         & 1,664                                                          & 0                                                                               & 0                                                                           \\ \hline
	\end{tabular}
	\caption{Comparison of the number of records per category between the Book32 and Book30 training and test sets}
	\label{tab:book_data_comparison}
	\end{table}

The reduced \emph{BookCover30} dataset has already been pre-processed to remove 2 underrepresented classes, and to scale and shape the images; this will likely have to be re-sourced before the processing has taken place as the various CNNs we are going to use have different input shapes. There is also some other issues with the dataset; in particular that no manual filtering was done to ensure that the images are representative of the actual book cover, an example where this is not the case can be seen in \cref{fig:bad_cover}. The data also looks to cover multiple languages so it is possible that there will be duplication of items (although often foreign editions of books have different covers) and this could also lead to additional difficulty if there is a difference in design choices between different countries. Furthermore, this dataset and paper were first published in 2016; while using the same dataset would be useful for comparison it is possible that the same script they used to generate their data could be repurposed to allow for a more up-to-date dataset to be produced, with higher resolution images and a potentially more representative sample of modern books. More discussion on this is repeated in \cref{sub:part_0_data_collection_and_preparation}.

We mention here for completeness that during our research we did come across a potential alternative source of data, that of the Open Library \cite{openAbout} who as part of the Internet Archive project have the goal to have\emph{"One web page for every book ever published"}. The regular data dumps \cite{openDataDumps} are a possible alternative source to using the Iwana dataset, however this dataset would come with some uncertainty attached. Whilst the volume of records is undoubtedly larger in size, it is not clear without processing all of the records how many of these actually have a genre classification at all, and given that their book covers data dump has not been updated since 2016 it does not provide any more recent data than Iwana's does. We would also be remiss to not mention that as this is an open-source project, the quality of the data itself would be more in question than with the \emph{Book32} records. Finally, it would remove our ability to use Iwana et al's paper as a reference in terms of performance of the networks as not only would the training data itself be vastly different, but the genres are also different in this source.

Overall the \emph{Book32} dataset is the best quality and most label-complete dataset available that we could find, and to produce a new one from scratch even using Open Library as a foundation would take a large amount of work for likely very little gain, so it is this dataset we will use in our work.
\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.4]{0345535529.jpg}
	\caption{An image in the \emph{BookCover30} dataset showing an anthology box collection rather than the cover}
	\label{fig:bad_cover}
\end{figure}
% subsection existing_datasets (end)
\subsection{Debugging CNNs}
\label{sub:debug_cnn}
CNNs, along with most other deep-learning methods, have been seen for a long time as \emph{black box} methods i.e. it is not possible to understand what a specific neuron or layer within the network is doing. The attitude to black box methods has soured over the years for 2 major reasons. The first being that black box methods are illegal in some fields such as underwriting; a lender must be able to explain to an applicant why they were rejected for a loan, and black box methods do not allow for this. The second is that they are almost impossible to debug and understand where the network may be making an error. Many examples of this were first shown by Nguyen et al \cite{Nguyen} in 2015, where images could be evolved to trick CNNs that return high precision results for obviously garbage images. More recently Su et al. \cite{Su} managed to show that perturbation of even a single pixel in an image can generate a change in the predicted class from otherwise high performing classifiers. These examples, combined with a drive for more explainable AI, has led to a few key developments in the debugging of CNNs via representative visualisation.

In 2018 Qin et al. \cite{Qin2018a} produced a survey of the major visualisation methods and whilst there has been some developments, mostly in the implementation of these methods in common software packages, their work remains a good summary of the state of the work. They call out four key visualisation methods:
\begin{enumerate}
	\item Activation maximisation - A visualised image is generated such that it maximises a specific neuron's activation.
	\item DeconNet - Uses an inverted CNN structure to identify the pattern in the input image that triggers specific neuron activation.
	\item Network Inversion - Recreates an image based on the input image from a specific layer, helping identify what information is preserved by this layer.
	\item Network dissection  - Applies semantic concepts to each neuron such as scene, material, object etc.
\end{enumerate}
For the purpose of this work we focus only on activation maximisation as this is what we plan to use to generate representative images of some of the genres.

Activation maximisation was proposed by Erhan et al. \cite{erhan2009} in 2009 and the idea itself is quite simple; iteratively perturbate an image to maximise the activation of a given neuron. The resultant image then represents what that neuron has learned. The actual algorithm uses backpropagation instead of perturbation to improve the synthesised image over many iterations. Compared to normal backpropagation where we calculate the change in the network error with respect to each weight and bias and use that to update the values, here we calculate the change in activation with respect to the input and use this as the base for our update rule:
\begin{equation}
	input_{t} = input_{t-1} - \eta\times\frac{\partial ActivationMaximisationLoss}{\partial input_{t-1}}
\end{equation} 
where $ActivationMaximisationLoss$ is some loss measure that decreases as the activation increases, and $\eta$ is a learning rate. We stop when the activation of the neuron cannot be increased further. This can be applied to almost any neuron in the network with no alteration to the structure except for those in the final layer. The final layer activation function, the transformation applied to the inputs of the neuron, must be changed from the pre-existing SoftMax ($\sigma$) as defined in \cref{eq:softmax}, which takes a vector of activations for the $K$ output classes and normalises them to be between 0 and 1 with a sum of 1, and be replaced with a linear function instead (often just the linear sum of inputs is used). The reason for this is that to maximise the output of a SoftMax function for a single neuron, it is possible to minimise the values of the other neurons. This leads to a case where the generated image would not be representative of our target class, but just be more representative than it is of the other classes. By using the unbounded linear function we can continue to apply updates to the input until we see little to no increase in the activation for our target neuron, irrespective of the other class activations.

\begin{equation}
	\sigma(\vec{x})_i = \frac{e^{x_i}}{\sum_{j=1}^{K}e^{x_j}}
	\label{eq:softmax}
\end{equation}

Often, this approach is taken together with some form of normalisation to improve the interpretability of these images. As mentioned, pictures that are composed entirely of noise can trigger activation in the output neurons in a network with high certainty, so these normalisation methods are applied to reduce the production of these types of images and return more visually understandable results. Some of these normalisations, such as $l_2$ decay (which usually uses the sum of squared weight, but pixel values in this case) are used to prevent a few pixels dominating the image, or applying a Gaussian Blur can be done to ensure that no area of the image has too high an information frequency.

The final extension to this work is that of \emph{Deep Generative Network Activation Maximization} (DGN-AM)\cite{Nguyena} which adds in a generative network, a network with the goal of producing an item of the specified class rather than to classify an object into a class, to produce more realistic images as opposed to the abstract ones often produced by activation maximisation. They were first used in 2016 and have shown greatly improved results, but also require greatly increased network complexity and training time. DGN-AMs have also been shown to provide worse outputs when the generative network is not the same architecture as the CNN and so multiple generative networks would need to be created for this work which would be far too computationally expensive; as such we chose to not use them within this work. An example taken from Qin's survey is shown in \cref{fig:am_dgnam} highlighting the huge change that using DGN-AMs produce.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.4]{am_vs_dgnam.PNG}
	\caption{Activation Maximization vs Deep Generative Network Activation Maximization from "How convolutional neural network see the world - A survey of convolutional neural network visualization methods" by Qin et al, 2018}
	\label{fig:am_dgnam}
\end{figure}
% section lit_rev (end)
\section{Proposal} % (fold)
\label{sec:proposal}
\subsection{Planned work}
The proposed work is to be completed in 3 distinct parts, starting with the work detailed in \cref{sub:part_0_data_collection_and_preparation} focusing on the collection and reprocessing of the data first generated by Iwana et al. \cite{KenjiIwana} to ensure completeness and correctness. We then in \cref{sub:part_1_genre_predictor} propose to reproduce the work of Iwana et al. and then develop this further using recent advances in CNN architecture. Finally \cref{sub:part_3_inverting_the_network} discusses the plans to use activation maximisation on the CNN to produce images based on the output class.
\subsubsection{Part 1: Data Collection and Preparation} % (fold)
\label{sub:part_0_data_collection_and_preparation}
The work first produced by Iwana\cite{KenjiIwana} uses a reduced dataset of 57,000 book covers that span 30 classes with 1,900 records each. We will first step back to the \emph{Book32} dataset to have available to us the full use of 207,572 books; we will download and ensure via some manual testing that the URLs of the images are still the correct ones from the time of writing.

At this stage any images will all be pre-processed including scaling to be a sensible size. The complexity is that of the shape of book covers in general, whilst there is some standardisation across the industry, is not fixed as shown in \cref{fig:book_cover_sizes} and specific genres of books such as cookbooks are likely to take a different shape due to their usage and content. This means that by scaling the images we are likely to impact some classes more than others. Analysis will be conducted to identify the distribution of shapes and if this varies by class, as well as a review of various methods used within the literature to approach such a problem. In particular an approach similar to the NIST dataset as reproduced by Cohen at al. \cite{Cohena} could be used with some adaptations for colour images.

Once this has been completed we will need to re-sample the dataset down to the same size as the original work, however we may not use the exact same subset of books they selected if we believe we can responsibly remove unrepresentative images such as the one displayed in \cref{fig:bad_cover}. This will lead to an inability for exact comparison between our work and that of Iwana, however it should help improve the accuracy of the model by removing non-representative training and test data. Where possible the same records will be used within the training and test sets, and only those that we identify to be non-standard will be swapped for another record.
\begin{figure}
	\centering
	\captionsetup{justification=centering,margin=2cm}
	\includegraphics[scale=0.5]{Comparison_book_sizes}
	\caption{Comparison of book sizes based on \cite{AmericanLibraryAssociation.CommitteeonLibraryTerminology.1971}\\
		\footnotesize{By Cmglee - Own work, CC BY-SA 3.0, https://commons.wikimedia.org/w/index.php?curid=15264645}}
	\label{fig:book_cover_sizes}
\end{figure}
% subsection part_0_data_collection_and_preparation (end)
\subsubsection{Part 2: Genre Predictor via Transfer Learning} % (fold)
\label{sub:part_1_genre_predictor}
In their work, Iwana et al\cite{KenjiIwana} trained 2 CNNs (LeNet and AlexNet) on the training set and report the Top 1 and Top 3 predicted class accuracy by book genre on the test set. We will attempt to reproduce the results achieved by them using AlexNet on our adjusted dataset and just use the LeNet results for an approximate comparison. We will use the same method as they did to train AlexNet i.e. apply transfer learning to a pre-trained version of the network on the ImageNet dataset, removing the classification layers of the network, those densely connected non-convolutional layers, including the  final 1000 neuron dense layer (representing the 1000 classes) and replace it with a similar network ending with a layer of 30 densely connected neurons instead, 1 for each of our genres, freezing the weights of the feature-extraction part of the network before beginning training.

Once a baseline has been established we will intended to improve the per-class accuracy by using some of the newer architectures as shown in \cref{fig:cnn_history}, namely Inception-ResnetV2 and a ResNeXt, and see if these yield better results. We will also look at using any of the techniques in the literature for differing shaped input images at this point to see if it is possible to provide not just the cover image to the network, but also information about the shape and size of it as well in the chance that this may improve accuracy. We again will use transfer learning here to greatly reduce the computational cost and time to train these models. The specific details of these CNN architectures will be discussed more within the work itself.

\begin{figure}
	\centering
	\captionsetup{justification=centering,margin=2cm}
	\includegraphics[scale=0.22]{CNN_HISTORY.png}
	\caption{Key CNN architectures and their year of publication\cite{Karim2019}}
	\label{fig:cnn_history}
\end{figure}
% subsection part_1_genre_predictor (end) 
\subsubsection{Part 3: Feature Visualisation}
\label{sub:part_3_inverting_the_network}
Once we have compared the performance of the different CNN architectures on the dataset we will attempt to use activation maximisation as discussed in \cref{sub:debug_cnn} with the network and instead of predicting the genre based on an image, it will produce an image based on a requested genre. This is sometimes also referred to as feature visualisation, although rather than individual features we are trying to reproduce a whole input. We will only take the best and worst 2-3 performing genres based on the previous stage to attempt this for.

To do this we propose the use of the relatively new and still under active development python package \emph{tf-keras-vis} \cite{keras-vis}. This package offers 2 key feature visualisation techniques; activation maximisation for filter layers and final output layers, and attention (i.e. what area of the image is driving the output class the network has predicted). Using this package we should be able to produce images that maximise the activation of our target class neuron. As mentioned in \cref{sub:debug_cnn} it is probable that these images will be abstract and may be entirely uninterpretable, but they may shed some insight on how specific genres are easier to predict than others and so we believe it adds value to the project.

\subsection{Hardware and Software}
Even using transfer learning, the amount of epochs the networks will have to train for to get a reasonable accuracy is still likely to be quite high, and as these are complex deep networks this will take a lot of time. Because of this we propose the use of Google Colab \cite{Google} to enable the use of GPUs, and potentially Tensor Processing Units (TPUs), for an order of magnitude speed up in the training of these networks. There is a limit on the amount of time a single session can run for, and there is no guarantee on the specific G/TPU brand/architecture that will be used, but without this or an approach similar it is unlikely that we posses the computing power to complete the work. As a point of note, TPUs are likely to offer a faster training however require a non-trivial amount of additional setup and so the choice to use them will be evaluated on a case by case basis. Because of this it is not possible to specify the exact hardware that we will use for this work.

For software, the majority of the work in terms of downloading, pre-processing, modelling, evaluations, and visualising features will be done in python using Tensorflow 2.x with a Keras backend, with heavy support from both the sklearn package and the aforementioned tf-keras-vis package. Visualisation outside of features may be done using R after some pre-analysis in python due to the extensive variety of libraries that support the production of graphics in R compared to python, as well as our own familiarity with the language.  
% section proposal (end)
\section{Conclusion} % (fold)
\label{sec:conclusion}
In this work we have laid out a problem which can be summarised as \emph{Can a CNN judge a book by its cover} and highlighted some of the previous work already completed in this space, namely that of Iwana et al. We discussed the existing available datasets, known as \emph{BookCover30} and \emph{Book32}, and explored some of the limitations this dataset currently has and we may potentially wish to address. Following this we laid out a 3 stage plan for trying to answer the question, starting with data collection and pre-processing, then using transfer learning to train high-performing CNN architectures for this problem and evaluating their performance. Finally we spoke of plans to use feature visualisation techniques to produce idealised covers for the most and least accurate genres from our models. Overall the proposed work should produce, as far as we can tell at time of writing, the only follow-up work to the Iwana paper and update their work using the more recent developments in CNN architecture and computing power.
% section conclusion (end)
%\nocite{*}
\bibliographystyle{ieeetran}
\bibliography{BBK_MSc_Project} 
\end{document}

### 2020-07-11 Tex/MSc Data Science Project.tex
Initial commit

--- 2020-07-11 Tex/MSc Data Science Project Proposal.tex
@@
 %
-\title{MSc Data Science Proposal: Can a Convolutional Neural Network judge a book by its cover?}
+\title{MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?}
 \author{\\Ryan Hill MSci\\\\
@@
 \begin{abstract}
-	In this work we make a proposal for training convolutional neural networks to predict the genre of a book, based on its cover.
-We evaluate the prior work done in this area and the proposed improvements of the existing dataset known as \emph{BookCover30}.
-We break the work into 3 stages and provide detail on the proposed hardware and software usage within the project.
+	Coming soon...
 \end{abstract}
@@
 \label{sec:intro}
-\subsection{Background} % (fold)
-\label{sub:background}
-\emph{Don't judge a book by its cover} is a common English idiom, with its origin somewhere around the mid 19\textsuperscript{th} century, often cited as being from the newspaper \emph{Piqua Democrat} as follows:
-\begin{quote}
-	``Don’t judge a book by its cover, see a man by his cloth, as there is often a good deal of solid worth and superior skill underneath a jacket and yaller pants.''
-\end{quote}
-The phrase itself refers to the idea that one should not judge an item, or person, by how it first appears, but instead you should take the time to understand it better.
-That being said, a huge amount of cost and effort can go into the design and choice of a cover for a book; costs for cover design alone can range from a few dollars to a few thousand \cite{AdriasG}, and revenue from publishing in 2018 was around \$122bn \cite{Watson}.
-This value, combined with the fact that humans have been shown to make judgments within 100ms of exposure to another person \cite{Willis2006}, means it is reasonable to assume that this advice is not often taken literally and that as humans we do, in fact, judge a book by its cover.
 
-Beyond the focus of the judgement placed on book covers, they are designed to help the potential reader understand the type of story, or contents, a book may contain.
-As an example, consider the covers shown in \cref{fig:2bookcovers} that should, even if they had no words on them, help the reader identify immediately the type of content contained within.
-The type of this content is what we refer to as the \emph{genre} of the book.
-Even within this example it is clear that there could be some subjectivity, as the reader you may see a dragon and assume it is a fantasy novel, but the Teen \& Young Adult genre often encompasses many genres just due to the age of the intended audience.
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[width=.8\linewidth]{200px-Eragon_book_cover.png}  
-	  \caption{Cover of Eragon, Teen \& Young Adult Genre}
-    \end{subfigure}%
-    ~ 
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[width=.8\linewidth]{cookbook_cover.jpg} 
-	  \caption{Cover of Good Food Cookbook, Cookbooks, Food \& Wine Genre}
-    \end{subfigure}
-	\caption{Two book covers with distinctly different designs and genres}
-	\label{fig:2bookcovers}
-\end{figure*}
+% section CNN Theory and Architectures
 
-% subsection background (end)
-\subsection{Problem Statement} % (fold)
-\label{sub:problem_statement}
-In this work we will try to answer one seemingly simple question; \emph{can a Convolutional Neural Network predict the genre of a book based solely upon an image of the front of the book?} Or put another way, \emph{can a Convolutional Neural Network judge a book by its cover?} This may seem like a fairly simple question, but in many cases it might already be quite difficult for a human to distinguish which genre a given book belongs to as just discussed.
-Beyond answering this question we will also attempt to take the easiest and hardest to identify genres (as determined by the accuracy of the classifier) and use feature visualisation to generate ideal covers for these genres, as the Convolutional Neural Network \emph{sees} them.
+\section{Data collection and pre-processing} 
+\label{sec:Data_collection_and_pre-processing} 
+\subsection{Manual review} 
+\label{sub:Manual_review} 
+Rule of thumb: single front facing book cover, book cover did not contain pictures of multiple books, and was not just an arrangement of book covers.
+Covers designed specifically for a boxset were allowed.
+% section Manual_review end 
+% section Data_Collection_and_pre-processing end 
+% Section Data collection and pre-processing and exploratory analysis
+    % include size heatmap
+    % include average image per class
 
-The rest of this work is structed as follows; first we review the existing literature and history of computer vision, predicting genre from book covers, and feature visualisation in \cref{sec:lit_rev}.
-We then discuss in more detail the plans for how to tackle this problem statement, including the hardware and software we plan to use, in \cref{sec:proposal}.
-Finally, we summarise this work in \cref{sec:conclusion}.
-% subsection problem_statement (end)
-% section intro (end)
-\section{Literature Review} % (fold)
-\label{sec:lit_rev}
+% Section Model Training and Results
 
-\subsection{History of Computer Vision}
-\label{sec:computer_vision}
-The work in this section is mostly informed by the comprehensive article by R Demush \cite{Demush}.
-Computer vision was originally, like most theoretical work in artificial intelligence (AI) and machine learning, conceived in the 1960s.
-A 1966 summer project by MIT was intending to solve the computer vision problem within a few months, simply put by attaching a camera to a computer it would then `describe what it saw' \cite{Papert1966}.
-Clearly, that was a little over-ambitious.
+% Section Feature Visualisation
 
-One of the most influential papers in computer vision actually came in 1959 when two neurophysiologists, Hubel and Wiesel, published their work studying the response of visual neurons of a cat's brain, and how these experiences shape the neural architecture \cite{Hubel1959}.
-Paraphrasing some of their results, they discovered that there are both simple and complex neurons, with visual processing beginning with simple structures such as edges and then going to more complex structures.
-This is essentially the basis behind deep learning today, especially in computer vision.
-However, even with this knowledge, it wasn't until 1982 when neuroscientist David Marr published the next piece of the puzzle in the form of a book entitled \emph{Vision: A Computational Investigation into the Human Representation and Processing of Visual Information}\cite{Marr:1982:VCI:1095712}.
-While his representational framework was abstract and high-level, it was still ground-breaking work and helped us understand how we create more complex images from the simple building blocks, and gave us a path to understanding how we could build computer models to do this.
+% Section Discussion and future work
 
-Around the same time \emph{Neocognitron}, a multilayer artificial neural net, was proposed by Kunihiko Fukushima which contained several convolution layers, the types of layers still used today.
-In 1989 Yann LeCun took this network and developed LeNet-5 from it, a Convolutional Neural Network (CNN) that contains many elements used by modern CNNs.
-This model had the goal of classifying images, as has become the main use of CNNs in the modern era.
-This work \cite{Cun1989} also included the production of the now famous MNIST dataset of handwritten digits.
+% Section Conclusion
 
-For a few years, work continued to try and construct 3D models from 2D images, before instead moving towards object recognition around the turn of the new millennia.
-Development continued over the coming years, including a new focus on having a set of communal benchmark datasets to enable easy comparison.
-These included the original MNIST, before the development of Pascal VOC \cite{Everingham10}, and finally ImageNet \cite{imagenet_cvpr09}in 2010.
-ImageNet has now become the de-facto benchmark dataset for image classification with over a million manually cleaned images covering 1000 object classes.
-At this point, development started coming specifically within the architecture of CNNs, which is discussed briefly in \cref{sub:cnn_architecture}.
-
-\subsection{Prior Work} % (fold)
-\label{sub:prior_work}
-Design has been explored in many areas before with regards to its history and the various styles, however it is a relatively new area for machine learning in comparison to other fields.
-CNNs have been used to identify the artists of works \cite{Karayev2013}, they have been suggested for use in forgery detection \cite{Gideon2018}, and with the advances in Generative Adversarial Networks (GANs) we can also produce realistic examples or enhance images of a given classification \cite{Ledig2016}.
-
-CNNs have also been used to help classify artwork \cite{Zujovic2009}, clothing types\cite{Xiao2017}, and music \cite{Tzanetakis2002}; but only Iwana et al.
-\cite{KenjiIwana} have applied this work to book covers to attempt to classify the genre so far as we can tell.
-It is their work that we propose to build on.
-They produced a cleaned dataset, discussed further in \cref{sub:existing_datasets}, of 57,000 images spanning 30 equally represented genres and presented the classification results of an unseen test set of 10\% of the data after training AlexNet, a specific CNN architecture.
-They report a Top 1 accuracy of 24.7\% and a Top 3 accuracy of 40.3\%  across the 30 classes (i.e.
-a random chance of 3.3\% and 9.9\% respectively).
-Class specific Top 1 accuracies reach up to 68.9\% for the Test Preparation genre, but are as low as 6.8\% for the Politics \& Social Sciences genre.
-We will explore the results of their work further, and discuss the potential reasoning for this range of performance by class if we see similar results, in our main work.
-% subsection prior_work (end)
-\subsection{CNN Architectures and Transfer Learning}
-\label{sub:cnn_architecture}
-For all of the development of CNNs discussed in this section, they continue to maintain the same overall structure they started with, that of having a clear delineation on the network where they go from feature extraction to classification as shown in \cref{fig:cnn_struct}.
-The first part of the network consists of various convolution/filter and pooling layers, used to extract image features (e.g.
-edges) and to reduce the image dimension respectively.
-The second part of the network is a traditional densely connected artificial neural network used for the actual classification; a few fully connected layers followed by the output/target layer with as many neurons are there classes in the problem.
-Much of the development has gone into the feature extraction part of the network over the last 8 years or so; starting in 2012 with AlexNet \cite{Krizhevsky2017}, with the paper hailed as one of the most influential papers in applied computer vision.
-A few years later these were joined by what have been come to be known as the VGG CNNs (named after the Visual Geometry Group at Oxford)\cite{Simonyan2015}.
-The following year the family of Inception models (also known as GoogLeNet) start to appear \cite{Szegedy}; around the exact same time ResNets were proposed by He et al.
-at Microsoft \cite{He2015} which later led to the combination of these in 2017.
-We do not detail here the specifics of CNN layers, including the types of convolution or pooling, nor the differences in all these architectures as we will detail these in the work itself, but those chosen to be used within this work are mentioned in \cref{sub:part_1_genre_predictor}.
-
-\begin{figure}
-	\centering
-	\captionsetup{justification=centering}
-	\includegraphics[scale=0.3]{cnn_struct.jpeg}
-	\caption{An simplified example of a CNN architecture}
-	\label{fig:cnn_struct}
-\end{figure}
-
-We will be using transfer learning to reduce the effort involved in training our network.
-Transfer learning was first proposed in 1993 by Lorien Pratt \cite{Pratt1993} and has been said by Andrew Ng to be the next driver for commercial machine learning success.
-A comprehensive survey of transfer learning has been completed by Zhuang et al.
-\cite{Zhuang2019} so we do not detail here all of the developments and differences in various transfer learning algorithms, especially as we plan to take the most basic approach.
-Whilst many have worked to develop detailed algorithms that take into account the overlap of the source data domain and the target data domain and adjust a network accordingly, we are taking a more simple approach that is discussed in \cref{sub:part_1_genre_predictor} and reusing all but the (dense) classification layers of a network pretrained on ImageNet as the starting point for our training.
-% subsection cnn_architecture (end)
-\subsection{Existing Datasets} % (fold)
-\label{sub:existing_datasets}
-The dataset collected by Iwana et al.\cite{KenjiIwana} is the largest, and as far as we can tell only, clean dataset to collate together the book covers of over tens of thousands of books.
-The dataset exists in two distinct forms; the \emph{BookCover30} dataset and the \emph{Book32} dataset that contain 57,000 and 207,572 records respectively, split across 30 and 32 classes again respectively.
-Both datasets have the same columns; \emph{Amazon Index, Filename, Image URL, Title, Author, Category Id, Category}.
-For \emph{BookCover30} there is also a reduced dataset containing just the filename and category ID that can be used for training and testing after image download.
-A comparison of the volumes, as well as the split between train and test sets, and a key between the category ID and category label can be seen in \cref{tab:book_data_comparison}.
-In the case that a book belonged to multiple genres, the reported genre was chosen at random.
-
-\begin{table}[]
-	\centering
-	\begin{tabular}{|l|l|l|l|l|}
-	\hline
-	\textbf{Label} & \textbf{Category Name}        & \textbf{\begin{tabular}[c]{@{}l@{}}Size\\ Book32\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Size \\ BookCover30\\ Training\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Size \\ BookCover30\\ Test\end{tabular}} \\ \hline
-	0              & Arts \& Photography           & 6,460                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	1              & Biographies \& Memoirs        & 4,261                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	2              & Business \& Money             & 9,965                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	3              & Calendars                     & 2,636                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	4              & Children's Books              & 13,605                                                         & 1710                                                                            & 190                                                                         \\ \hline
-	5              & Comics \& Graphic Novels      & 3,026                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	6              & Computers \& Technology       & 7,979                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	7              & Cookbooks, Food \& Wine       & 8,802                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	8              & Crafts, Hobbies \& Home       & 9,934                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	9              & Christian Books \& Bibles     & 9,139                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	10             & Engineering \& Transportation & 2,672                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	11             & Health, Fitness \& Dieting    & 11,886                                                         & 1710                                                                            & 190                                                                         \\ \hline
-	12             & History                       & 6,807                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	13             & Humor \& Entertainment        & 6,896                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	14             & Law                           & 7,314                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	15             & Literature \& Fiction         & 7,580                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	16             & Medical Books                 & 12,089                                                         & 1710                                                                            & 190                                                                         \\ \hline
-	17             & Mystery, Thriller \& Suspense & 1,998                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	18             & Parenting \& Relationships    & 2,523                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	19             & Politics \& Social Sciences   & 3,402                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	20             & Reference                     & 3,268                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	21             & Religion \& Spirituality      & 7,559                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	22             & Romance                       & 4,291                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	23             & Science \& Math               & 9,276                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	24             & Science Fiction \& Fantasy    & 3,800                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	25             & Self-Help                     & 2,703                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	26             & Sports \& Outdoors            & 5,968                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	27             & Teen \& Young Adult           & 7,489                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	28             & Test Preparation              & 2,906                                                          & 1710                                                                            & 190                                                                         \\ \hline
-	29             & Travel                        & 18,338                                                         & 1710                                                                            & 190                                                                         \\ \hline
-	30             & Gay \& Lesbian                & 1,339                                                          & 0                                                                               & 0                                                                           \\ \hline
-	31             & Education \& Teaching         & 1,664                                                          & 0                                                                               & 0                                                                           \\ \hline
-	\end{tabular}
-	\caption{Comparison of the number of records per category between the Book32 and Book30 training and test sets}
-	\label{tab:book_data_comparison}
-	\end{table}
-
-The reduced \emph{BookCover30} dataset has already been pre-processed to remove 2 underrepresented classes, and to scale and shape the images; this will likely have to be re-sourced before the processing has taken place as the various CNNs we are going to use have different input shapes.
-There is also some other issues with the dataset; in particular that no manual filtering was done to ensure that the images are representative of the actual book cover, an example where this is not the case can be seen in \cref{fig:bad_cover}.
-The data also looks to cover multiple languages so it is possible that there will be duplication of items (although often foreign editions of books have different covers) and this could also lead to additional difficulty if there is a difference in design choices between different countries.
-Furthermore, this dataset and paper were first published in 2016; while using the same dataset would be useful for comparison it is possible that the same script they used to generate their data could be repurposed to allow for a more up-to-date dataset to be produced, with higher resolution images and a potentially more representative sample of modern books.
-More discussion on this is repeated in \cref{sub:part_0_data_collection_and_preparation}.
-
-We mention here for completeness that during our research we did come across a potential alternative source of data, that of the Open Library \cite{openAbout} who as part of the Internet Archive project have the goal to have\emph{"One web page for every book ever published"}.
-The regular data dumps \cite{openDataDumps} are a possible alternative source to using the Iwana dataset, however this dataset would come with some uncertainty attached.
-Whilst the volume of records is undoubtedly larger in size, it is not clear without processing all of the records how many of these actually have a genre classification at all, and given that their book covers data dump has not been updated since 2016 it does not provide any more recent data than Iwana's does.
-We would also be remiss to not mention that as this is an open-source project, the quality of the data itself would be more in question than with the \emph{Book32} records.
-Finally, it would remove our ability to use Iwana et al's paper as a reference in terms of performance of the networks as not only would the training data itself be vastly different, but the genres are also different in this source.
-
-Overall the \emph{Book32} dataset is the best quality and most label-complete dataset available that we could find, and to produce a new one from scratch even using Open Library as a foundation would take a large amount of work for likely very little gain, so it is this dataset we will use in our work.
-\begin{figure}
-	\centering
-	\captionsetup{justification=centering}
-	\includegraphics[scale=0.4]{0345535529.jpg}
-	\caption{An image in the \emph{BookCover30} dataset showing an anthology box collection rather than the cover}
-	\label{fig:bad_cover}
-\end{figure}
-% subsection existing_datasets (end)
-\subsection{Debugging CNNs}
-\label{sub:debug_cnn}
-CNNs, along with most other deep-learning methods, have been seen for a long time as \emph{black box} methods i.e.
-it is not possible to understand what a specific neuron or layer within the network is doing.
-The attitude to black box methods has soured over the years for 2 major reasons.
-The first being that black box methods are illegal in some fields such as underwriting; a lender must be able to explain to an applicant why they were rejected for a loan, and black box methods do not allow for this.
-The second is that they are almost impossible to debug and understand where the network may be making an error.
-Many examples of this were first shown by Nguyen et al \cite{Nguyen} in 2015, where images could be evolved to trick CNNs that return high precision results for obviously garbage images.
-More recently Su et al.
-\cite{Su} managed to show that perturbation of even a single pixel in an image can generate a change in the predicted class from otherwise high performing classifiers.
-These examples, combined with a drive for more explainable AI, has led to a few key developments in the debugging of CNNs via representative visualisation.
-
-In 2018 Qin et al.
-\cite{Qin2018a} produced a survey of the major visualisation methods and whilst there has been some developments, mostly in the implementation of these methods in common software packages, their work remains a good summary of the state of the work.
-They call out four key visualisation methods:
-\begin{enumerate}
-	\item Activation maximisation - A visualised image is generated such that it maximises a specific neuron's activation.
-	\item DeconNet - Uses an inverted CNN structure to identify the pattern in the input image that triggers specific neuron activation.
-	\item Network Inversion - Recreates an image based on the input image from a specific layer, helping identify what information is preserved by this layer.
-	\item Network dissection  - Applies semantic concepts to each neuron such as scene, material, object etc.
-\end{enumerate}
-For the purpose of this work we focus only on activation maximisation as this is what we plan to use to generate representative images of some of the genres.
-
-Activation maximisation was proposed by Erhan et al.
-\cite{erhan2009} in 2009 and the idea itself is quite simple; iteratively perturbate an image to maximise the activation of a given neuron.
-The resultant image then represents what that neuron has learned.
-The actual algorithm uses backpropagation instead of perturbation to improve the synthesised image over many iterations.
-Compared to normal backpropagation where we calculate the change in the network error with respect to each weight and bias and use that to update the values, here we calculate the change in activation with respect to the input and use this as the base for our update rule:
-\begin{equation}
-	input_{t} = input_{t-1} - \eta\times\frac{\partial ActivationMaximisationLoss}{\partial input_{t-1}}
-\end{equation} 
-where $ActivationMaximisationLoss$ is some loss measure that decreases as the activation increases, and $\eta$ is a learning rate.
-We stop when the activation of the neuron cannot be increased further.
-This can be applied to almost any neuron in the network with no alteration to the structure except for those in the final layer.
-The final layer activation function, the transformation applied to the inputs of the neuron, must be changed from the pre-existing SoftMax ($\sigma$) as defined in \cref{eq:softmax}, which takes a vector of activations for the $K$ output classes and normalises them to be between 0 and 1 with a sum of 1, and be replaced with a linear function instead (often just the linear sum of inputs is used).
-The reason for this is that to maximise the output of a SoftMax function for a single neuron, it is possible to minimise the values of the other neurons.
-This leads to a case where the generated image would not be representative of our target class, but just be more representative than it is of the other classes.
-By using the unbounded linear function we can continue to apply updates to the input until we see little to no increase in the activation for our target neuron, irrespective of the other class activations.
-
-\begin{equation}
-	\sigma(\vec{x})_i = \frac{e^{x_i}}{\sum_{j=1}^{K}e^{x_j}}
-	\label{eq:softmax}
-\end{equation}
-
-Often, this approach is taken together with some form of normalisation to improve the interpretability of these images.
-As mentioned, pictures that are composed entirely of noise can trigger activation in the output neurons in a network with high certainty, so these normalisation methods are applied to reduce the production of these types of images and return more visually understandable results.
-Some of these normalisations, such as $l_2$ decay (which usually uses the sum of squared weight, but pixel values in this case) are used to prevent a few pixels dominating the image, or applying a Gaussian Blur can be done to ensure that no area of the image has too high an information frequency.
-
-The final extension to this work is that of \emph{Deep Generative Network Activation Maximization} (DGN-AM)\cite{Nguyena} which adds in a generative network, a network with the goal of producing an item of the specified class rather than to classify an object into a class, to produce more realistic images as opposed to the abstract ones often produced by activation maximisation.
-They were first used in 2016 and have shown greatly improved results, but also require greatly increased network complexity and training time.
-DGN-AMs have also been shown to provide worse outputs when the generative network is not the same architecture as the CNN and so multiple generative networks would need to be created for this work which would be far too computationally expensive; as such we chose to not use them within this work.
-An example taken from Qin's survey is shown in \cref{fig:am_dgnam} highlighting the huge change that using DGN-AMs produce.
-
-\begin{figure}
-	\centering
-	\captionsetup{justification=centering}
-	\includegraphics[scale=0.4]{am_vs_dgnam.PNG}
-	\caption{Activation Maximization vs Deep Generative Network Activation Maximization from "How convolutional neural network see the world - A survey of convolutional neural network visualization methods" by Qin et al, 2018}
-	\label{fig:am_dgnam}
-\end{figure}
-% section lit_rev (end)
-\section{Proposal} % (fold)
-\label{sec:proposal}
-\subsection{Planned work}
-The proposed work is to be completed in 3 distinct parts, starting with the work detailed in \cref{sub:part_0_data_collection_and_preparation} focusing on the collection and reprocessing of the data first generated by Iwana et al.
-\cite{KenjiIwana} to ensure completeness and correctness.
-We then in \cref{sub:part_1_genre_predictor} propose to reproduce the work of Iwana et al.
-and then develop this further using recent advances in CNN architecture.
-Finally \cref{sub:part_3_inverting_the_network} discusses the plans to use activation maximisation on the CNN to produce images based on the output class.
-\subsubsection{Part 1: Data Collection and Preparation} % (fold)
-\label{sub:part_0_data_collection_and_preparation}
-The work first produced by Iwana\cite{KenjiIwana} uses a reduced dataset of 57,000 book covers that span 30 classes with 1,900 records each.
-We will first step back to the \emph{Book32} dataset to have available to us the full use of 207,572 books; we will download and ensure via some manual testing that the URLs of the images are still the correct ones from the time of writing.
-
-At this stage any images will all be pre-processed including scaling to be a sensible size.
-The complexity is that of the shape of book covers in general, whilst there is some standardisation across the industry, is not fixed as shown in \cref{fig:book_cover_sizes} and specific genres of books such as cookbooks are likely to take a different shape due to their usage and content.
-This means that by scaling the images we are likely to impact some classes more than others.
-Analysis will be conducted to identify the distribution of shapes and if this varies by class, as well as a review of various methods used within the literature to approach such a problem.
-In particular an approach similar to the NIST dataset as reproduced by Cohen at al.
-\cite{Cohena} could be used with some adaptations for colour images.
-
-Once this has been completed we will need to re-sample the dataset down to the same size as the original work, however we may not use the exact same subset of books they selected if we believe we can responsibly remove unrepresentative images such as the one displayed in \cref{fig:bad_cover}.
-This will lead to an inability for exact comparison between our work and that of Iwana, however it should help improve the accuracy of the model by removing non-representative training and test data.
-Where possible the same records will be used within the training and test sets, and only those that we identify to be non-standard will be swapped for another record.
-\begin{figure}
-	\centering
-	\captionsetup{justification=centering,margin=2cm}
-	\includegraphics[scale=0.5]{Comparison_book_sizes}
-	\caption{Comparison of book sizes based on \cite{AmericanLibraryAssociation.CommitteeonLibraryTerminology.1971}\\
-		\footnotesize{By Cmglee - Own work, CC BY-SA 3.0, https://commons.wikimedia.org/w/index.php?curid=15264645}}
-	\label{fig:book_cover_sizes}
-\end{figure}
-% subsection part_0_data_collection_and_preparation (end)
-\subsubsection{Part 2: Genre Predictor via Transfer Learning} % (fold)
-\label{sub:part_1_genre_predictor}
-In their work, Iwana et al\cite{KenjiIwana} trained 2 CNNs (LeNet and AlexNet) on the training set and report the Top 1 and Top 3 predicted class accuracy by book genre on the test set.
-We will attempt to reproduce the results achieved by them using AlexNet on our adjusted dataset and just use the LeNet results for an approximate comparison.
-We will use the same method as they did to train AlexNet i.e.
-apply transfer learning to a pre-trained version of the network on the ImageNet dataset, removing the classification layers of the network, those densely connected non-convolutional layers, including the  final 1000 neuron dense layer (representing the 1000 classes) and replace it with a similar network ending with a layer of 30 densely connected neurons instead, 1 for each of our genres, freezing the weights of the feature-extraction part of the network before beginning training.
-
-Once a baseline has been established we will intended to improve the per-class accuracy by using some of the newer architectures as shown in \cref{fig:cnn_history}, namely Inception-ResnetV2 and a ResNeXt, and see if these yield better results.
-We will also look at using any of the techniques in the literature for differing shaped input images at this point to see if it is possible to provide not just the cover image to the network, but also information about the shape and size of it as well in the chance that this may improve accuracy.
-We again will use transfer learning here to greatly reduce the computational cost and time to train these models.
-The specific details of these CNN architectures will be discussed more within the work itself.
-
-\begin{figure}
-	\centering
-	\captionsetup{justification=centering,margin=2cm}
-	\includegraphics[scale=0.22]{CNN_HISTORY.png}
-	\caption{Key CNN architectures and their year of publication\cite{Karim2019}}
-	\label{fig:cnn_history}
-\end{figure}
-% subsection part_1_genre_predictor (end) 
-\subsubsection{Part 3: Feature Visualisation}
-\label{sub:part_3_inverting_the_network}
-Once we have compared the performance of the different CNN architectures on the dataset we will attempt to use activation maximisation as discussed in \cref{sub:debug_cnn} with the network and instead of predicting the genre based on an image, it will produce an image based on a requested genre.
-This is sometimes also referred to as feature visualisation, although rather than individual features we are trying to reproduce a whole input.
-We will only take the best and worst 2-3 performing genres based on the previous stage to attempt this for.
-
-To do this we propose the use of the relatively new and still under active development python package \emph{tf-keras-vis} \cite{keras-vis}.
-This package offers 2 key feature visualisation techniques; activation maximisation for filter layers and final output layers, and attention (i.e.
-what area of the image is driving the output class the network has predicted).
-Using this package we should be able to produce images that maximise the activation of our target class neuron.
-As mentioned in \cref{sub:debug_cnn} it is probable that these images will be abstract and may be entirely uninterpretable, but they may shed some insight on how specific genres are easier to predict than others and so we believe it adds value to the project.
-
-\subsection{Hardware and Software}
-Even using transfer learning, the amount of epochs the networks will have to train for to get a reasonable accuracy is still likely to be quite high, and as these are complex deep networks this will take a lot of time.
-Because of this we propose the use of Google Colab \cite{Google} to enable the use of GPUs, and potentially Tensor Processing Units (TPUs), for an order of magnitude speed up in the training of these networks.
-There is a limit on the amount of time a single session can run for, and there is no guarantee on the specific G/TPU brand/architecture that will be used, but without this or an approach similar it is unlikely that we posses the computing power to complete the work.
-As a point of note, TPUs are likely to offer a faster training however require a non-trivial amount of additional setup and so the choice to use them will be evaluated on a case by case basis.
-Because of this it is not possible to specify the exact hardware that we will use for this work.
-
-For software, the majority of the work in terms of downloading, pre-processing, modelling, evaluations, and visualising features will be done in python using Tensorflow 2.x with a Keras backend, with heavy support from both the sklearn package and the aforementioned tf-keras-vis package.
-Visualisation outside of features may be done using R after some pre-analysis in python due to the extensive variety of libraries that support the production of graphics in R compared to python, as well as our own familiarity with the language.
-% section proposal (end)
-\section{Conclusion} % (fold)
-\label{sec:conclusion}
-In this work we have laid out a problem which can be summarised as \emph{Can a CNN judge a book by its cover} and highlighted some of the previous work already completed in this space, namely that of Iwana et al.
-We discussed the existing available datasets, known as \emph{BookCover30} and \emph{Book32}, and explored some of the limitations this dataset currently has and we may potentially wish to address.
-Following this we laid out a 3 stage plan for trying to answer the question, starting with data collection and pre-processing, then using transfer learning to train high-performing CNN architectures for this problem and evaluating their performance.
-Finally we spoke of plans to use feature visualisation techniques to produce idealised covers for the most and least accurate genres from our models.
-Overall the proposed work should produce, as far as we can tell at time of writing, the only follow-up work to the Iwana paper and update their work using the more recent developments in CNN architecture and computing power.
-% section conclusion (end)
 %\nocite{*}

### 2020-07-11 Tex/project.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mdframed}
\usepackage[margin=3cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\DeclareMathOperator{\Lagr}{\mathcal{L}}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage{hyperref}
\hypersetup{pdfpagemode=FullScreen,  
colorlinks=true}

\begin{document}
\renewcommand\citeform[1]{[#1]}

\title{On Higher Dimensional Black Holes and the Instability of the Extremal Reissner-Nordstr\"om}
\author{\\Ryan Hill \\\\
Supervisor: Dr Rodolfo Russo\\\\
Queen Mary, University Of London\\\\
\texttt{r.l.hill@se12.qmul.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
%
%
\graphicspath{{images/}}
%
\begin{abstract}
We study the derivation of the Schwarzschild and Reissner-Nordstr\"om (RN) black hole metrics in both four and $D$ dimensions before discussing the application of thermodynamics to these solutions. We then discuss the instability of the extremal RN black hole in four dimensions, before considering the same case for a five dimensional metric as a specific reduction of the Strominger-Vafa black hole; finding the same instability present in this as well. 
\end{abstract}
%
\clearpage
%
{\hypersetup{linkcolor=black}
\tableofcontents}
\thispagestyle{empty}
%
\clearpage
%
\setcounter{page}{1}
\section{Introduction}
The theoretical study of black holes is important to the understanding of General Relativity (GR), making predictions for observational experiments, and even aiding in the understanding of a quantum theory of gravity. This research project will focus on the study of two types of black holes, the Schwarzschild and Reissner-Nordstr\"om (RN) solutions, first deriving the metric for a general spherical solution in 4-dimensions in section.\ref{sub:general_spherical_solution} and then specialising to Schwarzschild solution in section.\ref{sub:schwarzschild_solution} before next deriving the RN solution in section.\ref{sub:reissner_nordstr"om_solution}. Then we will repeat the process for in D-dimensions; again starting with the general spherically symmetric case in section.\ref{sub:general_spherically_symmetric_solutions} before specialising once more to the Schwarzschild in section.\ref{sub:schwarzschild_solution_in_extra_dimensions}, and finally the RN in section.\ref{sub:reissner_nordstom_soltuion_in_extra_dimensions}.

We will then move on to look at the thermodynamics of these systems in section.\ref{sec:blackhole_thermodynamics} and we discuss how we can assign temperature and entropy to the horizon of a black hole. Next we move onto look at the instability of the extremal ($T=0$) RN black hole and how to show this explicitly in section.\ref{sec:intability_of_extremal_rn_black_hole} before finally doing the same but for a 5-D RN metric reduced from the Strominger-Vafa line element and showing the the instability still exists here in section.\ref{sec:5d_charged_black_holes_from_string_theory}. The reason we want to look at this further metric is there exists addition (super)symmetries in this case and it may have been that this did not give rise to the instability; however this is not the case and the instability remains. Finally we review the complete work in section.\ref{sec:conclusion} as well as providing some additional published results for further reading and then discussing possible continuations/extensions of this work.

A solid foundation of GR is already expected of the reader and, whilst not necessary, a basic understanding of differential forms will be particularly useful for the derivation of the higher dimensional metrics. We will be working mostly in natural units where $\hbar = c =\mu_{0} = \epsilon_{0} = k_b = 1$ where the only exception is that $G_{N}\ne 1$. For completeness a review of GR is provided in the next section. 
%
%
\subsection{Whistle-stop Review of General Relativity}
%Write here about what is required and why we only cover basic stuff
\subsubsection{Tensors} % (fold)
Tensors are an important part of GR and as such the review of them here will be brief as we expect the reader to have a solid foundation in them already. Tensors are mathematical objects of possibly multiple dimensions that leave vectors unchanged under a coordinate transform; they will change the components of the vector, but the overall vector is invariant. This can be written explicitly as a \emph{Tensor Transformation Law}
\begin{equation}
	{T^{\prime\mu}}_{\nu\rho} = \frac{\partial x^{\prime\mu}}{\partial x^\sigma}\frac{\partial x^\tau}{\partial x^{\prime\nu}}\frac{\partial x^\alpha}{\partial x^{\prime\rho}}{T^\sigma}_{\tau\alpha}
\end{equation}
where $x^\mu$ are the original coordinates and $x^{\prime\mu}$ are the transformed coordinates. This was an example but can be extrapolated to higher or lower rank tensors, some examples of which include
\begin{itemize}
	\item Rank 0: Scalars
	\item Rank 1: Vectors 
	\item Rank 2: 2-dimensional matrix
\end{itemize}
and so on. It is important to note that these are just examples of a tensor with such a rank but are not all of those rank tensors. For example, all vectors are rank 1 tensors but not all rank 1 tensors are vectors, i.e. vectors $\subset$ rank 1 tensors. The reason we use these objects is that the laws of physics must be invariant in all frames, rather than just the inertial ones required for Special Relativity, and as such GR is formulated using these objects to allow us to consider accelerating frames.
%
%
% subsubsection tensors (end)
\subsubsection{The Metric} % (fold)
\label{subsub:the_metric}
In General Relativity (GR) the metric is the most fundamental tensor; it describes a specific spacetime and contains information mainly relating to distances and causality for a specific coordinate system. For any general metric it will be denoted $g_{\mu\nu}$, where $\mu,\nu\ =0,1 \ldots D-1$ and $D$ is the dimension of the spacetime. The flat space metric (also known as the Minkowski Metric) will be denoted with $\eta_{\mu\nu}$ and in $D=4$ 
\begin{equation}
	\eta_{\mu\nu} = 
	\begin{pmatrix}
		-1 & 0 & 0 & 0 \\
		0 & 1 & 0 & 0 \\
		0 & 0 & 1 & 0 \\
		0 & 0 & 0 & 1 
	\end{pmatrix}
\end{equation}
where we have chosen the $(- + + + )$ convention as most GR work uses. The metric is a symmetric $D\times D$ matrix and as such $g_{\mu\nu} = g_{\nu\mu}$, a property we will use often. The next item we need to discuss is the line element
\begin{equation}
	ds^2 = dx^\mu dx^\nu g_{\mu\nu}
\end{equation}
where $x^\mu$ are the coordinates in that spacetime, $ds^2$ is an interval between two points in spacetime and is invariant under coordinate transformations i.e
\begin{equation}
	ds'^2=ds^2
\end{equation}
where $ds'^2$ is the line element in a new coordinate system.
% subsubsection the_metric (end)
\subsubsection{The Christoffel Symbols} % (fold)
\label{subsub:the_christoffel_symbolse}
A problem arises when we want to differentiate tensors as the usual partial derivative of a tensor does not transform as a tensor, and as such we use the covariant derivative, $D_\mu$. This introduces a different problem, how do we define $D_\mu$ such that it transforms as a tensor? If we examine how the partial derivatives transforms it is trivial to show that 
\begin{align}
	\partial_\mu V^\nu &\rightarrow \frac{\partial}{\partial x^{\prime\mu}} V^\nu(x^{\prime\mu})\\
	&= \frac{\partial x^\rho}{\partial x^{\prime\mu}} \frac{\partial x^{\prime\nu}}{\partial x^\sigma} \frac{\partial V^\sigma}{\partial x^\rho} + \frac{\partial x^\rho}{\partial x^{\prime\mu}}\frac{\partial^2 x^{\prime\nu}}{\partial x^\rho \partial x^\sigma}V^\sigma
\end{align}
where the second term is obviously not obeying the tensor transformation law. The obvious solution to this is to define the covariant derivative with a term that, under a change of coordinates, exactly cancels the second term. Doing this means that we would need
\begin{equation}
	\Gamma^{\prime\nu}_{\mu\rho} = \frac{\partial x^\sigma}{\partial x^{\prime\mu}} \frac{\partial x^\tau}{\partial x^{\prime\rho}} \frac{\partial x^{\prime\nu}}{\partial x^\omega}\Gamma^{\omega}_{\sigma\tau} - \frac{\partial x^\tau}{\partial x^{\prime\mu}}\frac{\partial^2 x^{\prime\nu}}{\partial x ^{\prime\rho}  \partial x^\tau}. \label{eq:chris_satisfy}
\end{equation}
There are many possible choices of this term, known as a connection, that would satisfy this equation; one such choice is the Levi-Civita connection
\begin{equation}
	\Gamma^{\mu}_{\nu\rho}=\frac{1}{2}g^{\mu\sigma}[\partial_\rho g_{\sigma\nu} + \partial_\nu g_{\sigma\rho} - \partial_\sigma g_{\nu\rho}], \label{eq:Christoffel_symbol}
\end{equation}
which we shall call the Christoffel symbol, where I have verified independently that this satisfies (\ref{eq:chris_satisfy}). Now we have defined this connection we can write the general form of the covariant derivative acting on a rank $m+n$ tensor as 
\begin{multline}
	D_\mu {A^{\nu_1,\nu_2,...,\nu_m}}_{\sigma_1,\sigma_2,\dots,\sigma_n}  = \partial_\mu{A^{\nu_1,\nu_2,...,\nu_m}}_{\sigma_1,\sigma_2,\dots,\sigma_n} \\+ \Gamma^{\nu_1}_{\mu\rho}{A^{\rho,\nu_2,\dots,\nu_m}}_{\sigma_1,\sigma_2,\dots,\sigma_n} + \Gamma^{\nu_2}_{\mu\rho}{A^{\nu_1,\rho,...,\nu_m}}_{\sigma_1,\sigma_2,\dots,\sigma_n} + \dots \\
	 - \Gamma^{\rho}_{\mu\sigma_1}{A^{\nu_1,\nu_2,\dots,\nu_m}}_{\rho,\sigma_2,\dots,\sigma_n} - \Gamma^{\rho}_{\mu\sigma_2}{A^{\nu_1,\nu_2,\dots,\nu_m}}_{\sigma_1,\rho,\dots,\sigma_n} - \dots.
\end{multline}
Finally we can define the geodesic equation, that is the path between two points in a spacetime that has a constant tangent vector. Said another way it is the shortest path between two points e.g. in flat space this is a straight line. The geodesic equation can be derived by considering the requirement that the tangent vector be constant and then \emph{covariantising} the derivatives. Expanding the relevant terms we get 
\begin{equation}
	\frac{d^2x^\mu}{ds^2} + \Gamma^{\mu}_{\nu\rho}\frac{dx^\nu}{ds}\frac{dx^\rho}{ds} = 0.
\end{equation}
% subsubsection the_christoffel_symbols (end)
\subsubsection{Curvature} % (fold)
\label{subsub:curvature}
We now finally move on to the tensors needed to describe the curvature of a given spacetime, the first is the Riemann Curvature Tensor, ${R^\rho}_{\mu\nu\sigma}$ which is defined as
\begin{align}
	[D_\nu,D_\sigma]A^\rho &= {R^\rho}_{\mu\nu\sigma} A^\mu \\
	\implies {R^\rho}_{\mu\nu\sigma} &= \partial_\nu \Gamma^{\rho}_{\sigma\mu} - \partial_\sigma \Gamma^{\rho}_{\nu\mu} + \Gamma^{\rho}_{\nu \lambda}\Gamma^{\lambda}_{\sigma\mu} - \Gamma^{\rho}_{\sigma \lambda}\Gamma^{\lambda}_{\nu \mu} \label{eq:riemann}
\end{align}
where the intermediate steps are tedious but trivial. We can also define the entirely covariant form of this as
\begin{equation}
 	R_{\rho\mu\nu\sigma}=g_{\rho\lambda}{R^\rho}_{\mu\nu\sigma}= \frac{1}{2}\left(\partial_{\mu}\partial_{\nu}g_{\sigma\rho}-\partial_{\nu}\partial_{\rho}g_{\mu\sigma} -\partial_{\mu}\partial_{\sigma}g_{\nu\rho} + \partial_{\sigma}\partial_{\rho}g_{\sigma\nu}  \right) + g_{\rho\lambda}\left(\Gamma^{\lambda}_{\nu \alpha}\Gamma^{\alpha}_{\sigma\mu} - \Gamma^{\lambda}_{\sigma \alpha}\Gamma^{\alpha}_{\nu \mu}\right)
\end{equation}
where we use the definition of the Christoffel from (\ref{eq:Christoffel_symbol}) to simplify the first two terms. As this is a rank-4 tensor it has $D^4$ components which even in a 4-dimensional space time means 256 components; fortunately we can see from the covariant form that this tensor has some properties which greatly reduce the number of unique components and these are
\begin{align}
	R_{\mu\nu\sigma\rho} &= -R_{\nu\mu\sigma\rho} = -R_{\mu\nu\rho\sigma} \label{eq:riemann-prop-1}\\
	R_{\mu\nu\sigma\rho} &= R_{\sigma\rho\mu\nu} \\
	R_{\mu\nu\rho\sigma} &+ R_{\mu\rho\sigma\nu} + R_{\mu\sigma\nu\rho} = 0 \label{eq:riemann-prop-3}
\end{align}
which, again in 4 dimensions, brings the number of unique components down to 20.
We also have the Ricci tensor, which is a rank-2 tensor and is a contraction of the Riemann tensor of the form
\begin{equation}
	R_{\mu\nu} = {R^\lambda}_{\mu\lambda\nu}. \label{eq:ricci}
\end{equation}
Then finally we have the Ricci scalar which tells you about the local curvature of the universe and is a contraction of the Ricci tensor with the inverse metric, i.e.
\begin{equation}
	R = R_{\mu\nu}g^{\mu\nu} \label{eq:ricscalar}
\end{equation}
which is related to the curvature of the universe by
\begin{align}
R > 0 &\iff \text{Positive curvature}\\
R = 0 &\impliedby \text{Flat space}\\
R < 0 &\iff \text{Negative curvature}
\end{align}
where it is important to note that not all $R=0$ values are flat space, but all flat space has $R=0$.
% subsubsection curvature (end)
\subsection{The Einstein Field Equations} % (fold)
\label{sub:The_einstein_field_equations}
%Talk about the importance of the equation and why this is actually fully derived
%
%
%
\subsubsection{The Einstein-Hilbert Action} % (fold)
\label{subsub:the_einstein-hilbert_action}
The Einstein Field Equations (EFEs) are a set of coupled, non-linear, partial differential equations that can provide us with a link between the geometry of spacetime, and what fills the universe (energy, matter, etc). To derive the EFEs we must start with the Einstein-Hilbert action
\begin{equation}
	\label{eq:HE_action}
	S = \int \sqrt{-g} \left(\frac{R}{16\pi G} + \Lagr_F\right)d^Dx
\end{equation}
where $g$ is the determinant of the metric, $R$ is the Ricci scalar, $G$ is Newton's gravitational constant, and $\Lagr_F$ is a general Lagrangian term that describes a field, e.g. electromagnetism or matter, and can be set to 0 if you are considering the vacuum.\\
For now let us keep $\Lagr_F$ general and we can specialise to a particular case later. We shall begin as usual by varying the action w.r.t $g^{\mu\nu}$
\begin{equation}
	\frac{\delta S}{\delta g^{\mu\nu}} = \int \left(\frac{1}{16\pi G} \frac{\delta(\sqrt{-g}R)}{\delta g^{\mu\nu}} + \frac{\delta(\sqrt{-g}\Lagr_F)}{\delta g^{\mu\nu}}\right)d^Dx.
\end{equation}
Now by expanding out the first term and pulling out some terms we can write this in the form
\begin{equation}
	{\delta S} = \int \left(\frac{1}{16\pi G} \left(\frac{\delta(R)}{\delta g^{\mu\nu}} + \frac{R}{\sqrt{-g}}\frac{\delta(\sqrt{-g})}{\delta g^{\mu\nu}}\right) + \frac{1}{\sqrt{-g}}\frac{\delta(\sqrt{-g}\Lagr_F)}{\delta g^{\mu\nu}}\right)\sqrt{-g}\;\delta g^{\mu\nu} d^Dx = 0
\end{equation}
using the principle of least action to equate this to zero. Then finally we know this must be true for all $\delta g^{\mu\nu}$ and as such can finally conclude that
\begin{equation}
	\frac{\delta(R)}{\delta g^{\mu\nu}} + \frac{R}{\sqrt{-g}}\frac{\delta(\sqrt{-g})}{\delta g^{\mu\nu}} = -\frac{16\pi G}{\sqrt{-g}}\frac{\delta(\sqrt{-g}\Lagr_F)}{\delta g^{\mu\nu}}. \label{eq:EFE_unfinished}
\end{equation}
% subsubsection the_einstein-hilbert_action (end)
\subsubsection{Variation of the Ricci Scalar} % (fold)
\label{subsub:variation_of_the_ricci_scalar}
Let's now focus the LHS and calculate each of the two terms individually, starting with the variation of the Ricci Scalar. Before we can work on the Ricci scalar, we need to work out how it depends on the metric, which means we need to return to the higher rank tensor, the Riemann tensor, and more precisely the equation defining it in terms of Christoffel symbols as we gave in (\ref{eq:riemann}), once we have this we can vary the Riemann tensor. Remembering that we can commute the variation and the partial derivative, which gives us
\begin{equation}
	\delta {R^\rho}_{\mu\nu\sigma} = \partial_\nu \delta\Gamma^{\rho}_{\mu\sigma} - \partial_\sigma \delta\Gamma^{\rho}_{\mu\nu} + \delta\Gamma^{\rho}_{\nu \lambda}\Gamma^{\lambda}_{\mu\sigma} + \Gamma^{\rho}_{\nu \lambda}\delta\Gamma^{\lambda}_{\mu\sigma} - \delta\Gamma^{\rho}_{\sigma \lambda}\Gamma^{\lambda}_{\mu \nu} - \Gamma^{\rho}_{\sigma \lambda}\delta\Gamma^{\lambda}_{\mu \nu}.
\end{equation}
It's important to understand that $\delta\Gamma$ is just the difference of two connections and is fortunately a tensor, as such we can take the covariant derivative of it
\begin{equation}
	D_\lambda(\delta \Gamma^\rho_{\nu\mu}) = \partial_\lambda(\delta \Gamma^{\rho}_{\nu\mu}) +\Gamma^{\rho}_{\lambda\sigma}\delta \Gamma^{\sigma}_{\nu\mu} - \Gamma^{\sigma}_{\lambda\mu}\delta \Gamma^{\rho}_{\sigma\nu} - \Gamma^{\sigma}_{\lambda\nu}\delta \Gamma^{\rho}_{\sigma\mu}
\end{equation}
which looks similar to what we had before, but we don't quite have the terms we want. Next we take the difference of two of these terms and, as we will see, this will give us the equation we are looking for;
\begin{multline}
 	D_\nu(\delta \Gamma^\rho_{\mu\sigma}) - D_\sigma(\delta \Gamma^\rho_{\mu\nu}) = \left[\partial_\nu(\delta \Gamma^{\rho}_{\mu\sigma}) + \Gamma^{\rho}_{\nu\lambda}\delta \Gamma^{\lambda}_{\mu\sigma} - \Gamma^{\lambda}_{\nu\sigma}\delta \Gamma^{\rho}_{\lambda\mu} - \Gamma^{\lambda}_{\nu\mu}\delta \Gamma^{\rho}_{\lambda\sigma}\right] \\- \left[\partial_\sigma(\delta \Gamma^{\rho}_{\mu\nu}) + \Gamma^{\rho}_{\sigma\lambda}\delta \Gamma^{\lambda}_{\mu\nu} - \Gamma^{\lambda}_{\sigma\mu}\delta \Gamma^{\rho}_{\lambda\nu} - \Gamma^{\lambda}_{\sigma\nu}\delta \Gamma^{\rho}_{\lambda\mu}\right] = \delta {R^\rho}_{\mu\nu\sigma}. \label{eq:delta_riemann}
\end{multline} 
The next step is comparably much easier as we only have to contract the upper and lower-middle index, that is
\begin{align}
	R_{\mu\nu} &= {R^\rho}_{\mu\rho\nu}\\
	\delta R_{\mu\nu} &= D_\rho(\delta \Gamma^\rho_{\mu\nu}) - D_\nu(\delta \Gamma^\rho_{\mu\rho}),
\end{align}
where we have relabelled the indices from (\ref{eq:delta_riemann}) such that $\nu\leftrightarrow\sigma$ then $\sigma\rightarrow\rho$ so that $\rho$ is summed over. Then finally we have to compute the Ricci scalar,
\begin{align}
R &= g^{\mu\nu} R_{\mu\nu}\\
\implies \delta R &= \delta g^{\mu\nu}R_{\mu\nu} + g^{\mu\nu} \delta R_{\mu\nu}\\
&= R_{\mu\nu}\delta g^{\mu\nu} + g^{\mu\nu}(D_\rho(\delta \Gamma^{\rho}_{\nu\mu}) - D_\nu(\delta \Gamma^{\rho}_{\rho\mu}))
\end{align}
but by using the fact that the covariant derivative of the metric is zero and by relabelling dummy indices this can be written as
\begin{equation}
	\delta R = R_{\mu\nu}\delta g^{\mu\nu} + D_\sigma(g^{\mu\nu}\delta \Gamma^{\sigma}_{\mu\nu} - g^{\mu\sigma}\delta \Gamma^{\rho}_{\mu\rho}).
\end{equation}
The second term in this equation is now clearly a total derivative, and as such does not contribute to the action and can be ignored giving us that $\delta R = R_{\mu\nu}\delta g^{\mu\nu}$ and that the term from (\ref{eq:EFE_unfinished}) gives us
\begin{equation}
	\frac{\delta R}{\delta g^{\mu\nu}} = R_{\mu\nu}.
\end{equation}
% subsubsectionvariation_of_the_ricci_scalar (end)
\subsubsection{Variation of the Determinant of the Metric} % (fold)
\label{subsub:variation_of_the_determinant_of_the_metric}
Now we want to focus on the second term that contains the variation of $\sqrt{-\det(g_{\mu\nu})}$. We start simply by taking the chain rule to simplify the problem and get
\begin{equation}
	\delta(\sqrt{-g}) = \frac{-1}{2\sqrt{-g}}\delta(g). \label{eq:metric_var_unfinished}
\end{equation}
We will now use a trick to simplify the calculation and, without loss of generality, change to a coordinate system where the metric is diagonal meaning that 
\begin{align}
	g &= \prod_{\mu=0}^{D-1} g_{\mu\mu}\\
	\implies \delta g & = \delta(\prod_{\mu=0}^{D-1} g_{\mu\mu}).
\end{align}
Once expanded this gives us
\begin{align}
	\delta g & = \delta(g_{00})\prod_{\mu \neq 0 }g_{\mu\mu} + \delta(g_{11})\prod_{\mu\neq 1} g_{\mu\mu} \dots + \delta(g_{(D-1)(D-1)})\prod_{\mu\neq D-1} g_{\mu\mu}\\
	&= (\prod_{\mu=0}^{D-1} g_{\mu\mu})\sum_{\mu=0}^{D-1} \frac{\delta g_{\mu\mu}}{g_{\mu\mu}} = g g^{\mu\mu} \delta g_{\mu\mu}
\end{align}
where we have explicitly placed in the summation to make the factorisation easier to follow, and then used the fact that the metric is diagonal to write $1$ divided by the metric as the inverse metric. We can now return to a non-diagonal metric having done the otherwise difficult part of the calculation and using the fact that $g_{\mu\nu}\delta g^{\mu\nu} = - g^{\mu\nu} \delta g_{\mu\nu}$ we can write (\ref{eq:metric_var_unfinished}) as
\begin{align}
	\delta(\sqrt{-g}) = \frac{-\sqrt{-g}}{2}g_{\mu\nu}\delta g^{\mu\nu}.
\end{align}
Then bringing together all of this work we can write Einstein's Field Equations as
\begin{equation}
	R_{\mu\nu} - \frac{1}{2}g_{\mu\nu}R = 8\pi G\; T_{\mu\nu} \label{eq:efe}
\end{equation}
where the RHS can be written as $G_{\mu\nu}\equiv R_{\mu\nu} - \frac{1}{2}g_{\mu\nu}R$ and
\begin{equation}
	T_{\mu\nu}=\frac{-2}{\sqrt{-g}}\frac{\delta(\sqrt{-g}\Lagr_F)}{\delta g^{\mu\nu}}
\end{equation} 
is the \emph{stress-energy tensor} (SE tensor) which describes the physical properties (e.g. energy density, stress, etc) of the space, and in the vacuum $T_{\mu\nu} = 0$. This is the arguably the most important equation in GR, the RHS describes purely the geometry of the spacetime as where the LHS describes only the physical properties of the spacetime which might otherwise be impossible to relate.
Another form of the equation includes a cosmological constant, $\Lambda$ and is written as
\begin{equation}
	R_{\mu\nu} - \frac{1}{2}g_{\mu\nu}R +g_{\mu\nu}\Lambda= 8\pi G\; T_{\mu\nu}
\end{equation}
but for the entirety of this report we will always consider the cosmological constant to be equal to 0.
% subsubsection variation_of_the_determinant_of_the_metric (end)
% subsection deriving_the_einstein_field_equations (end)
\subsubsection{Testing the EFE for the 4 Dimensional EM Case} % (fold)
\label{sub:from_the_vacumm_to_electromagnetism}
Let us step away from the general expression for the SE tensor and specialise to the case where $\Lagr_D = \Lagr_{E.M}$ so that we can add electromagnetism to our spacetime. We will start by deriving the general form of $T_{\mu\nu}$ before checking our result by comparing with known results in 4 dimensions. First we need to write the Lagrangian density of electromagnetism as
\begin{equation}
	\Lagr_{E.M} = \frac{-1}{4} F_{\mu\nu} F_{\rho\sigma} g^{\mu\rho} g^{\nu\sigma}
\end{equation}
where $F_{\mu\nu} = \partial_\mu A_\nu - \partial_\nu A_\mu$ and $A$ is the electromagnetic 4-potential. By considering flat spacetime, then by considering both the form of this and the solution to the Euler-Lagrange equations it is trivial to show that you can derive all of Maxwell's equations in covariant form
\begin{align}
	D_{\mu}F^{\mu\nu}&=J^{\nu}\\
	D_{\mu}\left(\frac{1}{2}\epsilon^{\mu\nu\rho\lambda}F_{\rho\lambda}\right)&=0
\end{align}
where $\epsilon^{\mu\nu\rho\lambda}$ is the Levi-Civita symbol and $J=\left(\rho,\vec{J}\right)$ is the covariant four-vector that contains the charge density, $\rho$, and the current density $\vec{J}$.
Let us start our derivation of the SE tensor by splitting the variation into two terms
\begin{align}
	T_{\mu\nu}&=\frac{-2}{\sqrt{-g}}\frac{\delta(\sqrt{-g}\Lagr_{E.M})}{\delta g^{\mu\nu}}\\
	&= \frac{-2}{\sqrt{-g}}\left(\frac{\delta(\sqrt{-g})\Lagr_{E.M}}{\delta g^{\mu\nu}} + \frac{\sqrt{-g}(\delta \Lagr_{E.M})}{\delta g^{\mu\nu}} \right);
	\label{eq:SE_EM_unf}
\end{align}
fortunately for us we have already calculated the first term and so can already simplify this to
\begin{equation}
	T_{\mu\nu}= g_{\mu\nu}\Lagr_{E.M} - 2 \frac{\delta \Lagr_{E.M}}{\delta g^{\mu\nu}}. \label{eq:energy_mom_tensor_w_lagr}
\end{equation}
It is important to remember that this variation is with respect to the metric and that $F_{\mu\nu}$ does not depend on the metric and as such can be considered constant. As well as this we must be careful to change the indices on the terms in the Lagrangian as these are free indices, whereas the ones on the metric variation are fixed. With this all in mind we can write
\begin{equation}
	\delta \Lagr_{E.M} = \frac{-1}{4}F_{\alpha\beta}F_{\rho\sigma}( (\delta g^{\alpha\rho}) g^{\beta\sigma} + g^{\alpha\rho} (\delta g^{\beta\sigma})).
\end{equation}
Then we simply have that
\begin{align}
	\frac{\delta\Lagr_{E.M}}{\delta g^{\mu\nu}} &= \frac{-1}{4}F_{\alpha\beta}F_{\rho\sigma}({\delta^\alpha}_\mu {\delta^\rho}_\nu g^{\beta\sigma} + g^{\alpha\rho} {\delta^\beta}_\mu {\delta^\sigma}_\mu)\\
	&=\frac{-1}{4} (F_{\mu\beta}F_{\nu\sigma}g^{\beta\sigma} + F_{\alpha\nu}F_{\rho\mu}g^{\alpha\rho})
\end{align}
where ${\delta^\alpha}_\mu$ is the Kronecker delta function. Which when inserted into (\ref{eq:SE_EM_unf}) gives
\begin{align}
	T_{\mu\nu}&=\frac{-1}{4}g_{\mu\nu}F_{\alpha\beta}F_{\rho\sigma}g^{\alpha\rho}g^{\beta\sigma} + \frac{1}{2} g^{\beta\sigma}F_{\mu\beta}F_{\nu\sigma} + \frac{1}{2} g^{\alpha\rho}F_{\alpha\mu}F_{\rho\nu}\\
	&=\frac{-1}{4}g_{\mu\nu}F_{\alpha\beta}F^{\alpha\beta}+  {F_\mu}^\sigma F_{\nu\sigma} \label{eq:SE_EM_fin}
\end{align}
where we have relabelled some dummy indices and used the anti-symmetric property of $F$ (i.e $F_{\mu\nu} = -F_{\nu\mu}$). Finally if we wish to check the validity of our results we can specialise further and consider the 4 dimensional flat-space case where the $(0,0)$ component of the SE tensor should be the energy density of the E.M field. The E.M tensor in 4 dimensions is written as
\begin{equation}
	F_{\mu\nu} = 
		\begin{pmatrix}
		0 & -E_x & -E_y & -E_z \\
		E_x & 0 & B_z & -B_y \\
		E_y & -B_z & 0 & B_x \\
		E_z & B_y & -B_x & 0 
	\end{pmatrix}
\end{equation}
and from this it follows that $F_{\mu\nu}F^{\mu\nu} = 2(B^2-E^2)$. So we finally have that
\begin{align}
	T^{00}&=\frac{-1}{4}\eta^{00}F_{\alpha\beta}F^{\alpha\beta}+  {F^0}_\sigma F^{0\sigma}\\
	&=\frac{-1}{4}(-1)2(B^2-E^2) + E^2\\
	&= \frac{E^2+B^2}{2}
\end{align}
which is the energy density we expected.
% subsection from__the_vacumm_to_electromagnetism (end)
%
%
\section{Black Holes} % (fold)
\label{sec:black_holes}
\subsection{History} % (fold)
\label{sub:history}
Isaac Newton first published his theory of gravity in 1687 in his Philosophi{\ae} Naturalis Principia Mathematica but it wasn't until almost a century later in 1783 that Rev. John Mitchell, a scientist in the fields of geology and physics, first came up with the idea that an object might be so massive that the escape velocity would be greater than the speed of light and wrote as much to Henry Cavendish;
\begin{quote}
\begin{textit}{If the semi-diameter of a sphere of the same density as the Sun were to exceed that of the Sun in the proportion of 500 to 1, a body falling from an infinite height towards it would have acquired at its surface greater velocity than that of light, and consequently supposing light to be attracted by the same force in proportion to its vis inertiae, with other bodies, all light emitted from such a body would be made to return towards it by its own proper gravity.} \\
\end{textit}
-J. Michell, ``A letter to Henry Cavendish'' \cite{michell_leter}.
\end{quote}
Laplace later wrote about the idea of these \emph{Dark Stars} in the first two editions of his book \emph{Exposition du syst{\'e}me du Monde} but removed them from later editions and never seemed to mention the idea again\cite{gillispie2000pierre}. It was largely due to there being no understanding of how a massless wave could be influenced by gravity that the idea pretty much disappeared until the 20\textsuperscript{th} century when the idea resurfaced due to Einstein's Theory of General Relativity in 1915. It was only a few months later that Karl Schwarzschild found a solution to EFE for a spherical non-rotating mass. Work continued throughout the first half of the century in developing the properties of this object, showing the singularity at the Schwarzschild radius was simply a coordinate singularity as well as providing some astronomical limits on star masses that would provide unstable systems.
The 1960's and onwards were a golden age for GR with discoveries of solutions for both rotating and charged black holes, the discovery of neutron stars, the development of the \emph{No Hair Theorem}, as well as it being shown that singularities appear in generic solutions.
The 1970's gave us the thermodynamics of black holes (Sec \ref{sec:blackhole_thermodynamics}) and since then there has been steady work on studying black holes, both observational and theoretical, and the field is still active today.
% subsection history (end)
\subsection{Causality of Black Holes} % (fold)
\label{sub:understanding_black_holes}
We will see that the metrics for black holes will end up having a form of 
\begin{equation}
	ds^2= -f(r)dt^2 + f^{-1}(r)dr^2+\ldots
\end{equation}
which for a certain value of $r=r_{0}$, the horizon, has that $f(r_{0})=0$. What this means is that for an observer at infinity time will seem to stop at this value of r; if the observer were to watch an object fall towards the black hole it would seem to take an infinite amount of time for object to reach the horizon. As for the object itself, for most cases it would not feel a change in passing through the horizon as the curvature of space at this point is not likely to be much more than on earth, however what will change is the light cones. As the object falls towards the horizon, the lines of the light cones which in flat space have a 45\textsuperscript{} angle between the lines and either axis, start to move towards each other as $f(r)\to 0$. At the horizon the lines of the light cone aren't a cone at all, but a single line from past to future; once inside the horizon the light cone \emph{spreads out} once again but with time and space swapped (due to the sign of $f(r)$ flipping) which means that the future of such an object is to move towards $r=0$ just as for any object outside the horizon the future is to move towards greater $t$, see figure (\ref{fig:light_cones}) for a pictorial representation.
\begin{figure}
        \includegraphics[scale=0.4]{light_cones.pdf}
	\caption{Light cones of the spacetime up to and beyond the horizon }
	\label{fig:light_cones}
\end{figure}
%
%
% subsection understanding_black_holes (end)
\subsection{General Spherically Symmetric Solutions in 4 Dimensions} % (fold)
\label{sub:general_spherically_symmetric_solutions}
We would now like to begin finding solutions to (\ref{eq:efe}) and more specifically those solutions which give rise to black holes. To start with we will consider a spherically symmetric static solution. Whilst it is possible to write a more general spherically symmetric line element we will choice to use one where the angular function has been fixed as $r^{2}$ which leaves us with
\begin{equation}
	ds^2= -f(r)dt^2 + g(r)dr^2+r^2 d{\Omega}^2_{D-2}
\end{equation}
where $d{\Omega}^2_{D-2}$ is the angular contribution to the metric in $D$ dimensions; in 4 dimensions this is $d\theta^2+sin^2(\theta)d\phi^2$. With some knowledge of the solutions we will arrive at, as well as following the work by Ahmad et al\cite{Ahmad:2013tia} it is more convenient to write our line element in the form
\begin{equation}
	ds^2= -e^{2\Phi(r)}dt^2 + e^{2\Lambda(r)}dr^2 + r^2 d{\Omega}^2_{2} \label{eq:sch-met-func-4d}.
\end{equation}
Now the non-zero Christoffel symbols for this metric are
\renewcommand{\arraystretch}{2}
\begin{equation}
	\begin{array}{llll}
		\Gamma^{t}_{rt}=\Phi^\prime & & & \\
		\Gamma^{r}_{rr}=\Lambda^\prime &\Gamma^{r}_{tt}=\Phi^\prime e^{2(\Phi-\Lambda)} &\Gamma^{r}_{\theta\theta}=-re^{-2\Lambda} &\Gamma^{r}_{\phi\phi}=-r\sin^2(\theta) e^{-2\Lambda} \\
		\Gamma^{\theta}_{r\theta}=\frac{1}{r} 	&\Gamma^{\theta}_{\phi\phi}=-\sin(\theta)\cos(\theta) & &\\
		\Gamma^{\phi}_{r\phi}=\frac{1}{r} &\Gamma^{\phi}_{\theta\phi}=\cot(\theta) & &
	\end{array}
\end{equation}
where the prime denotes a derivative with respect to $r$. With these we can calculate the six unique non-zero Riemann Tensors which are\renewcommand{\arraystretch}{1}
\begin{align}
&{R^t}_{\theta\theta t}=\Phi^\prime r e^{-2\Lambda}\nonumber \\ 
&{R^r}_{\theta\theta r}= -r \Lambda^\prime e^{-2\Lambda}\nonumber\\ 
&{R^r}_{\phi\phi r}= -\Lambda^\prime r \sin^2(\theta) e^{-2\Lambda}\nonumber\\
&{R^r}_{ttr}= -(\Phi^{\prime\prime}+\Phi^{\prime 2} -\Lambda^\prime\Phi^\prime)e^{2(\Phi-\Lambda)}\nonumber\\
&{R^\phi}_{tt\phi}= \frac{-1}{r}\Phi^\prime e^{2(\Phi-\Lambda)}\nonumber\\
&{R^\phi}_{\theta\theta\phi}= e^{-2 \Lambda} -1.
\end{align}
The next step is to calculate the Ricci Tensor using (\ref{eq:ricci}) and remembering the symmetries of the Riemann tensor from (\ref{eq:riemann-prop-1}-\ref{eq:riemann-prop-3}) we find that the four non-zero Ricci tensors are the symmetric ones, i.e.
\begin{align}
R_{tt}&=\left(\Phi^{\prime\prime}+\Phi^{\prime2}-\Phi^\prime\Lambda^\prime+\frac{2}{r}\Phi^\prime\right)e^{2(\Phi-\Lambda)} \label{eq:riccibh1}	\\
R_{rr}&=-\Phi^{\prime\prime}-\Phi^{\prime2}+\Phi^\prime\Lambda^\prime+\frac{2}{r}\Lambda^\prime\label{eq:riccibh2}\\
R_{\theta\theta}&=(-r\Phi^\prime+r\Lambda^\prime -1)e^{-2\Lambda} +1	\label{eq:riccibh3}\\
R_{\phi\phi}&=[(-r\Phi^\prime+r\Lambda^\prime -1)e^{-2\Lambda} +1]\sin^2(\theta) = R_{\theta\theta} \sin^2(\theta).\label{eq:riccibh4}
\end{align}
For completeness we will calculate the Ricci Scalar using (\ref{eq:ricscalar})
\begin{equation}
	R=e^{-2\Lambda}\left(-2\Phi^{\prime\prime}-2\Phi^{\prime2}+2\Phi^\prime\Lambda^\prime- \frac{4}{r}\Phi^\prime +\frac{4}{r}\Lambda^\prime -\frac{2}{r^2}\right) +\frac{2}{r^2}
\end{equation}
and also all non-zero components of the Einstein tensor
\begin{align}
G_{tt} &= e^{2(\Phi-\Lambda)}\left(\frac{2}{r}\Lambda^\prime - \frac{1}{r^2}\right) +\frac{1}{r^2}e^{2\Phi}	\\
G_{rr} &= \frac{2}{r}\Phi^\prime + \frac{1}{r^2} - \frac{1}{r^2}e^{2\Lambda}	\\
G_{\theta\theta} &= e^{-2\Lambda}r^2\left(\Phi^{\prime\prime}+\Phi^{\prime2}-\Lambda^\prime\Phi^\prime +\frac{\Phi^\prime}{r}- \frac{\Lambda^\prime}{r}\right)\\
G_{\phi\phi} &= \sin^2(\theta)G_{\theta\theta}.
\end{align}
% subsection general_spherically_symmetric_solutions (end)
\subsection{Schwarzschild Solution in 4 Dimensions} % (fold)
\label{sub:schwarzschild_solution}
The Schwarzschild solution, first derived by Karl Schwarzschild in 1915 (published in 1916), is the most basic case we can consider; that of a non-charged, non-rotating black hole. To find this solution let us first consider the trace of the Einstein Tensor in 4 dimensions
\begin{equation}
	{G^\mu}_\mu = {R^\mu}_\mu- \frac{1}{2}R {\delta^\mu}_\mu = R- 2R =-R
\end{equation}
which allows us to write that 
\begin{equation}
	R_{\mu\nu}=G_{\mu\nu}+\frac{1}{2}G g_{\mu\nu}.
\end{equation}
We want to solve the Einstein equations for the vacuum solution, i.e $G_{\mu\nu}=0$ which we have just shown is equivalent to $R_{\mu\nu}=0$. We can trivially solve (\ref{eq:riccibh4}) when $\theta=n\pi$ and we know that $e^{2(\Phi-\Lambda)}\ne0$ so we can use this to simplify (\ref{eq:riccibh1}). Adding the reduced version of this to (\ref{eq:riccibh2}) gives us that
\begin{align}
	\frac{2}{r}(\Lambda^\prime+\Phi^\prime)=0 &\implies \Lambda^\prime = - \Phi^\prime\\
	\implies \Phi + \Lambda &= C
\end{align}
where $C$ is a constant. This means we can now write the line element as 
\begin{equation}
	ds^2 =-e^{-2\Lambda}c^{2C}dt^2 + e^{2\Lambda}dr^2 + r^2 d{\Omega}^2_{2}
\end{equation}
but by going to a coordinate system where we take $t\to e^Ct$ then we can effectively have that $C=0$. By now using this information in (\ref{eq:riccibh3}) we can arrive at a single first-order linear differential equation;
\begin{equation}
	2r\Lambda^\prime e^{-2\Lambda} - e^{-2\Lambda} +1 = 0
\end{equation}
which we can rewrite in the form
\begin{equation}
	\frac{d}{dr}(-re^{-2\Lambda}) =-1
\end{equation}
which easily provides us with the solution that
\begin{equation}
	re^{-2\Lambda}=re^{2\Phi}= r \pm a.
\end{equation}
It is worth at this point checking that this solution makes sense before calculating a value for $a$; to do so we will consider the solution as $r\to \infty$ as this should look like flat space. 
\begin{align}
\lim_{r\to\infty}1\pm \frac{a}{r} =1 \implies e^{-2\Lambda}=e^{2\Lambda}=e^{2\Phi} =1
\end{align}
which gives us back our flat metric as we expected. We can calculate the value of $a$ by comparing the action of a particle $\int m \;ds$ with the usual form of an action $\int (T-V)dt$ with this potential being the Newtonian gravitational potential 
\begin{equation}
	V=-\frac{GmM}{r}.
\end{equation} Doing so gives us that $a=\frac{2GM}{c^2}=r_s$, where $G$ is Newtons gravitational constant, $M$ is the mass of the black hole, $r_s$ is the Schwarzschild radius and $c$ is the speed of light (although for us of course $c=1$). The final form of the Schwarzschild line element is then
\begin{equation}
	ds^2 = -\left(1- \frac{2GM}{r}\right)dt^2 +\left(1- \frac{2GM}{r}\right)^{-1}dr^2 + r^2d\Omega^2_{2}.
\end{equation}
% subsection schwarzschild_solution (end)
\subsection{Reissner-Nordstr\"om Solution in 4 Dimensions} % (fold)
\label{sub:reissner_nordstr"om_solution}
The next solution we will look at is the case of a charged, non-rotating, black hole meaning that we are no longer solving for a vacuum case of the Einstein equations. Before we calculate our SE tensor let us re-write here, for convenience, the line element we are using as well as (\ref{eq:SE_EM_fin}) for the SE tensor in electromagnetism
\begin{align}
	ds^2 &= -e^{2\Phi}dt^2 + e^{2\Lambda}dr^2 + r^2 d{\Omega}^2_{2};\\
	T_{\mu\nu}&=\frac{-1}{4}g_{\mu\nu}F_{\alpha\beta}F^{\alpha\beta}+  {F_\mu}^\sigma F_{\nu\sigma}.
\end{align}
Now we begin the work of calculating the SE tensor components for the electric case as the Reissner-Nordstr\"om (RN) solution has a zero magnetic field. We remind ourselves that electromagnetic tensor, $F_{\mu\nu}$, can be written as 
\begin{equation}
	F_{\mu\nu} = \partial_{\mu}A_{\nu} - \partial_{\nu}A_{\mu}
\end{equation}
where $A$ is the 4-potential of the electromagnetic field with the first component being related to the electric field and the remaining components related to the magnetic field; as such the last 3 are equal to zero in our case. For now we shall consider a generic radial electric potential and later specialise to the Coulomb potential as this will make it easier to consider the higher dimensional case when we come to that. Using all of this information gives us that
\begin{align}
	A_{0}&=e^{2\Psi(r)}
	&F_{\mu\nu}(r)=
	\begin{cases}
		2\Psi ^{\prime} e^{2\Psi}, &\text{if } \mu\nu=tr\\
		-2\Psi ^{\prime} e^{2\Psi}, &\text{if } \mu\nu=rt\\
		0, 	&\text{otherwise}
	\end{cases}
\end{align}
where the electromagnetic tensor is obviously still traceless and anti-symmetric. We will also need the version where both indices are raised and using our general metric this gives us that
\begin{equation}
	F^{\mu\nu}(r)=
	\begin{cases}
		-2\Psi ^{\prime} e^{2\Psi-2\Phi-2\Lambda}, &\text{if } \mu\nu=tr\\
		2\Psi ^{\prime} e^{2\Psi-2\Phi-2\Lambda}, &\text{if } \mu\nu=rt\\
		0, 	&\text{otherwise}.
	\end{cases}
\end{equation}
Looking at our equation for the stress energy tensor we see that our next step is to calculate the inner product of $F$
\begin{align}
	F_{\mu\nu}F^{\mu\nu}&=F_{rt}F^{rt}+F_{tr}F^{tr}=2F_{rt}F^{rt}\\
	&=-8\Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda}.
\end{align}
We also need the mixed product
\begin{align}
	{F_{\mu}}^{\sigma} F_{\nu\sigma}&= F^{\gamma\sigma}F_{\nu\sigma}g_{\mu\gamma}\\
	&=g_{\mu\gamma}\left(F^{\gamma t}F_{\nu t}+ F^{\gamma r}F_{\nu r}\right)
\end{align}
where we have used the fact that the tensor is only non-zero for two values to simplify the expression. The next step is to calculate the full SE tensor which now most generally can be written as
\begin{equation}
	T_{\mu\nu}= 2g_{\mu\nu}\Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda} +g_{\mu\gamma}\left(F^{\gamma t}F_{\nu t}+ F^{\gamma r}F_{\nu r}\right)
\end{equation}
which we can see will be a diagonal matrix with elements
\begin{align}
	T_{tt}&=-2\Psi ^{\prime 2} e^{4\Psi-2\Lambda} +4\Psi ^{\prime 2} e^{4\Psi-2\Lambda}\\
	T_{rr}&=2 \Psi ^{\prime 2} e^{4\Psi-2\Phi} -4 \Psi ^{\prime 2} e^{4\Psi-2\Phi}\\
	T_{\theta\theta}&=2r^{2} \Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda}\\
	T_{\phi\phi}&= \sin^2(\theta) \;T_{\theta\theta}
\end{align}
which can be written in the matrix form as
\begin{equation}
	T_{\mu\nu} = 
		2 \Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda} \begin{pmatrix}
		e^{2\Phi} & 0 & 0 & 0 \\
		0 & -e^{2\Lambda} & 0 & 0 \\
		0 & 0 & r^{2} & 0 \\
		0 & 0 & 0 & r^{2}\sin^{2}(\theta)
	\end{pmatrix}.
\end{equation}
Next we use all the quantities we have calculated in the EFEs to solve for the unknown functions. As before the two angular equations are linearly dependent and as such we only have 3 equations which with some minor rearrangement are
\begin{align}
	e^{-2\Lambda}\left(\frac{2}{r}\Lambda^\prime - \frac{1}{r^2}\right) +\frac{1}{r^2}&= 16\pi G \Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda}\label{eq:charged_efe_1}\\
	e^{-2\Lambda}\left(\frac{2}{r}\Phi^\prime + \frac{1}{r^2}\right) - \frac{1}{r^2} &=-16\pi G \Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda}\\
	e^{-2\Lambda}\left(\Phi^{\prime\prime}+\Phi^{\prime2}-\Lambda^\prime\Phi^\prime +\frac{\Phi^\prime}{r}- \frac{\Lambda^\prime}{r}\right)&=16\pi G \Psi ^{\prime 2} e^{4\Psi-2\Phi-2\Lambda}.
\end{align}
From the first two equations it falls out that once again $\Lambda ^{\prime} =-\Phi ^{\prime}$ giving us as with the Schwarzschild case that $e^{2\Lambda}=e^{-2\Phi}$. Next we take the third equation, simplifying it using the relation between $\Lambda$ and $\Phi$, and equate it to the first to give us that
\begin{equation}
	e^{2\Phi}\left(\Phi ^{\prime\prime} + 2\Phi ^{\prime 2} +\frac{4\Phi ^{\prime} }{r} +\frac{1}{r^{2}}\right)= \frac{1}{r^{2}}.
\end{equation}
Much like before we can write this in terms of differentials of our original function
\begin{equation}
	\frac{1}{2}\frac{d^{2}}{dr^{2}} e^{2\Phi} +\frac{2}{r}\left(\frac{d}{dr}e^{2\Phi}\right) + \frac{e^{2\Phi}}{r^{2}}=\frac{1}{r^{2}} \label{eq:charged_diff_eq_to_solve}
\end{equation}
which is solvable if we try the ansatz 
\begin{equation}
	e^{2\Phi}=A+\frac{B}{r}+\frac{C}{r^{2}}.
\end{equation}
By plugging this ansatz into (\ref{eq:charged_diff_eq_to_solve}) we get that $A=1$, we know that $B=-2GM$ from the homogeneous case which leaves us with just $C$ to find. By using this ansatz in (\ref{eq:charged_efe_1}) we discover that
\begin{align}
	\frac{C}{r^{4}}&=16\pi G \Psi ^{\prime 2} e^{4\Psi}\\
	&= 4\pi G \left(\frac{d}{dr}e^{2\Psi}\right)^{2}.
\end{align}
Now is when we will choice a particular electric potential; by using the Coulomb potential we can set 
\begin{equation}
	A_{\mu}=\left(\frac{q}{4\pi r},0,0,0\right)
\end{equation}
and this gives us that $C=\frac{Gq^{2}}{4\pi}\equiv GQ^{2}$ giving us that our final expression for the Reissner-Nordstr\"om line element is
\begin{equation}
	ds^2 = -\left(1- \frac{r_s}{r} +\frac{r_Q^2}{r^2}\right)dt^2 +\left(1- \frac{r_s}{r} +\frac{r_Q^2}{r^2}\right)^{-1}dr^2 + r^2d\Omega^2_{(2)}
\end{equation}
where $r_s$ is the Schwarzschild radius as before and $r_Q^2=GQ^{2}$.
% subsection reissner_nordstr"om_solution (end)
\subsection{Spherical Solutions in Extra Dimensions} % (fold)
\label{sub:spherical_solution_in_extra_dimensions}
\subsubsection{General Spherical Solution} % (fold)
\label{sub:general_spherical_solution}
It would be useful for us to have an expression for the Schwarzschild and RN metrics and radii in higher dimensions and so we shall derive here the general form of these for $D\ge 4$. To do this we will be working through the derivation using differential forms and operators, including wedge products and connections. Whilst a comprehensive understanding of the topic is not required to follow the derivation as I shall provide all necessary properties and equations as we reach them, I will not be explaining these in any detail and will simplify definitions where helpful so should you wish to have a more rigorous introduction to the topic I suggest \emph{Differential Geometry, Gauge Theories, and Gravity} by M. G\"ockeler and T. Sch\"ucker\cite{ differential_geometry}, specifically chapters (1-5).

To begin with we assume a metric of a very similar form as (\ref{eq:sch-met-func-4d}) with the only change being the dimension of the last term
\begin{equation}
	ds^2= -e^{2\Phi(r)}dt^2 + e^{2\Lambda(r)}dr^2 + r^2 d{\Omega}^2_{d-1}
\end{equation}
where now $d{\Omega}^2_{d-1}$ is the squared solid angle element for $d\ge 3$ dimensions and $d=D-1$ where $D$ is as always the total dimension of our spacetime. More explicitly we can write
\begin{equation}
	d{\Omega}^2_{d-1} = d\xi^2_{d-1} + \sum_{i=1}^{d-2}\prod_{j=i+1}^{d-1}\sin^2(\xi_j)d\xi^2_i
\end{equation}
where $\xi_i$ are the angles of our hyper-spherical coordinates with ranges $0\le \xi_1 <2\pi$ and $0\le \xi_k \le \pi$ where $k=2,\dots,d-1$. It is worth pointing out here that we have reversed the order of the angular components in our metric, something which has no impact on any calculations as the order of coordinates is completely arbitrary, but something that makes writing the angular elements in a concise form much easier to do. With this our metric now looks like
\begin{equation}
		g_{\mu\nu}= 
		\begin{pmatrix}
		-e^{2\Phi(r)} & 0 & 0 & 0\dots & 0 & 0 \\
		0 & e^{2\Lambda(r)} & 0 & 0\dots & 0 & 0 \\
		0 & 0 & r^2\prod_{i=2}^{d-1}\sin^2(\xi_i) & 0\dots & 0 & 0 \\
		\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
		0 & 0 & 0 & \dots & r^2\sin^2(\xi_{d-1}) & 0 \\
		0 & 0 & 0 & \dots & 0 & r^2
	\end{pmatrix}.
\end{equation}
In the remainder of this subsection Greek indices will run over all spacetime dimensions, $r$ and $t$ will be reserved for the radial and time coordinates, and Latin indices will run only over all angular coordinates, where as always repeated indices are summed over; we shall also drop the explicit written dependence on $r$ for the unknown functions and all primes will represent derivatives with respect to $r$.

We shall make our choice of frame (which is analogous to a coordinate system for our purposes) such that we are using an orthonormal basis, ${\omega^\mu}$ (in most books an orthonormal frame is usually written as ${e^\mu}$ but as we are dealing with exponentials this would get confusing very quickly). These basis can be written explicitly as
\begin{align}
\omega^{t} &= e^{\Phi}dt 	\\
\omega^{r} &= e^{\Lambda}dr 	\\
\omega^{1} &= r \sin(\xi_{d-1})\sin(\xi_{d-2})\dots\sin(\xi_2) d\xi_1 	\\
\vdots 	\\
\omega^{k} &= r\sin(\xi_{d-1})\sin(\xi_{d-2})\dots \sin(\xi_{k+1})d\xi_k\\
\vdots \\
\omega^{d-1} &= r d\xi_{d-1}
\end{align}
with which we can compactly write $ds^2=\eta_{\mu\nu}\omega^\mu\otimes\omega^\nu$ where $\eta_{\mu\nu}$ is the usual Minkowski metric in $D$ dimensions.

The first step is to use Cartan's first structure equation which defines torsion, $T$, as
\begin{equation}
	T^{\mu}=d\beta^{\mu}+\Gamma^{\mu}_{\nu} \wedge \beta^{\nu} \label{eq:torsion}
\end{equation}
where $d$ is the exterior derivative (whose relevant properties we give below), $\beta$ is a general frame, $\Gamma$ is a 1-form connection (not a Christoffel symbol) and $\wedge$ is a \emph{wedge product} which allows for the product of \emph{forms}. The exact definition of the wedge product is not important to our calculations but the relevant property of it is \emph{graded commutativity}, i.e.
\begin{equation}
	\phi\wedge\psi=(-1)^{pq}\psi\wedge\phi
\end{equation}
where $\phi$ is a p-form and $\psi$ is a q-form; for our purposes we will only ever be using this with 1-forms so we will have $\omega^i\wedge\omega^{j}=-\omega^{j}\wedge\omega^{i}$ which importantly means that $\omega^i\wedge\omega^{i}=0$. The exterior derivative takes you from a p-form to a (p+1)-form and for our purposes you can consider to act similar to a usual derivative whilst adding a wedge product, although this is most definitely not true. The only other property we need to use is that 
\begin{equation}
	ddx=d^2x=0.
\end{equation}
Again, I have only supplied the relevant definitions and properties of these objects and they are much more deep and complicated than I have made them seem to be, but to properly introduce the required knowledge would require a much larger work.

Now, gravity is a torsion free theory which means we can rewrite (\ref{eq:torsion}) as
\begin{equation}
	d\omega^{\mu}=-\Gamma^{\mu}_{\nu}\wedge\omega^{\nu}. \label{eq:useful_torsion}
\end{equation}
Now that we have this form we can use it to start working out the connections;
\begin{align}
	d\omega^{t}&=d(e^\Phi)dt+e^\Phi \cancelto{0}{d^2t}\\
	&=\Phi^{\prime}e^{\Phi}dr\wedge dt\\
	&=\Phi^{\prime}e^{-\Lambda}\omega^{r}\wedge\omega^{t}=-\Phi^{\prime}e^{-\Lambda}\omega^{t}\wedge\omega^{r}
\end{align}
where in the last line we have used the definitions of $\omega^{r,t}$ to substitute in for $dr,dt$. Comparing this with (\ref{eq:useful_torsion}) gives us that 
\begin{equation}
	\Gamma^{t}_{r}=e^{-\Lambda}\Phi^{\prime} \omega^{t} + f_{tr}(r)\omega^{r}
\end{equation}
with $f_{tr}$ being an arbitrary function that arises from the fact that $\omega^i\wedge\omega^{i}=0$. This also means that 
\begin{equation}
	\Gamma^{t}_{k}= f_{tk}(r)\omega^{k}.
\end{equation}
Next it is easy to see that $d\omega^{r}=0$ and so we have that 
\begin{equation}
	\Gamma^{r}_{\alpha}=f_{r\alpha}(r)\omega^{\alpha}.
\end{equation}
The slightly trickier task is calculating $d\omega^k$ so let us split it into two pieces, the part that involves $r$ and the part that involves the angular coordinates;
\begin{align}
	d\omega^{k}&=dr\wedge\frac{1}{r}\omega^{k} + \text{Other terms}\\
	dr\wedge\frac{1}{r}\omega^{k} &= \frac{1}{r}e^{-\Lambda}\omega^{r}\wedge\omega^{k}
\end{align}
where in the first line we used that the $r$-derivative of $\omega^{k}$ is $\frac{1}{r}\omega^{k}$. This gives us
\begin{equation}
	\Gamma^{k}_{r}= \frac{1}{r}e^{-\Lambda}\omega^{k} + f_{kr}(r)\omega^{r}.
\end{equation}
Finally we have to calculate the part with the angular coordinates. Let us consider what happens on just one term first; the differential would act on one of the sine functions, creating a cosine and a $d\xi_j$ combining all the remaining sine functions and the factor of $r$ to be wedged with $d\xi_i$ (when $j>i$). We can recombine the $r$ and all the sine functions into the original $\omega^{i}$ but we are missing one sine and have an extra $\cos(\xi_{j}) d\xi_{j}$ which we can mostly solve by simply having each term multiplied by a cotangent function. The next step is to remove the $d\xi_{j}$ and replace it with an $\omega$ of some kind; considering that $j$ must be greater than $i$ it means that were $i$ to equal $(d-2)$ then the substitution would be easy as you get $d\xi_{d-1}=\frac{1}{r}\omega^{d-1}$, then for $i=(d-3)$ one would get $\xi_{d-2}=\frac{1}{r\sin(\xi_{d-1})}\omega^{d-2}$ and so on. Continuing this trend allows us to simply write that 
\begin{align}
	d\omega^{i}&=\frac{\cot(\xi_j)\omega^{j}\wedge\omega^{i}}{\sin(\xi_{d-1})\dots\sin(\xi_{j+1})}\\
	\implies \Gamma^{i}_{j} &= \frac{1}{r}\frac{\cot(\xi_j)}{\sin(\xi_{d-1})\dots\sin(\xi_{j+1})}\omega^{i} + f_{ij}(r)\omega^{j}.
\end{align}
Finally we can solve these generic functions by using the property that the connection is antisymmetric in lowered indices in an orthonormal basis, i.e.
\begin{equation}
	\Gamma_{\mu\nu}=-\Gamma_{\nu\mu}
\end{equation}
which with a little bit of thinking allows us to see that
\begin{equation}
	f_{\alpha\beta}(r)=
	\begin{cases}
		e^{-\Lambda}\Phi^{\prime}, &\text{if } \alpha=r \text{ and } \beta=t\\
		0, 	&\text{otherwise}
	\end{cases}
\end{equation}
which leaves us with 3 unique non-zero connections
\begin{align}
\Gamma^{t}_{r}&=\Gamma^{t}_{r}=	e^{-\Lambda}\Phi^{\prime} \omega^{t}\\
\Gamma^{k}_{r}&=-\Gamma^{r}_{k}=	\frac{1}{r}e^{-\Lambda}\omega^{k}\\
\Gamma^{i}_{j}&=-\Gamma^{j}_{i}=	\frac{1}{r}\frac{\cot(\xi_j)}{\sin(\xi_{d-1})\dots\sin(\xi_{j+1})}\omega^{i} \;\;\;\;\;\;\;\;\;\;\;(i<j).
\end{align}
Now that we have the connections we can use Cartan's second structure equation
\begin{equation}
	R^{\mu}_{\nu}=d \Gamma^{\mu}_{\nu}+\Gamma^{\mu}_{\lambda}\wedge \Gamma^{\lambda}_{\nu}
\end{equation}
to calculate the curvature.
\begin{align}
{R^{t}}_{r} &= d(e^{-\Lambda}\Phi^{\prime}\omega^{t}) + 0\\
			&= -e^{-\Lambda} \Lambda ^{\prime} \Phi ^{\prime} dr \wedge \omega^{t} + e^{-\Lambda}\Phi ^{\prime\prime}  dr \wedge \omega^{t} +e^{-\Lambda} \Phi ^{\prime} d\omega^{t}\\
			&=e^{-\Lambda}\left[ -\Lambda ^{\prime} \Phi ^{\prime} + \Phi ^{\prime\prime}\right]dr \wedge \omega^{t} + e^{-\Lambda}\Phi ^{\prime} (-\Gamma^{r}_{t}\wedge \omega^{r})\\
			&=e^{-\Lambda}\left[ -\Lambda ^{\prime} \Phi ^{\prime} + \Phi ^{\prime\prime}\right]e^{-\Lambda} \omega^{r}\wedge \omega^{t} +e^{-\Lambda}\Phi ^{\prime}  (-e^{-\Lambda}\Phi ^{\prime} \omega^{t}\wedge \omega^{r})\\
			&=e^{-2\Lambda}\left(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2\right) \omega^{r}\wedge \omega^{t}
\end{align}
where from the second to the third line we once again used (\ref{eq:useful_torsion}). Also we get that
\begin{align}
{R^{t}}_{k} &= \Gamma^{t}_{\lambda}\wedge \Gamma^{\lambda}_{k} = \Gamma^{t}_{r}\wedge \Gamma^{r}_{k}\\
			&= e^{-\Lambda}\Phi ^{\prime}  \omega^{t} \wedge -\frac{1}{r}e^{-\Lambda}\omega^{k}\\
			&= \frac{-1}{r}e^{-2\Lambda} \Phi ^{\prime} \omega^{t} \wedge \omega^{k}\\
{R^{r}}_{k} &=d \Gamma^{r}_{k}+\Gamma^{r}_{i}\wedge \Gamma^{i}_{k}.
\end{align}
If we quickly look at the last term it is easy to convince ourselves it is $0$ as $\Gamma^{r}_{i}$ contains an $\omega^{i}$ and so does $\Gamma^{i}_{k}$ and therefore it equates to zero. Continuing with the exterior derivative
\begin{align}
{R^{r}}_{k} &=d\left(\frac{1}{r}e^{-\Lambda}\omega^{k}\right)\\
			&= \frac{1}{r^{2}}e^{-\Lambda} dr\wedge \omega^{k}	+ \frac{1}{r}e^{-\Lambda}\Lambda ^{\prime} dr \wedge \omega^{k}	- \frac{1}{r}e^{-\Lambda} d\omega^{k}\\
			&=\left(\frac{1}{r^{2}}e^{-\Lambda} +\frac{1}{r}e^{-\Lambda}\Lambda ^{\prime} \right) dr \wedge \omega^{k} +\frac{1}{r}e^{-\Lambda}\left(\Gamma^{k}_{r} \wedge \omega^{r} + \Gamma^{k}_{j} \wedge \omega^{j}\right)\\
			&=\left(\frac{1}{r^{2}}e^{-\Lambda} +\frac{1}{r}e^{-\Lambda}\Lambda ^{\prime} \right)e^{-\Lambda}\omega^{r} \wedge \omega^{k} +\frac{1}{r}e^{-\Lambda}\left(\Gamma^{k}_{r} \wedge \omega^{r}\right)\\
			&=\left(\frac{1}{r^{2}}e^{-2\Lambda} +\frac{1}{r}e^{-2\Lambda}\Lambda ^{\prime} \right)\omega^{r} \wedge \omega^{k} +\frac{1}{r}e^{-2\Lambda}\omega^{k} \wedge \omega^{r}\\
			&= \frac{1}{r}e^{-2\Lambda} \Lambda ^{\prime} \omega^{r}\wedge \omega^{k}
\end{align}
where we dropped the last term in the third line as you can write that connection in terms of $\omega^{j}$. Finally we quote\cite{extra_dimension_thesis} without proof the result of the final calculation
\begin{equation}
	{R^{i}}_{j} = \frac{1}{r^{2}}(1-e^{-2\Lambda})\omega^{i}\wedge \omega^{j}.
\end{equation}
Next we will use the relation 
\begin{equation}
	{R^{\mu}}_{\nu} = {R^{\mu}}_{\nu\rho\sigma}\omega^{\rho}\wedge\omega^{\sigma}
\end{equation}
and the four symmetries of the Riemann tensor as given in (\ref{eq:riemann-prop-1}-\ref{eq:riemann-prop-3}) to give us the 4 unique Curvature tensors
\begin{align}
{R^{t}}_{rtr} &=-{R^{r}}_{trt} = - e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2)\\
{R^{t}}_{ktk} &=-{R^{k}}_{tkt} = -\frac{1}{r} \Phi ^{\prime} e^{-2\Lambda}\\
{R^{r}}_{krk} &= {R^{k}}_{rkr} = \frac{1}{r} e^{-2\Lambda} \Lambda ^{\prime} \\
{R^{i}}_{jij} &=				 \frac{1}{r^{2}}(1-e^{-2\Lambda}).
\end{align}
The penultimate step is to calculate the Ricci tensor
\begin{align}
R_{\mu\nu}&={R^{\lambda}}_{\mu\lambda\nu}\\
R_{tt} &= - e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2)+ \frac{1}{r} \Phi ^{\prime} e^{-2\Lambda}(d-1)\\
R_{rr} &= e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2) +\frac{1}{r} e^{-2\Lambda} \Lambda ^{\prime}(d-1)\\
R_{kk} &= \frac{1}{r} e^{-2\Lambda}(\Lambda ^{\prime} -\Phi ^{\prime} ) + \frac{1}{r^{2}}(1-e^{-2\Lambda})(d-2)
\end{align}
where the factors on the end terms come from the fact that $k$ can take $(d-1)$ values and $i$, whilst having the same range, is constrained to not equal $j$ so can take $(d-2)$ values. Finally we calculate the Ricci scalar $R={R^{\mu}}_{\mu}$
\begin{equation}
	R= -2e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2) +\frac{2}{r} e^{-2\Lambda}(\Lambda ^{\prime} -\Phi ^{\prime})(d-1)+ \frac{1}{r^{2}}(1-e^{-2\Lambda})(d-2)(d-1).
\end{equation}
It is now we can return fully to the methods we are used to and so we take the next step of calculating the Einstein tensor 
\begin{equation}
	G_{\mu\nu} = R_{\mu\nu}-\frac{1}{2}\eta_{\mu\nu}R.
\end{equation}
where we can use the Minkowski metric because of the basis we are working in. Unsurprisingly the only non-zero components are the diagonal elements which are given by
\begin{align}
G_{tt}&= \frac{1}{r} \Lambda ^{\prime} e^{-2\Lambda}(d-1) + \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-2)(d-1)\label{eq:einstein_tensor_higher_d_1} \\
G_{rr}&= \frac{1}{r} \Phi ^{\prime} e^{-2\Lambda}(d-1) - \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-2)(d-1) \\
G_{kk}&= e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2) -\frac{1}{r} e^{-2\Lambda}(\Lambda ^{\prime} -\Phi ^{\prime})(d-2)- \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-3)(d-2).\label{eq:einstein_tensor_higher_d_3}
\end{align}
\subsubsection{Schwarzschild Solution in Extra Dimensions} % (fold)
\label{sub:schwarzschild_solution_in_extra_dimensions}
Until now everything we have done is true for a spherically symmetric static solution but just as when we first derived the Schwarzschild metric, we now impose the condition that $G_{\mu\nu} =0$ which leaves us with two linearly independent equations and we shall pick the time and radial components of the Einstein tensor for convenience. By taking $G_{tt}+G_{rr}=0$ we can get that
\begin{equation}
	\Phi ^{\prime} +\Lambda ^{\prime} =0
\end{equation}
as we had for the 4-dimensional case; this means we can now write the line element as 
\begin{equation}
	ds^2 =-e^{-2\Lambda}dt^2 + e^{2\Lambda}dr^2 + r^2 d{\Omega}^2_{d-1}
\end{equation}
By solving $G_{tt}=0$ in a similar way to before we can work out what $e^{-2\Lambda}$ is equal to;
\begin{align}
	\frac{1}{r} \Lambda ^{\prime} e^{-2\Lambda}(d-1) + \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-2)(d-1)&=0\\
	2re^{-2\Lambda}\Lambda ^{\prime} +(d-2) -(d-2)e^{-2\Lambda}&=0\\
	-2r^{d-2}e^{-2\Lambda}\Lambda ^{\prime} +(d-2)r^{d-3}e^{-2\Lambda} &= (d-2)\\
	\left[r^{d-2}e^{-2\Lambda}\right]^{\prime} &=(d-2)\\
	\implies e^{-2\Lambda}&=1 + \frac{C}{r^{d-2}}
\end{align}
where $C$ is a constant of integration similar to previously that we will calculate next. Remembering that we made it so $d=D-1$ we can finally arrive at the (almost) complete line element of a Schwarzschild black hole in D dimensions
\begin{equation}
	ds^2 =-\left(1+\frac{C}{r^{D-3}}\right)dt^2 + \left(1+\frac{C}{r^{D-3}}\right)^{-1}dr^2 + r^2 d{\Omega}^2_{D-2}.
\end{equation}
\subsubsection{Schwarzschild Radius in Extra Dimensions} % (fold)
\label{sub:schwarzschild_radius_in_d_dimensions}
To find the value of the constant in our new metric and also find the Schwarzschild radius in D dimensions we start by looking at the higher dimensional gravitational acceleration in the Newtonian case
\begin{equation}
	g_{D}=\frac{d^{2}r}{dt^{2}}= -\frac{G_{D}M}{r^{D-2}} \label{eq:newt-extra}
\end{equation}
where $G_{D}$ is the gravitational constant in D dimensions which we can define as 
\begin{equation}
	G_{D}=\frac{(D-3)}{(D-2)}\frac{8\pi G_{N}^{(D)}}{\Omega_{D-2}}=\frac{(D-3)}{(D-2)}\frac{\Gamma\left(\frac{(D-1)}{2}\right)}{2\pi^{\frac{(D-1)}{2}}}8\pi G_{N}^{(D)}
\end{equation}
where $\Omega_{D-2}$ is the surface area of a $(D-2)$ sphere, $\Gamma(x)$ is the gamma function and $G_{N}^{(D)}$ is a constant proportional to the 4-dimensional gravitational constant, roughly speaking it is $G_{N}$ multiplied by the characteristic length of the extra dimensions to the power of $(D-4)$\cite{ newton_higher_dimension} i.e.
\begin{equation}
	G_{N}^{(D)}\approx G_{N} l^{m}
\end{equation}
where $l$ is the length, usually of the order of the radius of the extra dimensions, and $m=(D-4)$. Next we remember that the acceleration of a free particle can be calculated using the geodesic equation which we rewrite here for convenience 
\begin{equation}
	\frac{d^2x^\mu}{ds^2} + \Gamma^{\mu}_{\nu\rho}\frac{dx^\nu}{ds}\frac{dx^\rho}{ds} = 0.
\end{equation}
where $\frac{d^2x^\mu}{ds^2}$ is the acceleration of the particle. If we consider the particle to be far from the black hole ($r^{D-3}>>C$) then in this limit we have that the proper time is roughly equal to the coordinate time $s=t$. In this case we get that $\frac{dx^{\mu}}{ds}=(1,0,0,0)$ and then comparing the equations we see that
\begin{equation}
	g_{D}=\frac{d^{2}r}{dt^{2}}=-\Gamma^{r}_{tt}
\end{equation}
which we can extend to the near case. Now using the definition of the Christoffel symbols we get that
\begin{equation}
	g_{D}=-\Gamma^{r}_{tt}=\frac{1}{2}\frac{1}{g_{rr}}\partial_{r}g_{tt}.
\end{equation}
If we keep terms down to the order of $r^{-(D-2)}$ then we get that 
\begin{equation}
	g_{D}\approx \frac{(D-3)C}{2r^{D-2}}
\end{equation}
which comparing with (\ref{eq:newt-extra}) we can see that
\begin{equation}
	C_{2}=\frac{-2G_{D}}{(D-3)}.
\end{equation}
Combining this with our definition for $G_{D}$ we get the final form of our line element as 
\begin{equation}
	ds^2 =-f(r)dt^2 + \frac{1}{f(r)}dr^2 + r^2 d{\Omega}^2_{D-2}
\end{equation}
where the function $f(r)$ is given by
\begin{equation}
	f(r)=\left(1-\frac{16\pi G_{N}^{(D)}M}{(D-2)\Omega_{D-2}}\frac{1}{r^{D-3}}\right).
\end{equation}
This line element is true for all cases when $D\ge 4$ as when $D=3$ we get a flat space solution outside of the matter.

The final calculation we need to do is to calculate the actual Schwarzschild radius for D dimensions, that is when $f(r_{S,D})=0$ which we can solve with
\begin{equation}
	r_{S,D}=\left(\frac{16\pi G_{N}^{(D)}M}{(D-2)\Omega_{D-2}}\right)^\frac{1}{(D-3)}
\end{equation} 
which when $D=4$ recovers the usual Schwarzschild radius of $2G_{N}M$ as expected.
% subsection schwarzschild_radius_in_d_dimensions (end)
% subsection spherical_solution_in_extra_dimensions (end)
\subsubsection{Reissner-Nordstr\"om Solution in Extra Dimensions} % (fold)
\label{sub:reissner_nordstom_soltuion_in_extra_dimensions}
If instead we wanted to derive the solution of a charged black hole in extra dimensions we can extend our general electrostatic SE tensor from 4 dimensions to $D=d+1$ and staying in the orthonormal basis, $\omega^{\mu}$, this is written as
\begin{equation}
		T_{\mu\nu}= 
		2\Psi ^{\prime 2} e^{4\Psi(r)}\begin{pmatrix}
		1 & 0 & 0 & 0\dots & 0 & 0 \\
		0 & -1 & 0 & 0\dots & 0 & 0 \\
		0 & 0 & 1 & 0\dots & 0 & 0 \\
		\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
		0 & 0 & 0 & \dots & 1 & 0 \\
		0 & 0 & 0 & \dots & 0 & 1
	\end{pmatrix}.
\end{equation}
By using this combined with the Einstein tensors as written in (\ref{eq:einstein_tensor_higher_d_1}-\ref{eq:einstein_tensor_higher_d_3}) we can see that as always the equation for the  time component is equal to minus the radial component which gives us the usual $\Lambda ^{\prime}  = - \Phi ^{\prime}$ and $e^{2\Phi}=e^{-2\Lambda}$. Now equating $G_{tt}$ to $G_{kk}$ we get that
\begin{multline}
\frac{1}{r} \Lambda ^{\prime} e^{-2\Lambda}(d-1) + \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-2)(d-1) = e^{-2\Lambda}(\Phi ^{\prime\prime} - \Phi ^{\prime} \Lambda ^{\prime} +{\Phi ^{\prime}}^2) \\-\frac{1}{r} e^{-2\Lambda}(\Lambda ^{\prime} -\Phi ^{\prime})(d-2)- \frac{1}{2r^{2}}(1-e^{-2\Lambda})(d-3)(d-2)
\end{multline}
which if we write everything in terms of $\Phi$ and by noticing the first term on the right hand side is the differential of a product we get a slightly nicer
\begin{equation}
	\frac{d}{dr}(e^{2\Phi}\Phi ^{\prime} )+ \frac{1}{r}e^{2\Phi}\Phi ^{\prime}(3d-5) - \frac{1}{r^{2}}(1-e^{2\Phi})(d-2)^{2}=0.\label{eq:higher_d_em_to_solve}
\end{equation}
By trying an ansatz of the form
\begin{align}
	e^{2\Phi}=f(r)&=A+\frac{B}{r^{d-2}}+\frac{C}{r^{2(d-2)}}\\
	f ^{\prime} &= \frac{-B(d-2)}{r^{d-1}} -\frac{C(2d-4)}{r^{2d-3}}\\
	f ^{\prime\prime} &= \frac{B(d^{2}-3d+2)}{r^{d}} + \frac{C(4d^{2}-14d+12)}{r^{2d-2}} 
\end{align}
when plugged into (\ref{eq:higher_d_em_to_solve}) gives us the result that $A=1$ leaving us with two unknowns. $B$ is obviously our higher dimensional Schwarzschild term leaving us to solve only for $C$. To do this we just need to substitute our solution into the EFEs for any component so we shall do it for the $tt$ component. Working through the cancellations we get that 
\begin{align}
	\frac{(D-3)(D-2)C}{2r^{(2D-4)}}&=16\pi G_{N}^{(D)} \Psi ^{\prime 2} e^{4\Psi}\\
	\implies \frac{C}{r^{(2D-4)}}&= \frac{8\pi G_{N}^{(D)}}{(D-3)(D-2)} \left(\frac{d}{dr}e^{2\Psi}\right)^{2}
\end{align}
where the differential term is the electric field. It can be shown, by use of QFT or even by an extension of Gauss' Law\cite{Zwiebach:789942}, that the electric potential in higher dimensions is of the form 
\begin{equation}
	A_{0}=\frac{1}{(D-3)\Omega_{D-2}}\frac{q}{r^{D-3}}=e^{2\Psi}
\end{equation}
which gives us that
\begin{equation}
	\frac{C}{r^{(2D-4)}}= \frac{8\pi G_{N}^{(D)}}{(D-3)(D-2)\Omega_{D-2}} \frac{q^{2}}{r^{(2D-4)}}.
\end{equation}
Rearranging we see that
\begin{equation}
	C=q^{2}\frac{8\pi G_{N}^{(D)}}{(D-3)(D-2)\Omega_{D-2}}\equiv{G_{N}^{(D)}Q^{2}}
\end{equation}
and this means we can write our complete line element as
\begin{equation}
	ds^2 =-\left(1- \frac{r_{S,D}^{D-3}}{r^{D-3}}+ \frac{r_{Q,D}^{2(D-3)}}{r^{2(D-3)}}	\right)dt^2 + \left(1- \frac{r_{S,D}^{D-3}}{r^{D-3}}+ \frac{r_{Q,D}^{2(D-3)}}{r^{2(D-3)}}	\right)^{-1}dr^2 + r^2 d{\Omega}^2_{D-2} \label{eq:RN_higher_dim_metric}
\end{equation}
where 
\begin{equation}
	r_{Q,D}=\left(G_{N}^{(D)}Q^{2}\right)^{\frac{1}{2(D-3)}}.
\end{equation}
% subsection reissner_nordstom_soltuion_in_extra_dimensions (end)
\subsection{Other Black Hole Solutions} % (fold)
\label{sub:other_black_hole_solutions}
% Mention the Kerr and the Kerr-Newman solutions, maybe write their metrics?
As well as the solutions we have given above there exists a further class of solutions that also consider the case of a rotating black hole, the Kerr black hole, as well as a charged and rotating case, the Kerr-Newman black hole. We write them here in 4 dimensions just for completeness but shall not consider either of the cases in the remainder of this work. First the Kerr metric in spherical polar coordinates is given by
\begin{multline}
	ds^{2}=-\left(1- \frac{r_{s}r}{\rho^{2}}\right)dt^{2}+\frac{\rho^{2}}{\Delta}dr^{2}+\rho^{2}d\theta^{2}+\\
	+\left(r^{2}+\alpha^{2}+\frac{r_{s}r\alpha^{2}}{\rho^{2}}\sin^{2}\theta\right)\sin^{2}\theta d\phi^{2}- \frac{2r_{s}2\alpha\sin^{2}\theta}{\rho^{2}}dtd\phi
\end{multline}
where $r_{s}$ is the Schwarzschild radius and 
\begin{align}
	\alpha=\frac{J}{M}\\
	\rho^{2}=r^{2}+\alpha^{2}\cos^{2}\theta\\
	\Delta=r^{2}-r_{s}r+\alpha_{2}
\end{align}
with $J$ being the angular momentum for the black hole. The Kerr-Newman case is
\begin{multline}
	ds^{2}=-\frac{\Delta-\alpha^{2}\sin_{2}\theta}{\rho^{2}}dt^{2}    +\frac{\rho^{2}}{\Delta}dr^{2}   +\rho^{2}d\theta^{2}+\\
	+(r^{4}+ 2r^{2}\alpha^{2}+ \alpha^{4}- \alpha^{2}\Delta\sin^{2}\theta)\frac{\sin^{2}\theta d\phi^{2}}{\rho^{2}}   + (\Delta -r^{2}-\alpha^{2})\frac{2\alpha\sin^{2}\theta dt d\phi}{\rho^{2}}
\end{multline}
where everything is as before except for $\Delta$ which now is defined as
\begin{equation}
	\Delta=r^{2}-r_{s}r+\alpha_{2} +r^{2}_{Q}
\end{equation}
with $r^{2}_{Q}$ being equal to $Q^{2}G$ as before.

Finally whilst we have only considered spherical solutions in higher dimensions it is possible to have different topological solutions in higher dimensions so that instead of a spherical black hole you could have a \emph{black ring} with $S^{1}\times S^{D-3}$ topology as shown in 5 dimensions by Emparan and Reall in 2002\cite{Emparan:2001wn}, or even a black hole with a ring surrounding it known as a \emph{black Saturn}. Working in higher dimensions greatly reduces the constraints on black hole topologies and the uniqueness theorem breaks down as well meaning that the number of possible solutions is not as well constrained; however we will work with just the solutions we have derived and leave it to the reader to read further about these solutions should they wish.
% subsection other_black_hole_solutions (end)

% section black_hole_solutions_to_efes (end)

\section{Black hole Thermodynamics} % (fold)
\label{sec:blackhole_thermodynamics}
\subsection{Area of 4-Dimensional Black Holes} % (fold)
\label{sub:area_of_black_holes}
We model all black holes as spheres and as such they obey the usual equation for their area. We take the radius of the black hole to be the radius of the event horizon(s); so for the Schwarzschild black hole we have that 
\begin{equation}
	A_s=4\pi r_s^2.
\end{equation}
To solve for the Reissner-Nordstr\"om case we must first find the radius of the horizons i.e. where $g_{tt}=0$; solving this quadratic equation gives us the solutions that
\begin{equation}
	r_\pm=\frac{r_s\pm \sqrt{r_s^2-4r_q^2}}{2}
\end{equation}
where once again we remind ourselves that $r_q^2=\frac{Gq^2}{4\pi}\equiv GQ^2$. Looking at this we see that we have 3 cases
\begin{enumerate}
	\item 2 horizons if $r_s^2-4r_q^2>0$
	\item 0 horizons if $r_s^2-4r_q^2<0$
	\item 1 horizon if $r_s^2-4r_q^2=0$
\end{enumerate}
where the third case is known as the extremal case which has interesting thermodynamic properties and occurs when $GM^2=Q^2$. The case of no horizons is what is known as a \emph{naked singularity} and whilst not forbidden by GR it is excluded as it is not possible in EM to have an object with it's charge greater than its mass. Again this leads us to a simple expressions for the area of the black hole
\begin{equation}
	A^{(\pm)}_{RN}=4\pi r_{\pm}.
\end{equation}
% subsection area_of_black_holes (end)
\subsection{Surface Gravity of 4-Dimensional Black Holes} % (fold)
\label{sub:surface_gravity_of_black_holes}
The surface gravity of an object, $\kappa$, is defined as 
\begin{equation}
	\kappa^2=\frac{1}{2}(D^a\chi^b)(D_a\chi_b)
\end{equation}
where the expression is to be evaluated at the horizon and $\chi$ is a Killing vector which satisfies the relation that
\begin{equation}
	D_{\mu}\chi_{\nu}+D_{\nu}\chi_{\mu}=0;
\end{equation}
another way of saying this is that if you shift the metric $g$ a small amount by $\chi$ then $g$ remains unchanged. The Killing vector for both our cases is $\chi^a \sim(1,0,0,0)$ as the metrics are independent of $t$ which also means that $\chi_a \sim (-(1-\frac{r_s}{r},0,0,0)$. To calculate the covariant derivative we only need to calculate the Christoffel symbols that have $t$ as one or more of their indices of which for the Schwarzschild solution there are only two
\begin{align}
\Gamma^{t}_{tr}&=\frac{1}{2}\frac{r_s}{r^2-rr_s}=\frac{1}{2}\left(1- \frac{r_s}{r}\right)^{-1}\frac{r_s}{r^2} \\
\Gamma^{r}_{tt}&=\frac{1}{2}\frac{rr_s-r_s^2}{r^3}=\frac{1}{2}\left(1- \frac{r_s}{r}\right)\frac{r_s}{r^2}.
\end{align}
The two non-zero derivatives are therefore
\begin{align}
D^r\chi^t&=D_r\chi^t g^{rr}=\frac{1}{2}\frac{r_s}{r^2}\\
D^t\chi^r&=D_t\chi^r g^{tt}=-\frac{1}{2}\frac{r_s}{r^2}
\end{align}
from which we can see that $D^r\chi^t=-D^t\chi^r$ and due to the fact that $g_{tt}=-(g_{rr})^{-1}$ we also have that $D_t\chi_r=-D^t\chi^r$ and $D_r\chi_t=-D^r\chi^t$. Putting this all together we can arrive at the final result
\begin{align}
	\kappa^2=(D^t\chi^r)^2&=\left. \frac{r_s^2}{4r^4}\right|_{r=r_s} = \frac{1}{4r_s^2}\\
	&\implies \kappa_s=\frac{1}{2r_s}.
\end{align}
Applying the same methods to the charged case we again calculate the non-zero, t-indexed, Christoffel symbols
\begin{align}
\Gamma^{t}_{tr}&=\frac{1}{2}\left(1- \frac{r_s}{r}- \frac{r_q^2}{r^2}\right)^{-1}\frac{rr_s-2r^2_q}{r^3} \\
\Gamma^{r}_{tt}&=\frac{1}{2}\left(1- \frac{r_s}{r}- \frac{r_q^2}{r^2}\right)\frac{rr_s-2r^2_q}{r^3}
\end{align}
giving the equation for the surface gravity as 
\begin{align}
	\kappa^2&=(D^t\chi^r)^2=\frac{(rr_s-2r_q^2)^2}{4r^6}\\
	&\implies \kappa_{RN}^{(\pm)}=\left. \frac{rr_s-2r_q^2}{2r^3}\right|_{r=r_\pm}.
\end{align}
Evaluating this in the case where $r=r_\pm$ gives us a seemingly \emph{ugly} expression of
\begin{equation}
	\kappa_{RN}^{(\pm)}=2 \;\frac{r_s^2\pm r_s\sqrt{r_s^2-4r_q^2}-4r_q^2}{\left(r_s\pm\sqrt{r_s^2-4r_q^2}\right)^3}.
\end{equation}
By noticing the similarity of the first and last term in the numerator with those inside the square root we find the it is possible to rewrite this expression as
\begin{align}
	\kappa_{RN}^{(\pm)}&=2 \;\frac{\left(r_s\pm\sqrt{r_s^2-4r_q^2}\right)\left(\pm\sqrt{r_s^2-4r_q^2}\right)}{\left(r_s\pm\sqrt{r_s^2-4r_q^2}\right)^3}\\
	&=2\;\frac{(r_\pm-r_\mp)}{4r_\pm^2}\\
	\implies \kappa_{RN}^{(\pm)}&=\frac{r_\pm - r_\mp}{2r_\pm^2}.
\end{align}
The most interesting feature of this quantity is that in this form it is easy to see what happens when $r_+=r_-$, i.e. in the extremal case; we can see that the surface gravity at the horizon of such a black hole is zero, something that will become even more interesting when we make the connection between surface gravity and temperature in the next section.
% subsection surface_gravity_of_black_holes (end)
\subsection{Temperature and Entropy of 4-Dimensional Black Holes} % (fold)
\label{sub:temperature_and_entropy}
The reason for calculating these properties of a black hole is because these are related to thermodynamic variables we can associate with the objects. It is important to note that these are the \emph{effective} properties of a black hole; specifically the temperature of the black hole is just a mathematical concept as the horizon is not itself an object. Starting with the entropy; the motivation for black holes having entropy was provided by Wheeler who proposed a creature (later called Wheeler's demon) that could sit outside a black hole and drop in objects with entropy which would decrease the total entropy outside the black hole, the accessible universe, and therefore violate the second law of thermodynamics. It was later shown by Floyd and Penrose, and Christodoulou that the area of a black hole irreversibly increased under nearly any transformation and it was finally shown in 1971 by Hawking that the area of a black hole never decreased classically. It was finally Bekenstein who suggest that there was a relation between the area of a black hole and its entropy which would solve the violation of the second law which was later verified by Hawking and as such the entropy $S_{BH}$ is often referred to as the \emph{Bekenstein-Hawking} entropy. Hawking also showed, by considering the collapse of a star into a black hole and then comparing this to a black body, that the temperature of a black hole was proportional to the surface gravity at the horizon. Whilst the coefficients of of these relations can be derived mathematically we shall be satisfied by finding them via a dimensional argument followed by the requirement that the first law be satisfied to find any remaining numerical factors. For the remainder of this subsection we will return to SI units meaning that, as an example, $c\ne 1$.
\subsubsection{Bekenstein-Hawking Entropy of a Black Hole} % (fold)
\label{sub:bekenstein_hawking_entropy_of_a_black_hole}
The surface are of a black hole has units of $m^2$ whilst entropy has units of $Kg\;m^{2}\;s^{-2}\;K^{-1}$ which means that for
\begin{equation}
	S_{BH}=\alpha A
\end{equation}
we have that $\alpha$ has units of $\frac{Kg}{s^{2}K}$. The only physical constant containing Kelvin is the Boltzmann constant $k_{b}$ which has total units of $\frac{Kg\;m^{2}}{s^{2}K}$ which means we now can write that
\begin{equation}
	S_{BH}=\beta k_{b}A
\end{equation}
where $\beta$ will now need to have units of $m^{-2}$. The only constant with units of just length is the Planck length, $l_{p}$ which is defined as
\begin{equation}
	l_{p} =	\sqrt{\frac{\hbar G}{c^{3}}}.
\end{equation}
This gives us the almost final expression for the entropy
\begin{equation}
	S_{BH}\propto \frac{k_{b}A}{l_{p}^{2}}.
\end{equation}
where the constant of proportionality will be determined from the first law. The appearance of $\hbar$ in this equation was surprising as it implied some quantum effects for an entirely classical object and approach.
\begin{figure}
	\centering
        \includegraphics[scale=0.5]{Entropy.pdf}
	\caption{Entropy of the 3 black hole horizons; notice that the extremal case is when the outer and inner horizons of the RN are equal, and has a non-zero entropy.}
	\label{fig:entropy}
\end{figure}
Figure (\ref{fig:entropy}) plots the shape of the black hole entropies as a function of mass, and in the RN case with a fixed charge, then we can see that the entropy for the Schwarzschild case increases with increasing mass as expected, as does the outer RN horizon. The inner horizon decreases in entropy for an increase in mass, which makes sense as it will get smaller and smaller with added mass, approaching but never reaching the singularity (where the entropy would be zero). The point at which these two curves meet is the extremal RN case i.e. when $GM^{2}=Q^{2}$. Note that the entropy for the RN black hole is always less than that of the Schwarzschild which is as it should be because due to the charge there would be less microstates giving rise to the same macrostate and therefore a lower entropy. 
\subsubsection{Hawking Temperature and The First Law} % (fold)
\label{sub:hawking_temperature_and_the_first_law}
Whilst we took the time with the entropy to make a point of the involvement of $\hbar$ we can just find the coefficient for the temperature by using Plank units meaning that for
\begin{equation}
	T_{H} = \gamma\kappa
\end{equation}
we have that the units of $\gamma$ must be $\frac{s^{2}}{m}K$. Using Plank units this gives us that
\begin{equation}
	\left(\sqrt{\frac{\hbar G}{c^{5}}}\right)^{2}\sqrt{\frac{c^3}{\hbar k_{b}^{2}}}\sqrt{\frac{\hbar c^{5}}{G k_{b}^{2}}}= \frac{\hbar}{c k_{b}}
\end{equation}
then we have the expression that 
\begin{equation}
	T_{H}\propto \frac{\hbar}{c k_{b}}\kappa.
\end{equation}
Next to find the proportionality constants we have to turn to the first law of thermodynamics which in infinitesimal form is written as
\begin{equation}
	dE=TdS -PdV
\end{equation}
where $P$ is the pressure and $V$ is the volume. It turns out that the pressure in this equation is given by the cosmological constant which in our case is $0$ and as such we can remove this term. By substituting our expressions into this equation we get that
\begin{equation}
	c^{2}dM= \frac{\hbar}{c k_{b}}\kappa \frac{k_{b}c^{3}}{\hbar G}dA \alpha
\end{equation}
where $\alpha$ is our numerical constant that we are trying to solve for. By working with the Schwarzschild solution and plugging the factors of $c$ into the area and surface gravity such that 
\begin{align}
	A_{s} &= 16\pi \frac{G^{2}M^{2}}{c^{4}} \\
	\implies dA_{s} &= 16\pi \frac{G^{2}}{c^{4}} 2dM\\
	\kappa_{s} &= \frac{c^{4}}{4GM}
\end{align}
then by cancelling down we get the result that
\begin{equation}
	dM=8\pi \alpha dM
\end{equation}
which means that $\alpha=\frac{1}{8\pi}$. It turns out the factors are $\frac{1}{4}$ for the entropy relation and $\frac{1}{2\pi}$ for the temperature giving us
\begin{align}
	S_{BH}=\frac{k_{b}c^{3}A}{4\hbar G}  \label{eq:entropy}\\
	T_{H}= \frac{\hbar\kappa}{2\pi ck_{b}}.
\end{align}
\begin{figure}
	\centering
        \includegraphics[scale=0.6]{Temperature_all.pdf}
	\caption{Temperature of the three black hole horizons; the shape of the outer RN curve is not obvious at this zoom, but can be seen to tend towards the Schwarzschild temperature for a high mass.}
	\label{fig:temp_all}
\end{figure}
To show that these relations hold for the charged case we can to show that they satisfy the first law with a charged particle as well. Starting with taking the entropy for the $r_{+}$ horizon of the RN black hole
\begin{equation}
	S= \frac{k_{b}4\pi r_{+}^{2}}{4 l_{p}^{2}}=\frac{k_{b}\pi}{l_{p}^{2}c^{4}}\left(GM+\sqrt{G^{2}M^{2}-GQ^{2}}\right)^{2}
\end{equation}
then when we take the variation of the entropy we get that
\begin{align}
	dS&=\left(\frac{k_{b}\pi}{l_{p}^{2}c^{4}}\right)2\left(GM+\sqrt{G^{2}M^{2}-GQ^{2}}\right)\left[GdM+\frac{2MG^{2}dM}{2\sqrt{G^{2}M^{2}-GQ^{2}}}+\frac{-2GQdQ}{2\sqrt{G^{2}M^{2}-GQ^{2}}}\right]\\
	&=\frac{k_{b}2\pi}{Gc\hbar}\left(\frac{G(GM+\sqrt{G^{2}M^{2}-GQ^{2}})^2}{\sqrt{G^{2}M^{2}-GQ^{2}}}dM - \frac{GQ(GM+\sqrt{G^{2}M^{2}-GQ^{2}})}{\sqrt{G^{2}M^{2}-GQ^{2}}}dQ\right).
\end{align}
Next using the definition of the surface gravity for the RN case we can write this as
\begin{equation}
	dS=\frac{k_{b}2\pi c}{\hbar}\left(\frac{1}{\kappa}c^{2}dM- \frac{Q}{\kappa r_{+}}dQ\right)
\end{equation}
which using our definition of the temperature comes out as
\begin{equation}
	TdS=dE-\Phi dQ
\end{equation}
where $\Phi=\frac{Q}{r_{+}}$ which can also be though of as 
\begin{equation}
	\Phi=\left. -A_{0}\right|_{r=r_{+}},
\end{equation}
the electric potential of the horizon, which means we have arrived at the first law for the case of a charged particle. 
\begin{figure}
	\centering
        \includegraphics[scale=0.6]{Temperature_RN_outer.pdf}
	\caption{Temperature of the outer RN horizon}
	\label{fig:temp_outer_RN}
\end{figure}

Figure (\ref{fig:temp_all}) shows the curve of temperature for all black hole horizons (with again a fixed charge for the RN case); the most striking feature of this is that for the Schwarzschild the temperature decreases with increasing mass, something that will be discussed more in the next section. Next, figure (\ref{fig:temp_outer_RN}) shows an enlarged version of just the curve for the outer RN horizon. This graph features two distinct regions; the \emph{near-extreme} region which is from the start of the curve to the maxima peak, and the \emph{far, non-extreme} region which is everything after the peak. The second region behaves similarly to the Schwarzschild case where an increase in mass causes a decrease in temperature which makes sense as this is the $M>>Q$ limit. The near-extreme region however behaves as one would classically expect; an increase in mass gives an increase in temperature. The final key point is that at the extremal limit the temperature of the horizon is equal to zero. 
% subsection hawking_temperature_ (end)
% subsection bekenstein_hawking_entropy_of_a_black_hole (end)
% subsection temperature_and_entropy (end)
% subsection discussion_of_ (end)
\subsection{Specific Heat Capacity in 4 Dimensions} % (fold)
\label{sub:specific_heat_capacity_in_4_dimensions}
Returning now back to our usual units where $\hbar = c = k_b = 1$ we can consider the specific heat of black holes. Specific heat is defined as the change in heat with respect to a change in temperature i.e. $\frac{dE}{dT}$. Fortunately for us the change in heat is proportional to the change in energy which means the change in mass, and the change in temperature is proportional to the change in surface gravity. By plotting the temperature as a function of mass we can see the behaviour of the specific heat, i.e if it is positive or negative. Returning to the graph shown in figure (\ref{fig:temp_all}) we can see the specific heat for the black hole is negative which is very unphysical; the more energy you add the colder the black hole seems to get! In the low mass limit a small change in mass gives a very large decrease in temperature and at the large mass limit it takes a large change in mass to produce any change in temperature. Looking now again at figure (\ref{fig:temp_outer_RN}) we can speak more specifically about the specific heat of this horizon in the two regions. We have plotted temperature on the y-axis and as such the gradient of this graph is actually the inverse of the specific heat which means that at some critical mass the specific heat changes from $\infty$ to $-\infty$ (i.e the peak of the curve), at the extremal point the specific heat is 0 (implying that this is a stable state of the system where adding energy would not increase the temperature which is consistent with the third law of thermodynamics), and finally that at an infinite mass the specific heat is once again (negatively) infinite meaning it is impossible to ever reach a zero temperature black hole by just adding more (uncharged) mass, another agreement with the third law if in a slightly counter-intuitive situation. 
% subsection specific_heat_capacity_in_4_dimensions (end)
\subsection{Extending these Quantities to Extra Dimensions} % (fold)
\label{sub:extending_these_quantities_to_higher_dimensions}
Extending the temperature and entropy to higher dimensions requires us to extend the definitions of area and surface gravity with the later being only slightly more difficult than the first. Just as when we wanted the surface area of the horizon for the 4-dimensional case we calculated it as the surface of a 2-sphere of radius $r$ we can find the area of a D-dimensional black hole horizon by calculating the surface of a $(D-2)$-sphere
\begin{equation}
	A^{(D)}=\Omega_{(D-2)}r^{D-2}=\frac{2\pi^{(\frac{D-1}{2})}}{\Gamma(\frac{D-1}{2})}r^{D-2}
\end{equation}
which dimensionally is still correct as the gravitational constant in the Planck length also gains a dimension of length for each extra dimension we add. As for the surface gravity, if we were to write the function that multiplies the $dt^{2}$ as $f(r)$ then it is easy to show that by following the same logic as when calculating $\kappa$ for the 4-dimensional case that we can write the surface gravity more generally as
\begin{equation}
	\kappa=\left. \frac{f'(r)}{2}\right|_{r=horizon}.
\end{equation}
As we only concerned ourselves with the sign of the specific heat and as this won't change (as only the power and not the sign of the mass terms changes) we shall not be considering this quantity more specifically in higher dimensions but believe it enough to say that the sign of the quantity will remain unchanged.

\section{Instability of Extremal RN Black Hole in 4D} % (fold)
\label{sec:intability_of_extremal_rn_black_hole}
Following on from the consideration of the specific heat of an extremal RN black hole i.e. that at 0 temperature it should be stable under the introduction of a massless scalar perturbation, we can test this more rigorously and as we will see it turns out not to be stable under a general perturbation. There are a few tools we need to discuss/derive before we can show this instability which we will finally show from the work of Aretakis\cite{Aretakis:2011ha}\cite{Aretakis:2011hc} as given clearly by Lucietti et al\cite{Lucietti:2012xr}.
\subsection{Non-Extreme Instability} % (fold)
\label{sub:non_extreme_instability}
We will quickly mention here why we do not concern ourselves with this issue when it comes to non-extreme black holes. Whilst outside the horizon the energy density of a spacelike surface is non-increasing and as this is a function of $\partial_{\mu}\psi$ it means all behaves as expected; when evaluated on the horizon the energy density becomes degenerate and the constraints on $\psi$ disappear. Fortunately this behaviour does not need to be considered by using the horizon redshift, a quantity proportional to $e^{-\kappa v}$ where $\kappa$ is our surface gravity and $v$ is our Killing time vector. By doing this it has been shown that $\psi$ and all its derivatives decay on and outside of the horizon.
% subsection non_extreme_instability (end)
\subsection{Klein Gordon Equation in Curved Spacetime} % (fold)
\label{sub:klein_gordon_equation_in_curved_spacetime}
Before we can consider adding a perturbation we need to know the equation that governs it; as we will work with a massless scalar this is the Klein-Gordon (KG) equation. The Lagrangian density of the KG equation in Minkowski space is 
\begin{equation}
	\Lagr =\frac{1}{2}\eta^{\mu\nu}(\partial_{\mu}\psi)(\partial_{\nu}\psi) -\frac{1}{2}m^{2}\psi^{2}
\end{equation}
which to extending to curved space would involve promoting the Minkowski metric to the general metric tensor and replacing all partial derivatives with covariant derivatives; however the covariant derivative of a scalar field is just the partial derivative so these just remain as $\partial_{\mu}\psi$. This would give us the minimally coupled KG equation, however there is another scalar term we could couple to $\psi^{2}$ that would become 0 in flat space, the Ricci scalar. Putting all this together gives us that the new Lagrangian density is
\begin{equation}
	\Lagr =\frac{1}{2}g^{\mu\nu}(\partial_{\mu}\psi)(\partial_{\nu}\psi) -\frac{1}{2}(m^{2}+\xi R)\psi^{2} \label{eq:KG_action}
\end{equation}
where $\xi$ is some general coupling constant. Remembering that the action becomes
\begin{equation}
	S=\int d^{4}x \sqrt{-g}\Lagr
\end{equation}
we can get the Euler-Lagrange equations by requiring $\delta S=0$ which then gives us
\begin{equation}
	\partial_{\mu}\frac{\partial\sqrt{-g}\Lagr}{\partial(\partial_{\mu}\psi)}-\frac{\partial\sqrt{-g}\Lagr}{\partial\psi}=0.
\end{equation}
By plugging (\ref{eq:KG_action}) into these and using the fact that $\sqrt{-g}$ does not depend on $\psi$ or $\partial_{\mu}\psi$ we get that
\begin{equation}
	\partial_{\mu}(\sqrt{-g}g^{\mu\nu}\partial_{\nu}\psi)+\sqrt{-g}(m^{2}+\xi R)\psi=0.
\end{equation}
When we are in the case of minimal coupling and a massless field we can write this as
\begin{equation}
	\frac{1}{\sqrt{-g}}\partial_{\mu}(\sqrt{-g}g^{\mu\nu}\partial_{\nu}\psi)=(D_{\mu}g^{\mu\nu}\partial_{\nu})\psi=\nabla^{2}\psi=0 \label{KG_equation_curved}
\end{equation}
where $\nabla$ is the Laplace-Beltrami operator and this is then the form of the KG equation that we will use.
% subsection klein_gordon_equation_in_curves_spacetime (end)
\subsection{Eddington Finkelstein Coordinates} % (fold)
\label{sub:rn_black_hole_in_eddington_finkelstein_coordinates}
As previously mentioned, the singularity at the horizon for different black holes is only a coordinate singularity rather than a pure spacetime singularity and as such a change of coordinates can remove this singularity. The Eddington Finkelstein (EF) coordinates are ones where radially null geodesics define surfaces of constant time, i.e. in/outgoing radial light rays define lines of constant time. We start to make the change in coordinates by defining a tortoise coordinate $r^{*}$ where 
\begin{equation}
	\frac{dr^{*}}{dr}=g_{rr}=\left(1- \frac{2GM}{r}\right)^{-1}
\end{equation}
for the Schwarzschild case. It can be checked that 
\begin{equation}
	r^{*}=r+2GM\ln\left|\frac{r}{2GM}-1\right|
\end{equation}
satisfies this relation. We then define the ingoing EF coordinate, $v=t+r^{*}$ which allows us to write the line element as
\begin{equation}
	ds^{2}=-\left(1- \frac{2GM}{r}\right)dv^{2}+2dvdr+r^{2}d\Omega^{2}_{2}
\end{equation}
where the horizon is still at $r_{s}=2GM$ but we can see this no longer causes a singularity. Moving now to the extremal RN case we first need to evaluate $g_{tt}=g_{rr}^{-1}$ at $Q^{2}=GM^{2}$
\begin{equation}
	\left.\left(1- \frac{2GM}{r}+ \frac{GQ^{2}}{r^{2}}\right)\right|_{GM^{2}=Q^{2}}=\left(1- \frac{GM}{r}\right)^{2}.
\end{equation}
Again requiring that
\begin{equation}
	\frac{dr^{*}}{dr}=\left(1- \frac{GM}{r}\right)^{-2}
\end{equation}
gives us the solution that
\begin{equation}
	r^{*}=r +2GM\ln\left|\frac{r-GM}{GM}\right| -\frac{G^{2}M^{2}}{r-GM}
\end{equation}
which using the same definition for $v$ leaves our line element as
\begin{equation}
	ds^{2}=-\left(1- \frac{GM}{r}\right)^{2}dv^{2}+2dvdr+r^{2}d\Omega^{2}_{2}.
\end{equation}
% subsection rn_black_hole_in_eddington_finkelstein_coordinates (end)
\subsection{Aretakis Instability} % (fold)
\label{sub:aretakis_instability}
To see the Aretakis instability we will look at the equation of motion for a massless scalar field, $\psi$, in an extremal RN background. This means studying the KG equation in the case of the metric being the extremal RN metric. The definition of $\nabla^{2}$ has both the inverse metric and the determinant of the metric in it; to calculate these we will use two properties of block diagonal matrices. The first property is that the inverse of a block diagonal matrix is simply the matrix with each block being replaced by the inverse of that block so if
\begin{equation}
		M= 
		\begin{pmatrix}
		A & 0 & 0 & 0\\
		0 & B & 0 & 0\\
		0 & 0 & C & 0\\
		0 & 0 & 0 & D\\
	\end{pmatrix}
\end{equation}
where $A-D$ are smaller matrices then
\begin{equation}
		M^{-1}= 
		\begin{pmatrix}
		A^{-1} & 0 & 0 & 0\\
		0 & B^{-1} & 0 & 0\\
		0 & 0 & C^{-1} & 0\\
		0 & 0 & 0 & D^{-1}\\
	\end{pmatrix}.
\end{equation}
The second property we will use is that the determinant of a block diagonal matrix is also the product of the determinants of each block i.e.
\begin{equation}
	\det(M)=\det(A)\det(B)\det(C)\det(D).
\end{equation}
This allows us to write our inverse metric 
\begin{equation}
	g^{\mu\nu}=
	\begin{pmatrix}
		0 & 1 & 0 & 0\\
		1 & (1- \frac{GM}{r})^{2} & 0 & 0\\
		0 & 0 & \frac{1}{r^{2}} & 0\\
		0 & 0 & 0 & \frac{1}{r^{2}sin^{2}(\theta)}\\
	\end{pmatrix}
\end{equation}
and determinant as $\sqrt{-g}=r^{2}f(\Omega)$ where $f$ is some arbitrary function of the angles which won't be relevant to the calculations. The next step that will make our work easier is to expand $\psi$ in spherical harmonics
\begin{equation}
	\psi(v,r, \Omega)= \sum_{l=0}^{\infty}\psi_{l}(v,r)Y_{l}(\Omega)
\end{equation}
where we have removed the index $m$ as it will not be relevant for the work we are doing. Finally we remind ourselves that we can write $\nabla^{2}=\nabla^{2}(v,r)+\nabla^{2}(\Omega)$ and that $r^{2}\nabla^{2}(\Omega)Y_{l}=-l(l+1)Y_{l}.$ Putting this information together into (\ref{KG_equation_curved}) we finally get
\begin{align}
	\nabla^{2}\psi&=\frac{1}{r^{2}}\left[\partial_{v}\left(r^{2}\partial_{r}\psi_{l}\right)	+\partial_{r}\left(r^{2}\partial_{v}\psi_{l}\right) + \partial_{r}\left(\left(1- \frac{GM}{r}\right)^{2}r^{2}\partial_{r}\psi_{l}\right) \right]- \frac{1}{r^{2}}l(l+1)\psi_{l}=0\\
	&=\partial_{v}\left(r^{2}\partial_{r}\psi_{l}\right)	+\partial_{r}\left(r^{2}\partial_{v}\psi_{l}\right) +\partial_{r}(\Delta\partial_{r}\psi_{l})-l(l+1)\psi_{l}=0\\
	&=r^{2}\partial_{v}\partial_{r}\psi_{l}+2r\partial_{v}\psi +r^{2}\partial_{v}\partial_{r}\psi_{l}+\partial_{r}(\Delta\partial_{r}\psi_{l})-l(l+1)\psi_{l}=0\\
	&=2r\partial_{v}\partial_{r}(r\psi_{l})+\partial_{r}(\Delta\partial_{r}\psi_{l})-l(l+1)\psi_{l}=0 \label{KG_rn_pert}
\end{align}
where $\Delta=(r-GM)^{2}$. If we now set $l=0$ and evaluate at the horizon ($r=GM$) then this reduces to
\begin{align}
	GM[\partial_{r}(r\psi_{0})]_{r=GM}&=\text{const}\equiv (GM)^{2}H_{0}\\
	\implies H_{0}&=\frac{1}{GM}\partial_{r}(r\psi_{0})_{r=GM}
\end{align}
which is independent of $v$ and therefore conserved. For general initial conditions $H_{0}$ is non zero which means that $\psi$ and $\partial_{r}\psi$ cannot both tend to zero at late time. Aretakis showed that $\psi$ does decay at late time and therefore $\partial_{r}\psi$ does not decay leading to
\begin{equation}
	 \lim_{v\to\infty}(\partial_{r}\psi_{0})= H_{0}.
\end{equation}
We also want to look the energy density measured by a radially ingoing observer coming from the EM tensor, $T_{rr}$ which would usually decay due at the horizon due to the redshift. From (\ref{eq:energy_mom_tensor_w_lagr}) we get that
\begin{equation}
	T_{\mu\nu}=g_{\mu\nu}\Lagr+\partial_{\mu}\psi\partial_{\nu}\psi
\end{equation}
meaning that $T_{rr}=(\partial_{r}\psi)^{2}$ which does not decay at the horizon, this agrees with the absence of the horizon redshift. 
Next by acting on (\ref{KG_rn_pert}) with $\partial_r$ we can get
\begin{equation}
	2\partial_{v}\partial_{r}(r\psi_{l}) +2r\partial_{v}\partial_{r}^{2}(r\psi_{l}) +2\partial_{r}\psi_{l}=0.
\end{equation}
By setting $l=0$ and evaluating at the horizon we get that the first term is zero leaving us with
\begin{equation}
	\left[GM\partial_{v}\partial_{r}^{2}(r\psi_{0}) +\partial_{r}\psi_{0}\right]_{r=MG}=0.
\end{equation}
In the limit that $v\to\infty$ the second term goes to $H_{0}$ giving us that
\begin{align}
	(\partial_{v}\partial_{r}^{2}(r\psi_{0}))_{r=GM}\to \frac{-H_{0}}{GM}\\
	\implies (\partial_{r}^{2}(r\psi_{0}))_{r=GM} = \frac{-H_{0}v}{GM}
\end{align}
for late time. Expanding out in the limit as $v\to\infty$ we get that 
\begin{align}
	H_{0}+GM(\partial_{r}^{2}\psi_{0})_{r=GM}= \frac{-H_{0}v}{GM}\\
	(\partial_{r}^{2}\psi_{0})_{r=GM}\approx \frac{-H_{0}v}{(GM)^{2}}.
\end{align}
By repeated application of $\partial_{r}$ if can be shown that by keeping only the highest powers of $v$ that $(\partial_{r}^{k}\psi_{0})_{r=M}\propto v^{k-1}$ for late time.
If instead we want to consider the $l\ne 0$ case then by acting with $\partial_{r}^{l}$ on (\ref{KG_rn_pert}) then evaluating at the horizon we get that
\begin{equation}
	2[\partial_{v}\partial_{r}^{l}(r\partial_{r}(r\psi_{l}))]_{r=GM}+(2-l(l+1))(\partial_{r}^{l}\psi_{l})_{r=GM}.
\end{equation}
Aretakis has shown that $\partial_{r}^{k}\psi_{l}$ decays outside and on the horizon for $k\le l$ meaning that we can define 
\begin{equation}
	H_{l}[\psi]=\frac{1}{(GM)^{2}}[\partial_{r}^{l}(r\partial_{r}(r\psi_{l}))]_{r=GM}
\end{equation}
which can be seen is conserved as $[\partial_{v}H_{l}]_{r=M}=0$. It has also been shown that $\partial_{r}^{l+1}\psi_{l}$ generally does not decay at the horizon which gives us that $\partial_{r}^{j}\psi_{l}$ and higher derivatives are proportional to $v^{(j-l-1) }$ for $j\ge l+1$. 
% subsection aretakis_instability (end)
% section intability_of_extremal_rn_black_hole (end)
\section{Charged Black Holes From String Theory} % (fold)
\label{sec:5d_charged_black_holes_from_string_theory}
% section 5d_charged_black_holes_from_string_theory (end)
\subsection{Kaluza Klein Reduction} % (fold)
\label{sub:kaluza_klein_reduction}
Consider the 10 dimensional black hole given by the metric that is derived from string theory and given in the paper by Mohaupt\cite{Mohaupt:2000gc}
\begin{equation}
	ds^{2}=\frac{1}{\sqrt{Z_{1}Z_{2}}}\left(-dt^{2}+dy^{2}+F(dt-dy)^{2}\right)+\sqrt{Z_{1}Z_{2}}dx^{2}_{i}+\sqrt{\frac{Z_{1}}{Z_{2}}}dy_{i}^{2}
\end{equation}
with $i=1,2,3,4$ and
\begin{equation}
	Z_{1,2}=1+\frac{Q_{1,2}}{r^{2}} \;\;\;\text{and}\;\;\; F=\frac{Q_{P}}{r^{2}}
\end{equation}
which are all harmonic functions i.e. solve the Laplace equation in 4D. These \emph{charges} $Q_{1,2,P}$ are defined as $Q=\hat{Q}_{i}c_{i}$ with $\hat{Q}_{i}\in\mathbb{Z}$ and
\begin{equation}
	c_{1}=\frac{4G_{N}^{(5)}R}{\pi\alpha ^{\prime} g_{s}}\;\;\; c_{2}=g_{s}\alpha ^{\prime} \;\;\; c_{P}=\frac{4G_{N}^{(5)}}{\pi R} \label{eq:q_defintions}
\end{equation}
here $R$ is the radius of the $y$ dimension, $g_{s}$ is the string coupling constant and $\alpha^{\prime}$ is related to the tension of the string. The reason we can define them this way is that the integers $\hat{Q}_{1},\hat{Q}_{2}$ and $\hat{Q}_{P}$ literally count the number of $D1$-branes, $D5$-branes and the quantised KK momentum (the momentum in the compact dimension $y$) in the background spacetime. The motivation of working with such a metric is that when the charges are not equal we have a generalisation of the extreme RN metric and this metric has some additional symmetries including 2 supersymmetries and as such we might expect this metric to not show evidence of instabilities at the horizon; when the charges are equal we will see that we just have our usual extreme 5D-RN metric in a shifted coordinate system.

We want to reduce this metric to 5 large dimensions, the first step in doing so is to make it such that the dimensions $y_{i}$ and $y$ are small by requiring them to be closed circles;
\begin{align}
	y_{i}=y_{i}+2\pi l_{s}\\
	y=y+2 \pi R
\end{align}
where $l_{s}$ is the length of the strings and $R \gg l_{s}$.
Next we will remove the smaller dimensions, $y_{i}$; remember the action has a term proportional to $\sqrt{-g_{(10\times10)}}R_{(10)}$ which as the metric is block diagonal can be written as
\begin{equation}
	\sqrt{-g_{(10\times10)}}R_{(10)}=\sqrt{-g_{(6\times6)}}\sqrt{-g_{4\times4}}R_{(6)}=\sqrt{-\tilde{g}_{(6\times6)}}R_{(6)}
\end{equation}
where $\tilde{g}$ is the actual reduced metric and is not equal to $g$. Normally this would mean we have to deal with the trace of the small dimension part of the metric, but if we set $Z_{1}=Z_{2}$ then we have that $\sqrt{-g_{4\times4}}=1$ which means we don't have to concern ourselves with any difficulties when compactify these 4 dimensions and focus on the now 6 dimensional line element. The problem with removing the final compact dimension is the cross term between $y$ and $t$ meaning that we need to do a Kaluza-Klein reduction as just removing the $y$ coordinate as before would result in the remaining line element not satisfying the EFEs. To start with let us remind ourselves that
\begin{equation}
	ds^{2}=g_{\mu\nu}dx^{\mu}dx^{\nu}
\end{equation}
which means for each point in space we have $d$-differentials. Next we introduce a set of $d$ 1-forms $\{\omega^{\alpha}\}$ which we define as
\begin{equation}
	\omega^{a}={E^{a}}_{\mu}(x)dx^{\mu}
\end{equation}
where $E$ is a $d\times d$ matrix and is a point-wise function of the spacetime. If we require that these functions satisfy the relation
\begin{equation}
	{(^{t}E)_{\mu}}^{a}\eta_{ab}{E^{b}}_{\nu}=g_{\mu\nu},
\end{equation}
where $(^{t}E)$ is the transpose of $E$, then these forms are known as the vielbein. The metric can be though of as the square root of it as
\begin{align}
	\det(^{t}E\eta E)=\det(^{t}E)\det(\eta)\det(E)=\pm \det(E)^{2}=\det(g)\\
	\implies \sqrt{|\det(g)|}=|\det(E)|
\end{align}
where we have used the properties of the determinant relating to products and transposes. 
Now we will begin by assuming an ansatz of the form 
\begin{equation}
	{E^{a}}_{\mu}= 
		\begin{pmatrix}
		{E^{\hat{a}}}_{\hat{\mu}} & {E^{\hat{a}}}_{M}\\
		{E^{\alpha}}_{\hat{\mu}} & {E^{\alpha}}_{M} \\
	\end{pmatrix}
\end{equation}
where $\hat{a}$ and $\hat{\mu}$ run over our large dimensions and $\alpha$ and $M$ run over the small dimensions to compactify, in our case they will only ever equal $y$. We also choose to set $ {E^{\hat{a}}}_{M}$ to zero giving us a lower-triangular matrix to work with. It is worth noting that ${E^{\alpha}}_{\hat{\mu}}$ looks like a gauge field, $A_{\hat{\mu}}$, where $\alpha$ would count the number of gauge field; in our case just 1. The next step is to see how we can write the metric in terms of the vielbein; by writing 
\begin{equation}
	{E^{a}}_{\mu}= 
		\begin{pmatrix}
		{E^{\hat{a}}}_{\hat{\mu}} & 0\\
		e^{\sigma}A_{\mu} & e^{\sigma} \\
	\end{pmatrix}
\end{equation}
we can get that 
\begin{equation}
	g_{\mu\nu}= 
		\begin{pmatrix}
		\tilde{g}_{\mu\nu} +A_{\mu}A_{\nu}e^{2\sigma} & A_{\mu}e^{2\sigma}\\
		A_{\mu}e^{2\sigma} & e^{2\sigma} \\
	\end{pmatrix}
\end{equation}
where $\tilde{g}_{\mu\nu}$ is proportional to the $(D-1)$ metric and $\sigma$ is a scalar. Next we will calculate the actual $(D-1)$ metric by looking at how the action changes when this compactification is done. To do this we will make two simplifications; firstly we will assume there is no gauge field as this means we can write the $D$ dimensional metric as
\begin{equation}
		g_{\mu\nu}= 
		\begin{pmatrix}
		{(^{t}E)_{\hat{\mu}}}^{\hat{a}}\eta_{\hat{a}\hat{b}}{E^{\hat{b}}}_{\hat{\nu}} & 0\\
		0 & e^{2\sigma} \\
	\end{pmatrix},
\end{equation}
and secondly we will consider just a scalar term in the action to avoid having to deal with working with the Ricci scalar as the answer will be true for the Ricci scalar term as well. Having done this we can write the action as
\begin{equation}
	\int d^{d}x\sqrt{g_{(d)}}\sqrt{g_{(D-d)}}\;\partial_{\mu}\phi\;\partial_{\nu}\phi \;g^{\mu\nu}_{(d)}. \label{eq:eh_hilber_action_reduction}
\end{equation}
where $d$ is the number of non-compact dimensions. Next, because $\sigma$ is a scalar, we have that
\begin{equation}
	\sqrt{g_{(D-d)}}=e^{\sigma}.
\end{equation}
We will now perform a Weyl transformation to the metric i.e.
\begin{equation}
	g_{\mu\nu}^{(d)}=g_{\mu\nu}^{(5)}e^{\alpha\sigma} \label{eq:weyl_transform}
\end{equation}
with $\alpha$ just being some scaling constant. By plugging this into (\ref{eq:eh_hilber_action_reduction}) we get that 
\begin{equation}
	\int d^{d}x\sqrt{-g_{(5)}}\;e^{\alpha\sigma \frac{d}{2}+\sigma -\alpha\sigma}\;\partial_{\mu}\phi\;\partial_{\nu}\phi \;g^{\mu\nu}_{(5)}
\end{equation}
and then, as we want this action to be independent of the compact scalar $\phi$, we require the exponent to be equal to zero giving us that
\begin{equation}
	\alpha=\frac{1}{1- \frac{d}{2}}.
\end{equation}
Finally we use this value in (\ref{eq:weyl_transform}) to get that
\begin{equation}
	g_{\mu\nu}^{(5)}=g_{\mu\nu}^{(d)}(e^{\sigma})^{\frac{-1}{1-d/2}}=
	g_{\mu\nu}^{(d)}(\sqrt{g_{(D-d)}})^{\frac{1}{d/2-1}}=\tilde{g}_{\mu\nu}(e^{2\sigma})^{\frac{1}{d-2}}.
\end{equation}
Putting this together we get the equation for our reduced metric as
\begin{equation}
	g_{\mu\nu}^{(D-1)}=\left(g_{\mu\nu}^{D}-A_{\mu}A_{\nu}e^{2\sigma}\right)(e^2{\sigma})^{\frac{1}{d-2}}
\end{equation}
Going back to our line element we see that
\begin{align}
	g_{tt}&= -\frac{1-F}{Z_{1}}\\
	g_{ty}&=g_{yt}=A_{\mu}e^{2\sigma}=\frac{-F}{Z_{1}}\\
	g_{yy}&=e^{2\sigma}=\frac{1+F}{Z_{1}}\\
	g_{x_{i}x_{i}}&=Z_{1}
\end{align}
giving us that our new metric has elements
\begin{align}
	g_{tt}^{(5)}&=\left(	-\frac{1-F}{Z_{1}} -\frac{F^{2}}{Z_{1}^{2}}\frac{Z_{1}}{1+F}\right){\left(\frac{1+F}{Z_{1}}\right)}^{\frac{1}{3}}\\
	&=\frac{1}{Z_{1}^{4/3}}\frac{-1}{(1+F)^{2/3}}\\
	g_{x_{i}x_{i}}^{(5)}&=Z_{1}{\left(\frac{1+F}{Z_{1}}\right)}^{\frac{1}{3}}\\
	&=(1+F)^{1/3}Z_{1}^{2/3}.
\end{align}
The final thing we do is to now specialise to our RN case where there is only 1 charge, i.e. $Q_{P}=Q_{1} \implies Z_{1}=(1+F)\equiv Z$, meaning that we can write our reduced metric as 
\begin{equation}
	ds^{2}_{reduced}=-Z^{-2}dt^{2}+Z dx_{i}^{2}.
\end{equation}
% subsection kaluza_klein_reduction (end)
\subsection{Entropy of the String Metric} % (fold)
\label{sub:entropy_of_the_string_metric}
To calculate the entropy of the metric we will need to calculate the surface area of the black hole in this space time. The infinitesimal surface area of a sphere in 5 dimensional hyper-spherical coordinates ($r,\theta,\phi,\xi$) in a curved space time is given by 
\begin{equation}
	A=r^{3}\int\sqrt{\det(g_{\Omega\Omega})d\Omega^{2}_{3}}.
\end{equation}
Had we not equated any of the $Z$ functions the angular metric element would be 
\begin{equation}
	\left(Z_{1}Z_{2}F\right)^{1/3}d\Omega^{2}_{3}
\end{equation}
which means that we have
\begin{equation}
	A=r^{3}\sqrt{Z_{1}Z_{2}F}\int_{0}^{2\pi}\int_{0}^{\pi}\int_{0}^{\pi} \sin(\theta)\sin(\phi)\; d\theta d\phi d\xi=2\pi^{2}r^{3}\sqrt{Z_{1}Z_{2}F}.
\end{equation}
We want to evaluate this at the horizon which in these coordinates is at $r=0$ so if we expand the terms in the square root and take the limit towards the horizon we get that
\begin{equation}
	A=2\pi^{2}\lim_{r\to 0}\left(r^{3}\sqrt{\frac{Q_{1}Q_{2}Q_{P}}{r^{6}}}\right)=\sqrt{Q_{1}Q_{2}Q_{P}}.
\end{equation}
By plugging this into our equation for the entropy (\ref{eq:entropy}) and by using the definition of the $Q$'s given earlier in (\ref{eq:q_defintions}) we can see that the BH entropy for this 5D black hole is given by
\begin{equation}
	S=2\pi\sqrt{\hat{Q}_{1}\hat{Q}_{2}\hat{Q}_{P}}
\end{equation}
which is entirely dependent on the integer numbers defined above. It is possible to count the possible number of states, as was done by Mohaupt, and you come to the same result; this is important as it shows the the BH entropy formula gives the same value as a counting of states which is classically the way entropy is calculated.
% subsection entropy_of_the_string_metric (end)
\subsection{Comparison of Metrics} % (fold)
\label{sub:comparison_of_metrics}
To make sure that this is actually the RN black hole in 5 dimensions we need to compare the metrics and show that there is a coordinate transform from one to the other. First we need to set the coordinates $x_{i}$ as $r$ and the angular coordinates $\Omega$. The string metric is in a frame such that the horizon is at $r=0$ so we should expect the that the transformation would be 
\begin{equation}
	r^{2}=r^{2}_{s}+GQ^{2}
\end{equation}
where $r_{s}$ is the string metric radius, $r$ is the radius in the metric as given by (\ref{eq:RN_higher_dim_metric}) and we will also set it such that $Q_{1}=GQ$. We can verify this for each dimension starting with time 
\begin{align}
	g_{tt}^{s}&=-\left(1+\frac{GQ}{r_{s}^{2}}\right)^{-2}=-\left(\frac{r_{s}^{2}+GQ}{r_{s}^{2}}\right)^{-2}\\
	&=-\left(\frac{r^{2}}{r^{2}-GQ^{2}}\right)^{-2}=-\left(1-\frac{GQ^{2}}{r^{2}}\right)^{2}=g_{tt}.
\end{align}
The next one to compare will be the angular elements as its the easiest of the two remaining
\begin{align}
	g_{\Omega\Omega}^{s}&=Zr^{2}_{s}=\left(1+ \frac{GQ^{2}}{r^{2}_{s}}\right)r^{2}_{s}=r^{2}=g_{\Omega\Omega}.
\end{align}
The final one is the radial coordinate so we first need to write that
\begin{equation}
	dr_{s}^{2}=\frac{r^{2}}{r^{2}-GQ^{2}}
\end{equation}
which gives us that
\begin{equation}
	g_{rr}^{s}dr^{2}_{s}=\left(1+\frac{GQ^{2}}{r^{2}_{s}}\right)dr_{s}^{2}=\left(\frac{r^{2}}{r^{2}-GQ^{2}}\right)^{2}dr^{2}=\left(\frac{1}{1- \frac{GQ^{2}}{r^{2}}}\right)^{2}dr^{2}=g_{rr}dr^{2}.
\end{equation}
As we can see all the elements of the metric agree perfectly with that coordinate transform and therefore we have proved that the metric we derived from string theory is in fact the same as the 5 dimensional RN metric we derived before. 
% subsection comparison_of_metrics (end)
\subsection{Instability in 5D} % (fold)
\label{sub:instability_in_5d}
To see if there exists an instability in 5D we first need to know what the metric looks like in the EF coordinates. We approach the problem in the exact same way by requiring that 
\begin{equation}
	\frac{dr^{*}}{dr}=\left(1- \frac{GM}{r^{2}}\right)^{-2}
\end{equation}
which gives us that 
\begin{equation}
	r^{*}=r+\frac{GMr}{2GM-2r^{2}}-\frac{3}{2}\sqrt{GM}\arctanh\left(\frac{r}{\sqrt{GM}}\right)
\end{equation}
and then using the ingoing EF coordinate defined as $v=t+r^{*}$ we can then just write the line element in the form
\begin{equation}
	ds^{2}=-\left(1- \frac{GM}{r^{2}}\right)^{2}dv^{2}+2dvdr+r^{2}d\Omega^{2}_{3}
\end{equation}
and our inverse metric as 
\begin{equation}
	g^{\mu\nu}=
	\begin{pmatrix}
		0 & 1 & 0 & 0 & 0\\
		1 & (1- \frac{GM}{r^{2}})^{2} & 0 & 0& 0\\
		0 & 0 & \frac{1}{r^{2}} & 0& 0\\
		0 & 0 & 0 & \frac{1}{r^{2}sin^{2}(\theta)}& 0\\
		0 & 0 & 0 & 0 & \frac{1}{r^{2}sin^{2}(\theta)\cos^{2}(\phi)}\\
	\end{pmatrix}.
\end{equation}
Finally, we now have that $\sqrt{-g}=r^{3}f(\Omega)$ as we now have three angular elements in the metric. Now using (\ref{KG_equation_curved}) we get that
\begin{equation}
	\nabla^{2}\psi=\frac{1}{r^{3}}\left[\partial_{v}\left(r^{3}\partial_{r}\psi_{l}\right)	+\partial_{r}\left(r^{3}\partial_{v}\psi_{l}\right) + \partial_{r}\left(\left(1- \frac{GM}{r^{2}}\right)^{2}r^{3}\partial_{r}\psi_{l}\right) \right]- \frac{1}{r^{3}}l(l+2)\psi_{l}=0. \label{eq:5d_rn_kg}
\end{equation}
The reason the last term has changed to be $l(l+2)$ is as that additive term is related to the number of angular dimensions. By setting $l=0$ and evaluating at the horizon $(r^{2}=GM)$ we find that only the first two terms survive and we see that
\begin{equation}
	\partial_{v}\left(3GM\psi_{0}+2(GM)^{3/2}\partial_{r}\psi_{0}\right)_{r^{2}=GM}=0
\end{equation}
which means that the terms in the bracket must be a conserved quantity. If we choose to define this such that
\begin{equation}
	(3GM\psi_{0}+2(GM)^{3/2}\partial_{r}\psi_{0})_{r^{2}=GM}\equiv 2 (GM)^{3/2}H_{0}^{\prime} 
\end{equation}
then by taking the limit of late time i.e. $v\to \infty$ and if we make the assumption that $\psi$ still decays then we get that
\begin{equation}
	\lim_{v\to\infty}(\partial_{r}\psi_{0})_{r^{2}=GM}=H_{0}^{\prime} 
\end{equation}
just as we had in 4 dimensions. We also see that the energy density $T_{rr}=(\partial_{r}\psi)^{2}$ still does not decay again due to the lack of gravitational redshift. The next step is to once again act $\partial_{r}$ on (\ref{eq:5d_rn_kg}) and setting $l=0$ to get that
\begin{equation}
	6r^{2}\partial_{v}\partial_{r}\psi_{0}+2r^{3}\partial_{v}\partial_{r}^{2}\psi_{0}+6r\partial_{v}\psi_{0}+\partial_{r}^{2}\left(\left(1- \frac{GM}{r^{2}}\right)^{2}r^{3}\partial_{r}\psi_{0}\right)=0.
\end{equation}
We can see that there will only be one surviving contribution from the last term when evaluating at the horizon which means after simplifying our expression we have
\begin{equation}
	\left[\partial_{v}(3\psi_{0}+2\sqrt{GM}\partial_{r}\psi_{0})+\sqrt{GM}\partial_{v}\partial_{r}\psi_{0}+GM\partial_{v}\partial_{r}^{2}\psi_{0}+4(\partial_{r}\psi)\right]_{r^{2}=GM}=0.
\end{equation}
The reason we have written it in this form is so we can easily see that the first term is 0 and that when we take the late time limit we have that
\begin{equation}
	\partial_{v}\left[\sqrt{GM}\partial_{r}\psi_{0}+GM\partial_{r}^{2}\psi_{0}\right]+4H_{0}^{\prime} =0 \;\;\;\text{as}\;\; v\to\infty
\end{equation}
which means that the term in the brackets must be equal to $-4H_{0}^{\prime}v$. Seeing that the first term is also proportional to $H_{0}^{\prime}$ and remembering we are in the late time limit we finally see that
\begin{equation}
	(\partial_{r}^{2}\psi_{0})_{r^{2}=GM}\approx \frac{-4H_{0}^{\prime}v}{GM}
\end{equation}
where, as in the four dimensional case, we get that the second derivative of the field blows up at late time meaning there is an instability for this black hole.
% subsection instability_in_5d (end)
\section{Discussion} % (fold)
\label{sec:conclusion}
In this work we have derived the metrics as well as thermodynamic properties of both the Schwarzschild and Reissner-Nordstr\"om black holes in $D\ge4$ dimensions; we have also shown that the extremal RN black hole exhibits an instability when perturbed by a massless scalar field. This is cause for concern, we expect the extremal solution to be stable; even more so in 5 dimensions where the black hole can be embedded in string theory. What we have done is cover the groundwork of a larger area of study and we have left many unanswered questions; what is the final state of the black hole, do there exist non-generic perturbations that do not lead to an instability, and does this result extend to other extremal black holes regardless of dimension?

The first question, that of the final state, has been answered by numerical calculations done by Murata et al\cite{Murata:2013daa} and they found that for a general perturbation the late time result is that of a non-extreme RN black hole. However, they also found that for a specific tuning of the perturbation it is possible to have a time-dependent extremal RN black hole at a late time; one that outside the horizon acts as an extremal RN but on the horizon does not. The idea of a non-extreme black hole settling down to an extreme one may considered to be a violation of the third law of thermodynamics but by using a definition of the third law as given by Israel\cite{PhysRevLett.57.397} it is possible to have this situation without violating the third law.

The final question of whether this applies to other extremal black holes was also answered by Murata\cite{Murata:2012ct} who studied gravitational (and electromagnetic) perturbations on extremal black hole solutions by a more mathematical approach. They show that many of the extreme higher dimensional black holes exhibit the instability, and specifically all of such solutions in 5 dimensions.

Finally, we can briefly discuss one question that, as far as we can tell, has not yet been answered; what happens if we perturb the 3-charge string theory metric with a supersymmetric scalar? One might expect that by using a supersymmetric scalar that the supersymmetries of the system might be preserved leading to no instability on the horizon. It may very well be that this perturbation is analogous to the tuned parameter found by Murata et al that lead to the time-dependent extremal RN result.

It is clear that a lot of work has been done on the topic of the instabilities, both theoretically for generic perturbations, and numerically for both generic and tuned cases, and that we are closer to understanding why and how such instabilities occur. However there is still the need for theoretical study of particular perturbations to the black hole, be they supersymmetric or otherwise, in an attempt to find exact final-state solutions to broaden our understanding of this unexpected result.
% section conclusion (end)
\nocite{*}
\bibliographystyle{ieeetr}
\bibliography{black_hole_references} 
\end{document}

### 2020-07-18 Tex/MSc Data Science Project.tex
Add results data. Add training and eval notebooks

@@
 \setcounter{page}{1}
-\section{Introduction} % (fold)
+\section{Introduction} 
 \label{sec:intro}
 
+Introduce the project, describe the goal and the idea behind why this is important. Explain the structure of this work,
+
+\subsection{Related Works} 
+\label{sub:Related Works} 
+Discuss the work of Iwana as the main source, reference their methods and results but keep much for later where relevant for comparison with my work.
+% subsection Related Works end 
 % section CNN Theory and Architectures
+\section{Theory} 
+\label{sec:Theory} 
+\subsection{Components of Convolutional Neural Networks} 
+\label{sub:Components_of_Convolutional_Neural_Networks} 
+Describe the maths and the theory behind the various components of CNNs to the best of my ability, referring to the source papers for more complex combined units where sensible. 
+% subsection Components_of_Convolutional_Neural_Networks end 
+\subsection{Transfer Learning} 
+\label{sub:Transfer_Learning} 
+Describe the theory and maths behind transfer learning to the reader.
+% subsection Transfer_Learning end
 
+\subsection{Activation Maximisation} 
+\label{sub:Activation_Maximisation} 
+Describe the theory and/or maths of activation maximisation and the methods used within the package we use the regularisations applied. 
+% subsection Activation_Maximisation end 
+% section Theory end 
 \section{Data collection and pre-processing} 
 \label{sec:Data_collection_and_pre-processing} 
-\subsection{Manual review} 
-\label{sub:Manual_review} 
+\subsection{Image Download and Manual Review} 
+\label{sub:Image Download and Manual Review} 
+Detail the process of collecting the data and manually reviewing those chosen for use within the train/test set, as well as the validation split.
+
 Rule of thumb: single front facing book cover, book cover did not contain pictures of multiple books, and was not just an arrangement of book covers.
 Covers designed specifically for a boxset were allowed.
-% section Manual_review end 
+% subsection Image Download and Manual Review end 
+\subsection{Image Pre-processing Technique} 
+\label{sub:Image_Pre-processing_Technique} 
+Detail the approaches taken when considering how to pre-process the images and their results, including what the final choice was.
+% section Image_Pre-processing_Technique end 
 % section Data_Collection_and_pre-processing end 
-% Section Data collection and pre-processing and exploratory analysis
-    % include size heatmap
-    % include average image per class
+\section{Model Training} 
+\label{sec:Model_Training} 
+\subsection{Configuration} 
+\label{sub:Configuration} 
+Detail the models used, the hyperparameters set as well as the optimiser and the loss functions chosen. Denote the pre-trained sources and the reason for choosing these models.
+% subsection Configuration end 
+\subsection{Training Performance} 
+\label{sub:Training_Performance} 
+Detail and visualise the performance of the models through training; compare length of training to that of Iwana.
+% subsection Training_Performance end
 
-% Section Model Training and Results
+% section Model_Training end
 
-% Section Feature Visualisation
+\section{Evaluation} 
+\label{sec:Evaluation_and_Further_Exploration} 
+\subsection{Results} 
+\label{sub:Results} 
+Present and discuss the basic results for each model, including a comparison with Iwana with the caveat of a slightly different dataset and different pre-processing.
 
-% Section Discussion and future work
+% section Results end 
+\subsection{Further Analysis} 
+\label{sub:Further_Analysis} 
+Consider further, in depth, analysis on the best performing model, including some of the insight provided by Iwana.
 
-% Section Conclusion
+Idea: compare class average colour by predicted and actual and compare to accuracy?
+% section Further_Analysis end 
+\subsection{Feature Visualisation} 
+\label{sub:Feature_Visualisation} 
+Detail the approach taken to try and visualise the ideal input to trigger the most accurate classes in an attempt to understand what feature the model is picking up on. Show output and discuss impact.
+% section Feature_Visualisation end
+
+\subsection{Discussion} 
+\label{sub:Discussion} 
+Discussion around the reasons we have seen some of these results if not already covered in the previous subsections. 
+% subsection Discussion end 
+% section Evaluation_and_Further_Exploration end
+
+\section{Conclusion} 
+\label{sec:Conclusion} 
+Discuss the project as a whole, the impact of the results when compared to the literature, and what future work there could be off the back of this.
+% section Conclusion end
 
@@
 \bibliography{BBK_MSc_Project}
+
+\appendix
+
+\section{CNN Architectures} 
+\label{sec:CNN_Architectures} 
+Detail the exact configurations of these architectures
+\subsection{MobileNetV2} 
+\label{sub:MobileNetV2}
+
+% subsection MobileNetV2 end  
+\subsection{Inception-ResnetV2} 
+\label{sub:Inception-ResnetV2}
+
+% subsection Inception-ResnetV2 end 
+\subsection{ResNeXt50} 
+\label{sub:ResNeXt50}
+
+% subsection ResNeXt50 end 
+% section CNN_Architectures end
+
+\section{Technology} 
+\label{sec:Technology} 
+List of software and packages/libraries including their versions.
+% section Technology end 
 \end{document}

### 2020-07-23 Tex/MSc Data Science Project.tex
Add pre-processing to writeup

@@
 \label{sub:Image Download and Manual Review} 
-Detail the process of collecting the data and manually reviewing those chosen for use within the train/test set, as well as the validation split.
+The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}.
+The original csv containing all image URLs was split into files of no more than 20,000 records each to allow to the downloads to be run over multiple sessions do to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway.
+Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e.
+the genre classification.
 
-Rule of thumb: single front facing book cover, book cover did not contain pictures of multiple books, and was not just an arrangement of book covers.
-Covers designed specifically for a boxset were allowed.
+Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.
+At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, and example of such an image can be seen in \textcolor{red}{TODO: add image}\cref{fig:TBC}.
+The terms used in a regex search of the titles were \emph{"boxed set", "boxset", "box set", "anthology", "bundle", "\#-book", "\# book"}.
+264 records within the bookcover30 dataset matched at least one of these terms and were extracted into a new folder for manual review.
+The process was somewhat subjective but as a rule of thumb we rejected any image that was not a single front facing cover, that contained pictures of multiple books and was not an arrangement of covers; single covers that seemed to be specifically designed for a boxset were considered acceptable.
+In total 94 records needed to be removed due to these rules.
+To preserve the class balance of the train and test datasets, new records were chosen from the book32 dataset by making like-for-like swaps of the same class; any new covers were also reviewed to ensure they were not boxsets.
+No verification of the class assignment to images was performed so any errors that existed in the original data would have persisted and we continue to use the single class chosen by the original authors in the case of multi-class books.
+
+We chose to split the training data out into training and validation data using a 90/10 random split per class to ensure class balance in all of the training, validation, and test sets.
+In total the training set compromised of 46,170 records, the validation set 5130 records, and
+the test set 5700 records; with 1,539, 171, and 190 records per class respectively.
+
 % subsection Image Download and Manual Review end 
@@
 \label{sub:Image_Pre-processing_Technique}
-Detail the approaches taken when considering how to pre-process the images and their results, including what the final choice was.
+
+We next used 3 different image preparation methods on the datasets to attempt to identify the method that would lead to the best accuracy within our models.
+At this stage all methods output an image 299x299 pixels as this was the largest size we would need for any future models.
+By reducing the image size twice, first now and then again when using a specific model, we were potentially losing some information that would lead to a less accurate model; however we chose to do this to speed up training and testing by having to load much smaller data in compared to the original full sized images.
+Before pre-processing was done we collected the initial shape and average RGB values of each image for further analysis later.
+
+The first pre-processing method (\emph{NoPrep}) was to simply rescale the image directly down to a 299x299 shape using inter-cubic interpolation \textcolor{red}{TODO: explain this mathematically}.
+This was chosen at the time as a common all-purpose interpolation method, however after later research this method is often chosen to enlarging images, not shrinking them so the potential to use a better interpolation method could be used in future work.
+
+The second method (\emph{Padded}) was to first pad the images with a symmetric pure white border on either the left/right or top/bottom depending on if the image was portrait or landscape (no padding was done if the image was already a perfect square).
+These images were then downscaled to 299x299 using the same method as above.
+The final method (\emph{Cropped}) was to crop the image to a perfect square before the same downscaling method was used.
+
+Once the images had been processed we used a MobileNetV2 model on each processed training and validation dataset for 30 epochs with an early stopping patience of 15 (i.e.
+if no improvement was seen in validation accuracy over 15 epochs then we did not continue) and the best model for validation accuracy was kept.
+The model was trained using the default TensorFlow 2 settings of the adam optimiser with a loss function of sparse categorical cross entropy.
+More details of the set up are detailed in \cref{sec:Model_Training} as other than the epoch and early stopping the same configuration was used for the main training process.
+The test results of this model trained on each of the 3 datasets is presented in \cref{tab:preprocres} and the second method, padding the image with a white border, achieved the best result on the test data of 15.81\% accuracy.
+
+\begin{table}[]
+	\centering
+	\begin{tabular}{|l|r|r|}
+	\hline
+	\textbf{Pre-processing Method} & \multicolumn{1}{l|}{\textbf{Best epoch}} & \multicolumn{1}{l|}{\textbf{Test Accuracy}} \\ \hline
+	NoPrep                         & 8                                        & 14.39\%                                     \\ \hline
+	Padded                         & 6                                        & 15.81\%                                     \\ \hline
+	Cropped                        & 5                                        & 12.88\%                                     \\ \hline
+	\end{tabular}
+	\caption{Results of short training on MobileNetV2 on different image pre-processing methods}
+	\label{tab:preprocres}
+\end{table}
+
+It is important to note at this point that just because this method achieved the best result for this model in a short number of epochs is no guarantee that this would be the method that achieves the best results in other models or over longer epochs.
+We use this as a rough benchmark as given the constraints on computing power available to us we were only able to train each model in \cref{sec:Model_Training} once for a reasonable number of epochs, so a decision had to be made.
+It is also possible that hyperparameter tuning for each of these datasets may have led to different outcomes, but again due to the constraint on resources we decided it best to use domain accepted defaults and as such we progress with the padded dataset.
+
 % section Image_Pre-processing_Technique end 
@@
 Discussion around the reasons we have seen some of these results if not already covered in the previous subsections.
+
+Future work:
+use different interpolation method
 % subsection Discussion end 

### 2020-08-11 Tex/MSc Data Science Project.tex
R analysis and further writing

@@
 We chose to split the training data out into training and validation data using a 90/10 random split per class to ensure class balance in all of the training, validation, and test sets.
-In total the training set compromised of 46,170 records, the validation set 5130 records, and
-the test set 5700 records; with 1,539, 171, and 190 records per class respectively.
+In total the training set compromised of 46,170 records, the validation set 5130 records, and the test set 5700 records; with 1,539, 171, and 190 records per class respectively.
 
@@
 \label{sub:Configuration} 
-Detail the models used, the hyperparameters set as well as the optimiser and the loss functions chosen. Denote the pre-trained sources and the reason for choosing these models.
+All model architectures consisted of their respective base CNN structure as detailed in \cref{sec:CNN_Architectures} followed by a global average pooling 2D layer to reduce the dimensionality of our tensor and to ensure a flat 1D vector which is fully connected to a 30 neuron output layer with a softmax activation.
+All models were also trained using the same setup in terms of hyperparameters, optimisers and loss functions; the loss function was sparse categorical cross entropy (sparse due to the usage of a non one-hot encoded target vector), the optimiser was the Adam optimiser \cite{Kingma2015}.
+The optimiser was used with default hyperparameters as defined within the TensorFlow implementation, the most important being the learning rate with was left at the initial value of 0.001.
+The reason for the usage of the default hyperparameters and not using a more advanced custom optimiser, such as WAME \cite{Mosca2017}, was due to the constraints on access time to computing power.
+As is the training already had to be completed across multiple sessions and doing any tuning would have resulted in too large a time required to produce meaningful results; Adam produces decent results with the defaults and so they were used in this work.
+
+The models themselves, MobileNetV2, Inception-ResnetV2, and ResNeXt50 were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}.
+MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResnetV2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures.
+These models, combined with the original work done by Iwana et al, span a range of complexity and training speed allowing for a good comparison between performance.
+
+The models were training once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point.
+No cross-validation was done due to resource limitations.
 % subsection Configuration end 
@@
 \label{sub:Training_Performance} 
-Detail and visualise the performance of the models through training; compare length of training to that of Iwana.
+The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most.
+
+The performance of the models throughout training can be seen in \cref{fig:train_perf} where whilst training accuracy continues to increase over time, the validation accuracy doesn't see as much consistent movement which suggests the model is starting to overfit to the training data as opposed to improving the identification of intrinsic underlying features within the images themselves.
+Iwana does not report their training accuracy so we are not able to compare the performance of our training with theirs.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.5]{training_results.png}
+	\caption{Performance of models on training and validation data throughout training. Note that models that end short stopped training due to early stopping.}
+	\label{fig:train_perf}
+\end{figure}
+
+Iwana et al trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs.
+It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.
 % subsection Training_Performance end

### 2020-08-11 Files and scripts/color_dist.R

@@ -1,0 +1,144 @@
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+c_names <- c('Arts & Photography',
+             'Biographies & Memoirs',
+             'Business & Money',
+             'Calendars',
+             "Children's Books",
+             'Comics & Graphic Novels',
+             'Computers & Technology',
+             'Cookbooks, Food & Wine',
+             'Crafts, Hobbies & Home',
+             'Christian Books & Bibles',
+             'Engineering & Transportation',
+             'Health, Fitness & Dieting',
+             'History',
+             'Humor & Entertainment',
+             'Law',
+             'Literature & Fiction',
+             'Medical Books',
+             'Mystery, Thriller & Suspense',
+             'Parenting & Relationships',
+             'Politics & Social Sciences',
+             'Reference',
+             'Religion & Spirituality',
+             'Romance',
+             'Science & Math',
+             'Science Fiction & Fantasy',
+             'Self-Help',
+             'Sports & Outdoors',
+             'Teen & Young Adult',
+             'Test Preparation',
+             'Travel',
+             'Filename')
+
+classes <- c("0" = 'Arts & Photography',
+             "1" = 'Biographies & Memoirs',
+             "2" = 'Business & Money',
+             "3" = 'Calendars',
+             "4" = "Children's Books",
+             "5" = 'Comics & Graphic Novels',
+             "6" = 'Computers & Technology',
+             "7" = 'Cookbooks, Food & Wine',
+             "8" = 'Crafts, Hobbies & Home',
+             "9" = 'Christian Books & Bibles',
+             "10" = 'Engineering & Transportation',
+             "11" = 'Health, Fitness & Dieting',
+             "12" = 'History',
+             "13" = 'Humor & Entertainment',
+             "14" = 'Law',
+             "15" = 'Literature & Fiction',
+             "16" = 'Medical Books',
+             "17" = 'Mystery, Thriller & Suspense',
+             "18" = 'Parenting & Relationships',
+             "19" = 'Politics & Social Sciences',
+             "20" = 'Reference',
+             "21" = 'Religion & Spirituality',
+             "22" = 'Romance',
[…]

### 2020-08-11 Files and scripts/colour analysis.R

@@ -1,0 +1,44 @@
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+train_data <- read.csv("Files and scripts/dims_and_colours_to_analyse_train.csv")
+valid_data <- read.csv("Files and scripts/dims_and_colours_to_analyse_valid.csv")
+test_data <- read.csv("Files and scripts/dims_and_colours_to_analyse_test.csv")
+
+comb_data <- train_data %>% 
+  rbind(valid_data, test_data)
+
+pivot_data <- mutate(Aspect_Ratio = Height/Width) %>%
+  pivot_longer(Height:Aspect_Ratio, names_to = 'measure')
+
+pivot_data %>% 
+  filter(measure != 'Height', 
+         ((measure == 'Aspect_Ratio' & value < 10) | measure != 'Aspect_Ratio'),
+         Dataset == 'Train') %>%
+
+  ggplot(aes(x = reorder(Category, desc(Category)),  group = Category, fill = measure)) + 
+  geom_boxplot(aes(y= value)) +
+  facet_wrap(~measure, scales = 'free_x') +
+  coord_flip() +
+  theme_bw() +
+  labs(y = 'Measure Value', x = 'Category') +
+  scale_fill_manual(values = c('white', 'blue', 'green', 'red', 'white')) +
+  theme(legend.position = 'none')
+
+
+comb_data %>%
+  group_by(Category) %>%
+  mutate(Avg_Color = rgb(mean(Avg_Red)/255, mean(Avg_Green)/255, mean(Avg_Blue)/255)) %>% 
+  ungroup() %>%
+  select(Category, Avg_Color) %>% 
+  distinct() %>%
+  ggplot(aes(x = 1, y = reorder(Category, desc(Category)), fill = Avg_Color)) + 
+  geom_col() +
+  scale_fill_identity() +
+  theme_minimal() + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank()) +
+  scale_x_continuous(expand = c(0, 0)) +
+  labs(y = 'Category',
+       title = 'Average Cover Colour by Category')

### 2020-08-11 Files and scripts/combined_file.R

@@ -1,0 +1,377 @@
+# Libraries ---------------------------------------------------------------
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(ggforce)
+
+
+# Functions ---------------------------------------------------------------
+get_avg_pred_cols <- function(df){
+  df %>%
+    pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') %>%
+    group_by(Filename) %>%
+    filter(Prediction_Perc == max(Prediction_Perc)) %>%
+    left_join(test_data) %>%
+    group_by(Predicted_Class) %>% 
+    summarise(Pred_Avg_red = mean(Avg_Red)/255, 
+              Pred_Avg_green = mean(Avg_Green)/255,
+              Pred_Avg_blue = mean(Avg_Blue)/255) %>%
+    return()
+}
+
+get_class_confidence <- function(df){
+  df %>%
+    pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') %>%
+    group_by(Filename) %>%
+    arrange(desc(Prediction_Perc)) %>%
+    slice(1:2) %>%
+    summarise(uncertainty = min(Prediction_Perc)/max(Prediction_Perc)) %>%
+    left_join(test_data) %>%
+    group_by(Category) %>% 
+    summarise(avg_uncertainty = mean(uncertainty)) %>%
+    return()
+}
+
+get_class_overlap <- function(df){
+  temp_df <- df %>%
+    pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') 
+  
+  left_join(temp_df, temp_df, by = c('Filename' = 'Filename')) %>%
+    mutate(class_overlap = Prediction_Perc.x/Prediction_Perc.y,
+           class_overlap = if_else(class_overlap > 1, 1/class_overlap, class_overlap)) %>%
+    filter(Predicted_Class.x >= Predicted_Class.y)  %>% # deal with double on all but equal joins
+    
+    group_by(Predicted_Class.x, Predicted_Class.y) %>% 
+    summarise(avg_overlap = mean(class_overlap)) %>%
+    return()
+}
+
+
+# Reference ---------------------------------------------------------------
+
+c_names <- c('Arts & Photography',
+             'Biographies & Memoirs',
+             'Business & Money',
+             'Calendars',
+             "Children's Books",
[…]

### 2020-08-11 Files and scripts/core_training_analysis.R

@@ -1,0 +1,41 @@
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(ggforce)
+
+mob_training_data <- read.csv("Files and scripts/mobilenet_model_history_log.csv")
+inc_training_data <- read.csv("Files and scripts/inception_resenetv2_model_history_log.csv")
+res_training_data <- read.csv("Files and scripts/resnext_model_history_log.csv")
+
+comb_data <- mob_training_data %>% mutate(Model = 'MobileNetV2') %>%
+  rbind(inc_training_data %>% mutate(Model = 'Inception-ResnetV2')) %>%
+  rbind(res_training_data %>% mutate(Model = 'ResNeXt50'))
+
+
+data_to_plot <- comb_data %>% 
+  rename(Train_accuracy = accuracy, 
+         Train_loss = loss, 
+         Train_sparse_top_3_categorical_accuracy = sparse_top_k_categorical_accuracy,
+         val_sparse_top_3_categorical_accuracy = val_sparse_top_k_categorical_accuracy) %>%
+  pivot_longer(Train_accuracy:val_sparse_top_3_categorical_accuracy, names_to = "temp") %>%
+  separate(temp, c("data_type", "measure"), "_", extra = 'merge') %>%
+  mutate(data_type = if_else(data_type == 'Train', 'Training', 'Validation'),
+         measure = case_when(measure == 'accuracy' ~ 'Accuracy',
+                             measure == 'loss' ~ 'Categorical Cross Entropy',
+                             measure == 'sparse_top_3_categorical_accuracy' ~ 'Top 3 Accuracy')) %>%
+  group_by(Model, data_type, measure) %>%
+  mutate(best_val_acc = if_else(data_type == 'Validation' & measure == 'Accuracy' & value == max(value), 1, 0)) %>%
+  ungroup()
+
+data_to_plot %>%
+  ggplot(aes(x = epoch, y= value, group = Model)) + 
+  geom_line(aes(col = Model), size = 1.25) + 
+  geom_mark_ellipse(aes(filter = best_val_acc == 1), expand = unit(2, "mm")) +
+  facet_grid(measure~data_type, scales = 'free_y') +
+  theme_bw() +
+  labs(x = 'Epoch',
+       y = 'Measure Value',
+       caption = 'Circled validation accuracies are the best for each model and were used as the final model for testing') +
+  scale_color_brewer(palette="Set1") +
+  theme(legend.position = 'bottom',
+        plot.caption = element_text(face = 'italic'))

### 2020-08-11 Files and scripts/overlap and results spread.R

@@ -1,0 +1,191 @@
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+c_names <- c('Arts & Photography',
+             'Biographies & Memoirs',
+             'Business & Money',
+             'Calendars',
+             "Children's Books",
+             'Comics & Graphic Novels',
+             'Computers & Technology',
+             'Cookbooks, Food & Wine',
+             'Crafts, Hobbies & Home',
+             'Christian Books & Bibles',
+             'Engineering & Transportation',
+             'Health, Fitness & Dieting',
+             'History',
+             'Humor & Entertainment',
+             'Law',
+             'Literature & Fiction',
+             'Medical Books',
+             'Mystery, Thriller & Suspense',
+             'Parenting & Relationships',
+             'Politics & Social Sciences',
+             'Reference',
+             'Religion & Spirituality',
+             'Romance',
+             'Science & Math',
+             'Science Fiction & Fantasy',
+             'Self-Help',
+             'Sports & Outdoors',
+             'Teen & Young Adult',
+             'Test Preparation',
+             'Travel',
+             'Filename')
+
+classes <- c("0" = 'Arts & Photography',
+             "1" = 'Biographies & Memoirs',
+             "2" = 'Business & Money',
+             "3" = 'Calendars',
+             "4" = "Children's Books",
+             "5" = 'Comics & Graphic Novels',
+             "6" = 'Computers & Technology',
+             "7" = 'Cookbooks, Food & Wine',
+             "8" = 'Crafts, Hobbies & Home',
+             "9" = 'Christian Books & Bibles',
+             "10" = 'Engineering & Transportation',
+             "11" = 'Health, Fitness & Dieting',
+             "12" = 'History',
+             "13" = 'Humor & Entertainment',
+             "14" = 'Law',
+             "15" = 'Literature & Fiction',
+             "16" = 'Medical Books',
+             "17" = 'Mystery, Thriller & Suspense',
+             "18" = 'Parenting & Relationships',
+             "19" = 'Politics & Social Sciences',
+             "20" = 'Reference',
[…]

### 2020-08-11 Files and scripts/prep_training_analysis.R

@@ -1,0 +1,41 @@
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(ggforce)
+
+noprep_training_data <- read.csv("Files and scripts/noprep_model_history_log.csv")
+padded_training_data <- read.csv("Files and scripts/padded_model_history_log.csv")
+cropped_training_data <- read.csv("Files and scripts/cropped_model_history_log.csv")
+
+comb_data <- noprep_training_data %>% mutate(Model = 'No Prep') %>%
+  rbind(padded_training_data %>% mutate(Model = 'Padded')) %>%
+  rbind(cropped_training_data %>% mutate(Model = 'Cropped'))
+
+
+data_to_plot <- comb_data %>% 
+  rename(Train_accuracy = accuracy, 
+         Train_loss = loss, 
+         Train_sparse_top_3_categorical_accuracy = sparse_top_k_categorical_accuracy,
+         val_sparse_top_3_categorical_accuracy = val_sparse_top_k_categorical_accuracy) %>%
+  pivot_longer(Train_accuracy:val_sparse_top_3_categorical_accuracy, names_to = "temp") %>%
+  separate(temp, c("data_type", "measure"), "_", extra = 'merge') %>%
+  mutate(data_type = if_else(data_type == 'Train', 'Training', 'Validation'),
+         measure = case_when(measure == 'accuracy' ~ 'Accuracy',
+                             measure == 'loss' ~ 'Categorical Cross Entropy',
+                             measure == 'sparse_top_3_categorical_accuracy' ~ 'Top 3 Accuracy')) %>%
+  group_by(Model, data_type, measure) %>%
+  mutate(best_val_acc = if_else(data_type == 'Validation' & measure == 'Accuracy' & value == max(value), 1, 0)) %>%
+  ungroup()
+
+data_to_plot %>%
+  ggplot(aes(x = epoch, y= value, group = Model)) + 
+  geom_line(aes(col = Model), size = 1.25) + 
+  geom_mark_ellipse(aes(filter = best_val_acc == 1), expand = unit(2, "mm")) +
+  facet_grid(measure~data_type, scales = 'free_y') +
+  theme_bw() +
+  labs(x = 'Epoch',
+       y = 'Measure Value',
+       caption = 'Circled validation accuracies are the best for each model and were used as the final model for testing') +
+  scale_color_brewer(palette="Set1", name = 'Pre-Processing Method') +
+  theme(legend.position = 'bottom',
+        plot.caption = element_text(face = 'italic'))

### 2020-08-11 Tex/MSc Data Science Project.tex
Write results and conclusion section

@@
 
+In this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.
 \subsection{Related Works} 
@@
 More details of the set up are detailed in \cref{sec:Model_Training} as other than the epoch and early stopping the same configuration was used for the main training process.
-The test results of this model trained on each of the 3 datasets is presented in \cref{tab:preprocres} and the second method, padding the image with a white border, achieved the best result on the test data of 15.81\% accuracy.
+The progress on the training and validation data can be seen in \cref{fig:prep_perf} and the test results of this model trained on each of the 3 datasets is presented in \cref{tab:preprocres}.
+The second method, padding the image with a white border, achieved the best result on the test data of 15.81\% accuracy.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.5]{prep_training.png}
+	\caption{Performance of different pre-processing methods throughout training of MobileNetV2}
+	\label{fig:prep_perf}
+\end{figure}
 
@@
 It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.
+
+MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNeXt50 continued to see improvement until just before the 50th epoch/ ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
 % subsection Training_Performance end
@@
 \label{sub:Results} 
-Present and discuss the basic results for each model, including a comparison with Iwana with the caveat of a slightly different dataset and different pre-processing.
+Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison.
+A visual representation of this for our models only is shown in \cref{fig:test_perf}.
+As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset did not contain the set of records; but for a comparison it is suitable.
+Inception-ResnetV2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account.
+MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres.
+ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs.
+This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.
+
+At a per-genre level Inception-ResnetV2 shows the best results on 16 out of the 30 genres for top-1 predictions, followed by 10 for AlexNet.
+If we further explore the seemingly best performing classes in ResNeXt50 we can see by looking at the precision as presented in \cref{tab:test_prec} that in the genres where this model seems to perform well the precision is low (when compared to for example Inception-ResnetV2) which provides further evidence that for some classes it combines multiple genres and just predicts one of them i.e.
+whilst the model is able to correctly predict these \emph{chosen}	classes when it sees them, it is highly likely to predict other genres as one of these as well.
+
+Overall the results from Inception-ResnetV2 are promising given the low training time and adjusted dataset; the potential improvements are discussed further in \cref{sub:Discussion}.
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{lrrrrrrrrrrrrrr}
+								   & \multicolumn{2}{c}{\textbf{LeNet*}}                                     & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{AlexNet*}}                                   & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{MobileNetV2}}                                & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{Inception-ResnetV2}}                         & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{ResNeXt50}}                                  \\
+	\textbf{Genre}                 & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} \\ \hline
+	Arts \& Photography            & 5.8                                & 11.6                               &                               & 12.1                               & 31.1                               &                               & 11.1                               & 27.9                               &                               & \textit{13.2}                      & \textit{33.7}                      &                               & 3.7                                & 11.1                               \\
+	Biographies \& Memoirs         & 5.3                                & 18.4                               &                               & 13.2                               & 29.5                               &                               & 10.0                               & 30.5                               &                               & \textit{23.2}                      & 46.3                               &                               & 20.0                               & \textit{60.0}                      \\
+	Business \& Money              & 10.0                               & 25.3                               &                               & 12.6                               & 25.8                               &                               & 4.7                                & 18.4                               &                               & \textit{13.2}                      & \textit{33.7}                      &                               & 1.1                                & 3.2                                \\
+	Calendars                      & 18.9                               & 37.9                               &                               & \textit{47.9}                      & \textit{65.3}                      &                               & 13.7                               & 25.8                               &                               & 25.8                               & 37.4                               &                               & 18.9                               & 24.7                               \\
+	Children’s                     & 24.7                               & 42.1                               &                               & \textit{42.1}                      & \textit{61.6}                      &                               & 30.0                               & 52.1                               &                               & 18.9                               & 43.2                               &                               & 11.6                               & 34.7                               \\
+	Comics \& Graphic Novels       & 15.8                               & 33.7                               &                               & 47.4                               & 67.9                               &                               & 40.5                               & 56.8                               &                               & \textit{56.3}                      & \textit{72.6}                      &                               & 47.9                               & 60.5                               \\
+	Computers \& Technology        & 29.5                               & 42.8                               &                               & \textit{44.7}                      & \textit{59.5}                      &                               & 28.4                               & 43.7                               &                               & 41.1                               & 57.4                               &                               & 17.9                               & 26.8                               \\
+	Cookbooks, Food \& Wine        & 14.2                               & 32.6                               &                               & \textit{43.7}                      & \textit{57.4}                      &                               & 23.7                               & 42.1                               &                               & 40.0                               & 55.8                               &                               & 7.9                                & 23.2                               \\
+	Crafts, Hobbies \& Home        & 7.4                                & 22.1                               &                               & 17.4                               & 36.8                               &                               & 13.7                               & 30.5                               &                               & \textit{31.6}                      & \textit{60.0}                      &                               & 2.6                                & 15.3                               \\
+	Christian Books \& Bibles      & \textit{8.4}                       & 23.7                               &                               & 7.4                                & \textit{26.3}                      &                               & 2.6                                & 12.1                               &                               & 4.2                                & 20.5                               &                               & 0.0                                & 0.5                                \\
+	Engineering \& Transportation  & 10.0                               & 21.1                               &                               & 20.0                               & 34.7                               &                               & 15.3                               & 32.1                               &                               & \textit{20.0}                      & \textit{35.3}                      &                               & 6.8                                & 24.7                               \\
+	Health, Fitness \& Dieting     & 4.2                                & 15.8                               &                               & 12.6                               & 29.5                               &                               & 4.2                                & 17.9                               &                               & \textit{12.6}                      & \textit{37.4}                      &                               & 3.7                                & 25.8                               \\
+	History                        & 6.3                                & 16.8                               &                               & 12.6                               & 27.9                               &                               & 6.8                                & 21.6                               &                               & \textit{15.8}                      & \textit{38.4}                      &                               & 1.1                                & 2.1                                \\
+	Humor \& Entertainment         & 5.3                                & 16.3                               &                               & 10.5                               & 22.6                               &                               & 0.0                                & 7.4                                &                               & \textit{16.3}                      & \textit{43.7}                      &                               & 0.0                                & 2.6                                \\
+	Law                            & 14.7                               & 25.8                               &                               & 25.3                               & 38.4                               &                               & 16.3                               & 27.4                               &                               & 17.9                               & 37.9                               &                               & \textit{26.8}                      & \textit{45.3}                      \\
+	Literature \& Fiction          & 3.2                                & 12.1                               &                               & \textit{11.1}                      & 22.6                               &                               & 6.3                                & 24.7                               &                               & 10.5                               & \textit{38.9}                      &                               & 0.5                                & 2.6                                \\
+	Medical Books                  & 12.6                               & 30.0                               &                               & 19.5                               & 36.8                               &                               & 11.1                               & 27.4                               &                               & \textit{20.0}                      & \textit{45.8}                      &                               & 13.7                               & 38.9                               \\
+	Mystery, Thriller \& Suspense  & 23.7                               & 40.0                               &                               & 34.2                               & 48.9                               &                               & 12.1                               & 28.9                               &                               & \textit{34.7}                      & \textit{59.5}                      &                               & 19.5                               & 28.9                               \\
+	Parenting \& Relationships     & 14.7                               & 35.3                               &                               & \textit{24.2}                      & 39.5                               &                               & 24.2                               & \textit{43.7}                      &                               & 22.6                               & 42.1                               &                               & 4.7                                & 22.6                               \\
+	Politics \& Social Sciences    & 3.7                                & 18.4                               &                               & 6.8                                & 21.6                               &                               & \textit{26.3}                      & \textit{55.8}                      &                               & 2.6                                & 13.2                               &                               & 0.5                                & 12.6                               \\
+	Reference                      & 13.2                               & 26.8                               &                               & \textit{20.0}                      & \textit{34.2}                      &                               & 11.1                               & 21.6                               &                               & 8.9                                & 27.9                               &                               & 2.6                                & 18.4                               \\
+	Religion \& Spirituality       & 8.4                                & 27.9                               &                               & 16.3                               & 31.6                               &                               & 7.9                                & 26.3                               &                               & \textit{36.3}                      & \textit{55.8}                      &                               & 0.0                                & 0.5                                \\
+	Romance                        & 27.4                               & 43.2                               &                               & \textit{45.3}                      & \textit{60.5}                      &                               & 38.4                               & 52.1                               &                               & 40.5                               & 59.5                               &                               & 25.8                               & 52.1                               \\
+	Science \& Math                & 8.4                                & 26.3                               &                               & 14.2                               & 29.5                               &                               & 3.7                                & 16.3                               &                               & \textit{26.3}                      & \textit{56.3}                      &                               & 0.0                                & 0.5                                \\
+	Science Fiction \& Fantasy     & 14.7                               & 33.2                               &                               & \textit{35.8}                      & 52.6                               &                               & 21.6                               & 42.1                               &                               & 33.7                               & \textit{55.8}                      &                               & 18.9                               & 39.5                               \\
+	Self-Help                      & 13.7                               & 31.6                               &                               & 14.2                               & 33.2                               &                               & 3.7                                & 17.9                               &                               & 11.6                               & 28.4                               &                               & \textit{18.9}                      & \textit{46.8}                      \\
+	Sports \& Outdoors             & 5.3                                & 16.8                               &                               & 14.7                               & 28.4                               &                               & 1.1                                & 4.7                                &                               & \textit{27.4}                      & \textit{53.2}                      &                               & 0.0                                & 2.1                                \\
+	Teen \& Young Adult            & 7.9                                & 17.4                               &                               & 12.1                               & 28.4                               &                               & 7.4                                & 28.9                               &                               & \textit{13.2}                      & \textit{35.3}                      &                               & 0.0                                & 0.5                                \\
+	Test Preparation               & 47.9                               & 56.8                               &                               & \textit{68.9}                      & \textit{78.4}                      &                               & 34.2                               & 47.9                               &                               & 56.8                               & 74.2                               &                               & 50.5                               & 66.8                               \\
+	Travel                         & 19.5                               & 33.7                               &                               & 33.2                               & 48.4                               &                               & 7.4                                & 14.2                               &                               & \textit{44.7}                      & \textit{64.2}                      &                               & 24.2                               & 41.1                               \\ \hline
+	\textbf{Total Accuracy}         & \textbf{13.5}                      & \textbf{27.8}                      & \textbf{}                     & \textbf{24.7}                      & \textbf{40.3}                      & \textbf{}                     & \textbf{14.6}                      & \textbf{30.0}                      &                               & \textit{\textbf{24.7}}             & \textit{\textbf{45.4}}             &                               & \textbf{11.7}                      & \textbf{24.5}                      \\ \hline
+	\textit{Number of best genres} & \textit{1}                         & \textit{0}                         & \textit{}                     & \textit{10}                        & \textit{8}                         & \textit{}                     & \textit{1}                         & \textit{2}                         & \textit{}                     & \textit{16}                        & \textit{17}                        & \textit{}                     & \textit{2}                         & \textit{3}
+	\end{tabular}%
+	}
+	\caption{Top-1 and -3 Per-class recall (and overall accuracy) for the 3 trained models and the results presented by Iwana for comparison.
+Italics represent the best result for that genre in Top-1/3 recall/accuracy.
+\\
+	\tiny{*Iwana results are based on a slightly different training and test set}
+	}
+	\label{tab:results_acc}
+	\end{table}
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.5]{accuracy_by_model.png}
+	\caption{Top-1/3/5 recall (accuracy for all) of each model by genre}
+	\label{fig:test_perf}
+\end{figure}
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{lrrr}
+	\textbf{Predicted Genre}      & \multicolumn{1}{c}{\textbf{MobileNetV2}} & \multicolumn{1}{c}{\textbf{InceptionResnet-V2}} & \multicolumn{1}{c}{\textbf{ResNeXt50}} \\ \hline
+	Arts \& Photography           & 17.8                                     & 20.8                                            & 25.9                                   \\
+	Biographies \& Memoirs        & 11.5                                     & 19.3                                            & 5.5                                    \\
+	Business \& Money             & 11.0                                     & 13.4                                            & 20.0                                   \\
+	Calendars                     & 32.5                                     & 31.2                                            & 16.8                                   \\
+	Children's Books              & 24.2                                     & 36.7                                            & 13.2                                   \\
+	Christian Books \& Bibles     & 5.6                                      & 17.4                                            &                                        \\
+	Comics \& Graphic Novels      & 19.7                                     & 46.5                                            & 17.1                                   \\
+	Computers \& Technology       & 26.3                                     & 31.1                                            & 19.3                                   \\
+	Cookbooks, Food \& Wine       & 19.1                                     & 65.0                                            & 16.3                                   \\
+	Crafts, Hobbies \& Home       & 14.2                                     & 27.5                                            & 31.3                                   \\
+	Engineering \& Transportation & 10.2                                     & 54.3                                            & 12.3                                   \\
+	Health, Fitness \& Dieting    & 5.9                                      & 15.2                                            & 5.9                                    \\
+	History                       & 13.5                                     & 20.3                                            & 66.7                                   \\
+	Humor \& Entertainment        & 0.0                                      & 13.4                                            & 0.0                                    \\
+	Law                           & 16.1                                     & 24.1                                            & 9.4                                    \\
+	Literature \& Fiction         & 8.7                                      & 11.6                                            & 50.0                                   \\
+	Medical Books                 & 12.4                                     & 16.5                                            & 12.5                                   \\
+	Mystery, Thriller \& Suspense & 25.8                                     & 23.8                                            & 19.5                                   \\
+	Parenting \& Relationships    & 9.8                                      & 33.9                                            & 8.2                                    \\
+	Politics \& Social Sciences   & 5.5                                      & 11.1                                            & 5.3                                    \\
+	Reference                     & 19.4                                     & 26.2                                            & 10.2                                   \\
+	Religion \& Spirituality      & 9.6                                      & 16.7                                            &                                        \\
+	Romance                       & 14.6                                     & 44.0                                            & 14.0                                   \\
+	Science \& Math               & 11.9                                     & 11.8                                            & 0.0                                    \\
+	Science Fiction \& Fantasy    & 18.3                                     & 39.5                                            & 13.6                                   \\
+	Self Help                     & 12.3                                     & 22.9                                            & 7.2                                    \\
+	Sports \& Outdoors            & 22.2                                     & 26.0                                            & 0.0                                    \\
+	Teen \& Young Adult           & 9.5                                      & 14.2                                            & 0.0                                    \\
+	Test Preparation              & 50.4                                     & 43.7                                            & 13.0                                   \\
+	Travel                        & 50.0                                     & 17.3                                            & 8.3                                    \\ \hline
+	\textbf{Total Precision}      & \textbf{14.6}                            & \textbf{24.7}                                   & \textbf{11.7}                          \\ \hline
+	\end{tabular}%
+	}
+	\caption{Top-1 Per class and total precision for each model}
+	\label{tab:test_prec}
+	\end{table}
 
@@
 \label{sub:Further_Analysis} 
-Consider further, in depth, analysis on the best performing model, including some of the insight provided by Iwana.
+In the paper by Iwana they discuss the potential reasons for the disparity between the recall by genre and investigate some examples.
+They call out consistency of colour, objects, and text styles are key elements that impact the ability for a genre to be able to be predicted well by their models.
+Here we focus on the colour component as the strongest driver for predictability and investigate the ways in which the models may be driven by colour of the cover images.
 
-Idea: compare class average colour by predicted and actual and compare to accuracy?
+First, as displayed in \cref{fig:col_distributions}, we can see that whilst there is a wide range of average RGB values and width/aspect ratio (height is almost entirely consistent across genres), there are no standout values in colour alone that allow for easy genre classification.
+Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some of the more wider books than many other categories.
+Whilst none of this alone would be enough to provide good classification, it does suggest that explicit inclusion of dimensional information in a model could improve the performance when combined with the existing output.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.45]{distributions_col_ratio.png}
+	\caption{Distributions of various measures by genre}
+	\label{fig:col_distributions}
+\end{figure}
+
+The average overall colour per genre is presented in \cref{fig:avg_col} which does suggest some slight differences between the genres; similar to the callouts made by Iwana, Sci-Fi and Fantasy seem to be darker when compared to something like Cookbooks, Food and Wine.
+We can compare the average colour of the true classes vs those predicted by each model and visualise how these correlate with the recall per class as shown in \cref{fig:col_distance}.
+Here the colour distance is defined as
+
+\begin{equation}
+	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
+\end{equation}
+where R/G/B are the average red, green, and blue values for each predicted and class respectively.
+We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class.
+This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements.
+This supports the claims made by Iwana in their work.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.45]{avg_col_plot.png}
+	\caption{Average cover colour by genre}
+	\label{fig:avg_col}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.6]{col_dist_plot.png}
+	\caption{Colour Distance vs Recall per genre by model}
+	\label{fig:col_distance}
+\end{figure}
+
+Finally, we can review the \emph{class overlap} for each model.
+We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where A has a higher probability) i.e.
+\begin{equation}
+	CO(A, B) \stackrel{\mathrm{def}}{=} \min\left\{\frac{P(A)}{P(B)}, \frac{P(B)}{P(A)}\right\}
+\end{equation}
+The per genre pair class overlap is defined as the average class overlap across all predictions.
+As we can see in \cref{fig:class_overlap}, where only class overlaps of greater than 0.5 are displayed, the ResNeXt50 model sees large amounts of class overlap between many classes, which agrees with what we saw before that certain classes were being \emph{confused} for each other by the model.
+Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResnetV2 sees none.
+When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; Self-Help has overlap with Health, Fitness and Dieting, Science and Maths has overlap with Medical Books, and Religion and Spirituality has overlap with Christian Books and Bibles.
+All of these are logical overlaps and were a human tasked with classifying books based on their cover into these categories it is highly probable that different people would classify them differently.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.6]{class_pred_overlap.png}
+	\caption{Average class overlap by model}
+	\label{fig:class_overlap}
+\end{figure}
 % section Further_Analysis end 
@@
 \label{sub:Feature_Visualisation} 
-Detail the approach taken to try and visualise the ideal input to trigger the most accurate classes in an attempt to understand what feature the model is picking up on. Show output and discuss impact.
+The goal of using feature visualisation was to attempt to understand what, if any, underlying structure or information the models were using to aid their classification.
+To do this we use activation maximisation as implemented into the python package \emph{keras-vis}\cite{keras-vis} to do all the work.
+To produce the images the final layer activation function was changed from softmax to linear to ensure that class is being visualised as opposed to minimising other classes, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}.
+L2 normalisation was used as the default to attempt to avoid pure static images being produced.
+
+We ran the activation maximisation on the Inception-ResnetV2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images.
+The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine.
+During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure imagenet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).
+
+The results of this attempt can be seen in \cref{fig:act_max_output} and whilst it may be possible to argue that there are some differences between these images that you could prescribe to the common layout of these type of genres covers, it would be difficult to say this with any conviction.
+Overall the images are generic and provided no additional insight to understand what features the model is identifying to use for the classification.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.6]{visualize-dense-layer.png}
+	\caption{Activation maximisation outputs for the 3 best predicted classes of the Inception-ResnetV2 model}
+	\label{fig:act_max_output}
+\end{figure}
 % section Feature_Visualisation end
 
-\subsection{Discussion} 
+\subsection{Discussion \& Future Work} 
 \label{sub:Discussion} 
-Discussion around the reasons we have seen some of these results if not already covered in the previous subsections.
+Overall the Inception-ResnetV2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results.
+The counterpoint to this is that the AlexNet model as trained by Iwana et al was trained for 450,000 epochs where out model was trained for less than 50 epochs total.
+We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greatly difference between the results.
+Despite training on slightly less data (due to the validation set), we can see that more modern and powerful CNNs are able to predict book genres with slightly improved accuracy compared to older methods.
 
-Future work:
-use different interpolation method
+There are lots of directions this work could continue to go in the future; as mentioned before there is still more investigation that can be done surrounding the pre-processing of data, either varying the interpolation methods for downsizing or re-testing the pre-processing methods used specifically on the end model architecture for a longer period of time.
+Our models could be trained for more epochs with a greater requirement on early stopping to allow for long term improvements to be made that may not be evident over a few hundred epochs.
+Also, additional information such as the width and height or just aspect ratio could be passed to the model as well in an attempt to improve accuracy this way.
+The final fully connected part of the network could also be increased in complexity to contain more layers as opposed to just the single layer used in our work.
+
+If resources were no constraint some of the larger improvements to the work would be to tune the hyperparameters of the model, and even train the actual core part of the architecture itself rather than relying entirely on transfer learning and freezing the weights for this component.
+To do this would require a far greater dataset of images so a generator may have to be used or a larger dataset collected first.
+Finally, as shown by the results in \cref{tab:results_acc} there is not a single model that performs best across all genres, so an ensemble method may be able to increase the accuracy at the cost of increased complexity.
+Many of these suggestions are beyond the scope of many citizen data scientists resources and would also lead to a model that may have a prediction time that makes it prohibitive for any real-time use.
 % subsection Discussion end 
@@
 \label{sec:Conclusion} 
-Discuss the project as a whole, the impact of the results when compared to the literature, and what future work there could be off the back of this.
+In this work we have attempted to improve on the work by Iwana et al and better answer the question can a CNN judge a book by its cover.
+We reprocessed the dataset initially collected by Iwana and applied multiple pre-processing techniques in an attempt to identify the best method to deal with the impact of different shaped images.
+We then proceeded to train 3 modern models via transfer learning on a padded version of the dataset and evaluated the results.
+Finally we attempted to produce ideal class images using activation maximisation for the best model and genres.
+
+We have seen that using a more modern CNN architecture, namely Inception-ResnetV2, we are able to improve on the results published by Iwana slightly whilst training for substantially less epochs.
+We have also provided some further supporting insight into the components that factor into book cover design i.e.
+colour, and identified where classifications are failing where this might be due to overlap between classes.
+
+There is still lots to explore in this space as we discussed in \cref{sub:Discussion} and hopefully this work provides a springboard to enable more complex CNNs to be used for genre identification of book covers in the future.
+The work itself is unlikely to provide any value for real-world use as the accuracy is too low to be of use for either consumers or even cover designers as a form of QA.
+Genre prediction from cover images remains a complex topic and given the subjectivity of the classifications is unlikely to ever be fully solved, however the work produced here gets a little closer than we were before.
+So it seems for now that, if you are a CNN, you should follow the original advice and not try to judge a book by its cover.
 % section Conclusion end

### 2020-08-11 Files and scripts/calculate precision.R

@@ -1,0 +1,56 @@
+# Libraries ---------------------------------------------------------------
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+
+
+# Load Data ----------------------------------------------------------------
+c_names <- c('Arts & Photography',
+             'Biographies & Memoirs',
+             'Business & Money',
+             'Calendars',
+             "Children's Books",
+             'Comics & Graphic Novels',
+             'Computers & Technology',
+             'Cookbooks, Food & Wine',
+             'Crafts, Hobbies & Home',
+             'Christian Books & Bibles',
+             'Engineering & Transportation',
+             'Health, Fitness & Dieting',
+             'History',
+             'Humor & Entertainment',
+             'Law',
+             'Literature & Fiction',
+             'Medical Books',
+             'Mystery, Thriller & Suspense',
+             'Parenting & Relationships',
+             'Politics & Social Sciences',
+             'Reference',
+             'Religion & Spirituality',
+             'Romance',
+             'Science & Math',
+             'Science Fiction & Fantasy',
+             'Self-Help',
+             'Sports & Outdoors',
+             'Teen & Young Adult',
+             'Test Preparation',
+             'Travel',
+             'Filename')
+res_results <- read.csv("Files and scripts/Resnext_preds.csv",
+                        col.names = c_names, check.names = FALSE)
+test_data <- read.csv("Files and scripts/dims_and_colours_to_analyse_test.csv")
+
+
+
+# Calc Precision ----------------------------------------------------------
+
+res_results %>%
+  pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') %>%
+  group_by(Filename) %>%
+  arrange(desc(Prediction_Perc)) %>%
+  slice(1) %>% 
+  left_join(test_data) %>%
+  select(Filename:Category, -X) %>%
+  mutate(correct = if_else(Predicted_Class == Category, 1, 0)) %>%
+  group_by(Predicted_Class) %>%
+  summarise(precision = sum(correct)/n())

### 2020-08-11 Files and scripts/combined_file.R

@@ -367,7 +367,7 @@ comb_overlap_data %>%
 
 
 # Analysis ----------------------------------------------------------------
-
+#class overlap
 comb_overlap_data %>% 
   filter(avg_overlap < 1) %>%
   group_by(model) %>% 
@@ -375,3 +375,33 @@ comb_overlap_data %>%
   slice(1:5) %>%
   ungroup() %>%
   select(Predicted_Class.x, Predicted_Class.y) %>% distinct()
+
+#precision
+mob_results %>% mutate(model = 'MobileNetV2') %>%
+  rbind(inc_results %>% mutate(model = 'InceptionResnetV2')) %>%
+  rbind(res_results %>% mutate(model = 'ResneXt50')) %>%
+  pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') %>%
+  group_by(Filename, model) %>%
+  arrange(desc(Prediction_Perc)) %>%
+  slice(1) %>% 
+  left_join(test_data) %>%
+  select(Filename, Category, Predicted_Class, model) %>%
+  mutate(correct = if_else(Predicted_Class == Category, 1, 0)) %>%
+  group_by(Predicted_Class, model) %>%
+  summarise(precision = sum(correct)/n()) %>%
+  pivot_wider(names_from = model, values_from = precision) %>%
+  clipr::write_clip()
+
+mob_results %>% mutate(model = 'MobileNetV2') %>%
+  rbind(inc_results %>% mutate(model = 'InceptionResnetV2')) %>%
+  rbind(res_results %>% mutate(model = 'ResneXt50')) %>%
+  pivot_longer(`Arts & Photography`:Travel, names_to = 'Predicted_Class', values_to = 'Prediction_Perc') %>%
+  group_by(Filename, model) %>%
+  arrange(desc(Prediction_Perc)) %>%
+  slice(1) %>% 
+  left_join(test_data) %>%
+  select(Filename, Category, Predicted_Class, model) %>%
+  mutate(correct = if_else(Predicted_Class == Category, 1, 0)) %>%
+  group_by(model) %>%
+  summarise(precision = sum(correct)/n()) %>%
+  pivot_wider(names_from = model, values_from = precision) 

### 2020-08-15 Tex/MSc Data Science Project.tex
More sections written

@@
 \documentclass[12pt]{article}
+\usepackage{arydshln}
+\usepackage{multirow}
 \usepackage{graphicx}
+\usepackage{bm}
 \usepackage{amsmath}
@@
 \usepackage{subcaption}
+\usepackage{mathrsfs}
 \usepackage{mdframed}
@@
 \label{sec:Theory} 
+Whilst in this work we do not detail the full design and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks.
+In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}.
+We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks.
+For this section and the remainder of this work, unless otherwise specified, we observe the following notational conventions:
+\begin{itemize}
+	\item Unless otherwise stated all tensors (including vectors and matrices) are indexed from 1.
+	\item Roman indices will be used to reference specific components of a tensor within an equation, Greek indices will be used for dimensions to be operated over only.
+	\item Dimensions of each tensor when referring to an image or feature map will be referred to in the order of height, width, (colour) channel (interchangeable with depth), including in the indices.
+	\item Bold items will refer to full vectors of matrices, with it clear which one from the context.
+	\item $\vec{X}$ will always refer to the vector input patterns i.e. images of a dataset. Each element contains more dimensions so this is technically a 4D-tensor.
+	\item $\vec{Y}$ will always refer to the vector of target classes for each of those patterns. Each element is a single value.
+	\item Bracketed roman superscripts i.e. $w^{(i)}$ will be used to refer to, in this case the i\textsuperscript{th}, elements of these vectors. These are 3D-tensors.
+
+	\item Greek subscripts will be used to refer to elements along the dimensions of each pattern i.e.
+$x^{(i)}_{\alpha, \beta, \gamma}$ refers to the $alpha$ element of down on the tensor, the $\beta$ element across on the tensor, and the $\gamma$ channel of the tensor.
+	\item $\bigodot$ will refer to element wise multiplication of matrices and tensors across all dimensions where indices are not explicitly summed over in an equation.
+	\item $\bigoplus$ will refer to element wise addition of tensors across all dimensions where indices are not expli
+\end{itemize}
 \subsection{Components of Convolutional Neural Networks} 
 \label{sub:Components_of_Convolutional_Neural_Networks} 
-Describe the maths and the theory behind the various components of CNNs to the best of my ability, referring to the source papers for more complex combined units where sensible. 
+\subsubsection{Activation functions}
+Most layers of a neural network have some kind of activation function; these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range).
+Some normalisation is also usually applied which is discussed later.
+For our purposes there are only 3 activation functions that are relevant for this work.
+
+Linear activation is exactly what it sounds like, there is no transformation to the output.
+Rectified Linear Unit (ReLU) is a function that returns the input when the value is greater than or equal to zero, and zero otherwise.
+Finally, softmax is used almost exclusively in the final layer of a network to transform the outputs into values that sum to 1 (to be interpreted as the probability of that record belonging to the class that neuron represents).
+These are defined mathematically in \cref{eq:act_linear,eq:act_relu,eq:act_softmax}.
+
+\begin{align}
+	f_{linear}(x) &= x \label{eq:act_linear}\\
+	f_{ReLU}(x)   &= \max(0, x) \label{eq:act_relu}\\
+	f_{softmax}(\bm{x})_i &= \frac{e^{x_i}}{\sum_{\alpha=1}^{K}{e^{x_\alpha}}}\label{eq:act_softmax}
+\end{align}
+Where the non-bold $x$s are just individual outputs, but the softmax function requires all K outputs of the layer for calculation.
+Technically, most pre-built models in Tensorflow use ReLU6 as opposed to pure ReLU, this version is shown in \cref{eq:act_relu6}.
+\begin{equation}
+	f_{ReLU6}(x) = min(max(0, x), 6) \label{eq:act_relu6}
+\end{equation}
+
+\subsubsection{Convolutional layers}
+A convolutional layer, the key part of a neural network, is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN.
+A convolutional layer is defined by 3 key parameters, and the additional choice of padding.
+The 3 parameters that define a basic convolutional layer are the number of filters (analogous to the number of neurons in a dense layer) which is also the number of channels the output of that convolution will have, the size of the filters, and the stride (or step size) of the filter.
+
+A filter (also called a kernel) is applied atop of the input, starting at the top left of the input, and pointwise multiplication is applied to each set of elements, then their sum is computed.
+In the case of multiple input channels the same filter is applied over all channel and their outputs are summed.
+This value is the the output value of the top left cell of the output tensor.
+The filter is then moved left by the number of elements defined by the stride; when the end of the row has been reached the filter is move down by the same stride.
+This process is repeated until the filter has been applied to the entire input.
+Once all filters have been applied this set of outputs is referred to as the feature map.
+This process is demonstrated in \cref{fig:conv_process} and quantified mathematically as
+
+\begin{equation}
+	m_{i,j,k} = \sum_{\chi = 1}^{C}\sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, k} \times x_{(\alpha +(i-1)*s), (\beta + (j-1)*s),\chi}
+\end{equation}
+where $\bm{m}_k$ is the feature map for output channel $k$, C is the number of channels in the input, $x$ is the input, $f$ is the $k^{th}$ kernel of size $l\times l$, $s$ is the stride.
+In total for a square input and kernel this leads to an output of the dimensions described in
+
+\begin{equation}
+	\label{eq:feat_map1}
+	\dim(\mathscr{M}) = \left[\frac{m-f}{s} +1 \right] \times \left[\frac{m-f}{s} +1 \right] \times n_{f}
+\end{equation}
+where $\mathscr{M}$ is the full feature map, $m$ is the width and height of the input, $f$ is the width and height of the filter, $s$ is the stride, and $n_f$ is the number of filters.
+
+Using this method it means that, unless the filter size and stride are both equal to 1, certain parts of the input will never be exposed to the centre of the filter, and the output of the process will always lead to a smaller size.
+This can be an issue when multiple convolutional layers are applied in sequence as whilst the number of channels may increase the size of each channel will decrease.
+An approach that removes this problem is to first pad the input with 0s on all sides, the size of the padding being determined by the goal output size as well as the filter size and stride.
+By applying padding the output size of the image changes from that described in \cref{eq:feat_map1} to instead be
+\begin{equation}
+	\label{eq:feat_map2}
+	\dim(\mathscr{M}) = \left[\frac{m+2p-f}{s} +1 \right] \times \left[\frac{m+2p-f}{s} +1 \right] \times n_{f}
+\end{equation}
+where $p$ is the padding size on one side.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.25]{cnn_application.png}
+	\caption{An example of a 3x3 filter being applied to a single channel input, with stride of 1 and 0 padding. \cite{Yamashita2018}}
+	\label{fig:conv_process}
+\end{figure}
+
+Whilst the number of filters, filter size, and stride are hyperparameters that must be set by the user, the actual values of each filter in the layers are learned as part of the backpropagation training of the network.
+
+\subsubsection{Depth-Wise Convolutions}
+Depth wise convolutions are similar to normal convolutional layers in all ways except one; instead of applying the filter to each channel of the input and summing the outputs to produce a single output channel we instead keep these outputs separate and produce an output with as many channels as the input per filter.
+When followed by a 1x1 convolution (also known as a pointwise convolution) this has the effect of merging the layers via addition and combined these two processes actually have a reduced computational cost when compared to a traditional convolution \cite{Sandler2018}.
+
+\subsubsection{Pooling}
+Pooling is an operation used within convolutional neural networks to reduce the size of channels/feature maps within the network.
+Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area.
+These operations tend to either be maximum or average and is determined either by the user of the CNN architecture.
+These are applied per-channel and no aggregation across channels is completed afterwards.
+The dimensions of the output after pooling can be calculated using the same method as before displayed in \cref{eq:feat_map2}.
+Traditionally (and in our work) pooling is completed with no padding.
+Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned.
+
+\subsubsection{Batch Normalisation}
+Batch normalisation was introduced in a 2015 paper by Sergey Ioffee and Christain Szegedy\cite{Ioffe2015} and is a step introduced into neural networks, including CNNs, between the layers and the activation functions.
+Batch normalisation works by applying re-scaling and re-centering transformations to the output of a layer i.e.
+transforming the output of that layer for that batch to have a mean of 0 and a standard deviation of 1.
+This has the effect of greatly reducing training time and increases accuracy by way of reducing the change in the data distribution between layers and reducing the change of this over time via backpropagation.
+
+\subsubsection{Residual Networks}
+Residual networks (ResNets) were introduced by He et al in 2015\cite{He2015a} to allow for far deeper networks than previously used.
+The basic premise is that the layer input forms part of the output directly via a scaled addition to the convolved, normalised, and activated output.
+This allows for information to be carried further through the network and was shown to achieve better results against benchmark image classification datasets.
+Further detail and formulation is available in the original paper so we do not reproduce that here as a grasp of the main concept is all that is required.
+
 % subsection Components_of_Convolutional_Neural_Networks end 
@@
 \label{sub:Transfer_Learning} 
-Describe the theory and maths behind transfer learning to the reader.
+Transfer learning is a concept in machine learning that \emph{knowledge} gained by a network in one domain could be used within another domain, either out of the box or as a starting point for training.
+In our case, the models we use have already been trained on the more complex ImageNet dataset so would be suitable to use in transfer learning to save a lot of time training the model from scratch.
+
+In transfer learning it is possible to take one of two approaches; the first is to use the existing model weight and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e.
+all but the fully connected layers at the end, and just train the classification top of the model.
+In our work we took the second approach as given the size of the models we are using training the full model would have been computationally expensive, and given the relatively small amount of training data would also have when compared to the number of parameters in the model it would have been unlikely to have improved accuracy.
 % subsection Transfer_Learning end
@@
 The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}.
-The original csv containing all image URLs was split into files of no more than 20,000 records each to allow to the downloads to be run over multiple sessions do to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway.
+The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway.
 Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e.
@@
 Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.
-At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, and example of such an image can be seen in \textcolor{red}{TODO: add image}\cref{fig:TBC}.
-The terms used in a regex search of the titles were \emph{"boxed set", "boxset", "box set", "anthology", "bundle", "\#-book", "\# book"}.
+At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, an example of such an image can be seen in \cref{fig:bad_cover}.
+The terms used in a regex search of the titles were \emph{``boxed set'', ``boxset'', ``box set'', ``anthology'', ``bundle'', ``\#-book'', and ``\# book''}.
 264 records within the bookcover30 dataset matched at least one of these terms and were extracted into a new folder for manual review.
@@
 No verification of the class assignment to images was performed so any errors that existed in the original data would have persisted and we continue to use the single class chosen by the original authors in the case of multi-class books.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.5]{0345535529.jpg}
+	\caption{An example of a rejected \emph{``cover''} image as it did not meet our criteria.}
+	\label{fig:bad_cover}
+\end{figure}
 
@@
 
-The first pre-processing method (\emph{NoPrep}) was to simply rescale the image directly down to a 299x299 shape using inter-cubic interpolation \textcolor{red}{TODO: explain this mathematically}.
-This was chosen at the time as a common all-purpose interpolation method, however after later research this method is often chosen to enlarging images, not shrinking them so the potential to use a better interpolation method could be used in future work.
+The first pre-processing method (\emph{NoPrep}) was to simply rescale the image directly down to a 299x299 shape using inter-cubic interpolation.
+This method uses a 4x4 neighbourhood around the new pixel location to smooth the transition between boundaries as smoothly as possible and the implementation in the \emph{opencv} python library was used throughout.
+This was chosen at the time as a common all-purpose interpolation method, however after later research this method is often chosen to enlarging images, not shrinking them so the potential to use a better interpolation method (e.g.
+inter-area) could be used in future work.
 
@@
 \label{sub:MobileNetV2} 
+Full technical details of the specification is available in the paper by Sandler at al.
+\cite{Sandler2018} where the model is first published.
+The version we use is identical using the Keras implementation with the exception that we do not use a final Conv2D layer but instead use a dense fully connected layer of size 30 to use for our classification.
+The table and captions explaining the architecture is copied from their paper to here for convenience.
+The only other difference not detailed in these tables is batch normalisation is applied after each convolution but before the ReLU, and zero padding is applied where a stride of 2 is used.
+
+\begin{table}[]
+	\centering
+	\begin{tabular}{c|c|c}
+	\hline
+	\textbf{Input}                            & \textbf{Operator}           & \textbf{Output}                            \\ \hline \hline
+	$h \times w \times k$                     & 1x1 conv2d, ReLU6           & $h \times w \times tk$                     \\
+	$h \times w \times tk$                    & 3x3 dwise stride=$s$, ReLU6 & $\frac{h}{s} \times \frac{w}{s} \times k$  \\
+	$\frac{h}{s} \times \frac{w}{s} \times k$ & linear 1x1 covn2d           & $\frac{h}{s} \times \frac{w}{s} \times k'$ \\ \hline
+	Iff $s \equiv 1$ \& $k \equiv k'$ then $h \times w \times k$ & add original input           & $h \times w \times k'$ \\ \hline
+	\end{tabular}
+	\caption{Bottleneck residual block transforming from $k$ to $k'$ channels, with stride $s$, and expansion factor $t$.
+	}
+	\label{tab:bottleneck}
+	\end{table}
+
+\begin{table}[]
+	\centering
+	\begin{tabular}{c|c|c|c|c|c}
+	\hline
+	\textbf{Input}         & \textbf{Operator} & \textbf{t} & \textbf{c} & \textbf{n} & \textbf{s} \\ \hline \hline
+	$224^2$ $\times$ 3         & conv2d, ReLU6           & -          & 32         & 1          & 2          \\
+	$112^2$ $\times$ 32        & bottleneck        & 1          & 16         & 1          & 1          \\
+	$112^2$ $\times$ 16        & bottleneck        & 6          & 24         & 2          & 2          \\
+	$56^2$ $\times$ 24         & bottleneck        & 6          & 32         & 3          & 2          \\
+	$28^2$ $\times$ 32         & bottleneck        & 6          & 64         & 4          & 2          \\
+	$14^2$ $\times$ 64         & bottleneck        & 6          & 96         & 3          & 1          \\
+	$14^2$ $\times$ 96         & bottleneck        & 6          & 160        & 3          & 2          \\
+	$7^2$ $\times$ 190         & bottleneck        & 6          & 320        & 1          & 1          \\
+	$7^2$ $\times$ 320         & conv2d 1x1, ReLU        & -          & 1280       & 1          & 1          \\
+	$7^2$ $\times$ 1280        & avgpool 7x7       & -          & -          & 1          & -          \\
+	1 $\times$ 1 $\times$ 1280 & dense             & -          & 30         & -          & -          \\ \hline
+	\end{tabular}%
+	\caption{MobileNetV2 : Each line describes a sequence of 1 or more identical (modulo stride) layers, repeated n times.
+All layers in the same sequence have the same number c of output channels.
+The first layer of each sequence has a stride s and all others use stride 1.
+All
+	spatial convolutions use 3 x 3 kernels. The expansion factor t is always applied to the input size as described in \cref{tab:bottleneck}.
+	}
+	\label{tab:mobilenetv2_arch}
+	\end{table}
 
@@
 \label{sub:Inception-ResnetV2} 
+The architecture, with all the relevant filter and stride settings are detailed within the original paper by Szegedy et al.
+\cite{Szegedy2016} however the Keras implementation seems to differ slightly and the original paper does not provide a full spec in a single place so it is composed here.
+For simplicity batch normalisation is not explicitly written but is used after every convolution operator and ReLU activation is used throughout.
+A summarised version is available in \cref{fig:incepresnetv2}\cite{Alemi2016}.
+
+\begin{figure}
+	\centering
+	\captionsetup{justification=centering}
+	\includegraphics[scale=0.25]{inception_resnetv2.png}
+	\caption{High level view of the inception-resnetv2 architecture. The early classifier (bottom of image) is not used in our work.}
+	\label{fig:incepresnetv2}
+\end{figure}
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|c|c|c}
+	\textbf{Block Name}      & \textbf{LayerID} & \textbf{Input} & \textbf{Operator} & \textbf{Filters} & \textbf{Shape} & \textbf{Scale} \\ \hline \hline
+	Inception-ResNet-A block & 0                & -              & conv2d            & 32               & 1              & -              \\
+	Inception-ResNet-A block & 1                & -              & conv2d            & 32               & 1              & -              \\
+	Inception-ResNet-A block & 2                & 1              & conv2d            & 32               & 3              & -              \\
+	Inception-ResNet-A block & 3                & -              & conv2d            & 32               & 1              & -              \\
+	Inception-ResNet-A block & 4                & 3              & conv2d            & 48               & 3              & -              \\
+	Inception-ResNet-A block & 5                & 4              & conv2d            & 64               & 3              & -              \\
+	Inception-ResNet-A block & 6                & 0, 2, 5        & concat            & -                & -              & -              \\
+	Inception-ResNet-A block & 7                & -, 6           & add w/ scale      & -                & -              & (1, 0.17) \\ \hline  
+	\end{tabular}
+	}
+	\caption{Structure of the Inception-ResNet-A block.
+Shape unless otherwise specified is square, and scale is the multiplier of the inputs of the addition.
+- in input represents the input to the block.}
+	\label{tab:incepresa}
+	\end{table}
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|c|c|c}
+	\textbf{Block Name}        & \textbf{LayerID} & \textbf{Input} & \textbf{Operator} & \textbf{Filters} & \textbf{Shape} & \textbf{Scale} \\ \hline \hline
+	Inception-ResNet-B block & 0                & -              & conv2d            & 192              & 1              & -              \\
+	Inception-ResNet-B block   & 1                & -              & conv2d            & 128              & 1              & -              \\
+	Inception-ResNet-B block   & 2                & 1              & conv2d            & 160              & (1, 7)         & -              \\
+	Inception-ResNet-B block   & 3                & 2              & conv2d            & 192              & (7, 1)         & -              \\
+	Inception-ResNet-B block   & 4                & 0, 3           & concat            & -                & -              & -              \\
+	Inception-ResNet-B block   & 5                & `-, 4          & add w/ scale      & -                & -              & (1, 0.1)     \\ \hline  
+	\end{tabular}%
+	}
+	\caption{Structure of the Inception-ResNet-B block.
+Shape unless otherwise specified is square, and scale is the multiplier of the inputs of the addition.
+- in input represents the input to the block.}
+	\label{tab:incepresb}
+	\end{table}
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|c|c|c}
+	\textbf{Block Name}      & \textbf{LayerID} & \textbf{Input} & \textbf{Operator} & \textbf{Filters} & \textbf{Shape} & \textbf{Scale} \\ \hline \hline
+	Inception-ResNet-C block & 0                & -              & conv2d            & 192              & 1              & -              \\
+	Inception-ResNet-C block & 1                & -              & conv2d            & 192              & 1              & -              \\
+	Inception-ResNet-C block & 2                & 1              & conv2d            & 224              & (1, 3)         & -              \\
+	Inception-ResNet-C block & 3                & 2              & conv2d            & 256              & (3, 1)         & -              \\
+	Inception-ResNet-C block & 4                & 0, 3           & concat            & -                & -              & -              \\
+	Inception-ResNet-C block & 5                & `-, 4          & add w/ scale      & -                & -              & (1, 1) \\ \hline  
+	\end{tabular}%
+	}
+	\caption{Structure of the Inception-ResNet-C block.
+Shape unless otherwise specified is square, and scale is the multiplier of the inputs of the addition.
+- in input represents the input to the block.}
+	\label{tab:incepresc}
+	\end{table}
+
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|c|c|c|c}
+	\textbf{Block Name}               & \textbf{LayerID} & \textbf{Input} & \textbf{Operator} & \textbf{Filters} & \textbf{Shape} & \textbf{Strides}       & \textbf{N} \\ \hline\hline
+	Stem                              & 0                & -              & conv2d            & 32               & 3              & 2                      & 1          \\
+	Stem                              & 1                & 0              & conv2d            & 32               & 3              & 1                      & 1          \\
+	Stem                              & 2                & 1              & conv2d            & 64               & 3              & 1                      & 1          \\
+	Stem                              & 3                & 2              & maxpooling        & -                & 3              & 2                      & 1          \\
+	Stem                              & 4                & 3              & conv2d            & 80               & 1              & 1                      & 1          \\
+	Stem                              & 5                & 4              & conv2d            & 192              & 3              & 1                      & 1          \\
+	Stem                              & 6                & 5              & maxpooling        & -                & 3              & 2                      & 1          \\\hdashline
+	Inception-A block                 & 7                & 6              & conv2d            & 96               & 1              & 1                      & 1          \\
+	Inception-A block                 & 8                & 6              & conv2d            & 48               & 1              & 1                      & 1          \\
+	Inception-A block                 & 9                & 8              & conv2d            & 64               & 5              & 1                      & 1          \\
+	Inception-A block                 & 10               & 6              & conv2d            & 64               & 1              & 1                      & 1          \\
+	Inception-A block                 & 11               & 10             & conv2d            & 96               & 3              & 1                      & 1          \\
+	Inception-A block                 & 12               & 11             & conv2d            & 96               & 3              & 1                      & 1          \\
+	Inception-A block                 & 13               & 6              & avgpooling        & -                & 3              & 1                      & 1          \\
+	Inception-A block                 & 14               & 13             & conv2d            & 64               & 1              & 1                      & 1          \\
+	Inception-A block                 & 15               & 7, 9, 12, 14   & concat            & -                & -              & -                      & -          \\\hdashline
+	\textit{Inception-ResNet-A block} & 16               & 15             & -                 & -                & -              & -                      & 10         \\\hdashline
+	Reduction-A block                 & 17               & 16             & conv2d            & 384              & 3              & 2                      & 1          \\
+	Reduction-A block                 & 18               & 16             & conv2d            & 256              & 1              & 1                      & 1          \\
+	Reduction-A block                 & 19               & 18             & conv2d            & 256              & 3              & 1                      & 1          \\
+	Reduction-A block                 & 20               & 19             & conv2d            & 284              & 3              & 2                      & 1          \\
+	Reduction-A block                 & 21               & 16             & maxpooling        & -                & 3              & 2                      & 1          \\
+	Reduction-A block                 & 23               & 17, 20, 21     & concat            & -                & -              & -                      & -          \\\hdashline
+	\textit{Inception-ResNet-B block} & 24               & 23             & -                 & -                & -              & -                      & 20         \\
+	Reduction-B block                 & 25               & 24             & conv2d            & 256              & 1              & 1                      & 1          \\
+	Reduction-B block                 & 26               & 25             & conv2d            & 384              & 3              & 2                      & 1          \\
+	Reduction-B block                 & 27               & 24             & conv2d            & 256              & 1              & 1                      & 1          \\
+	Reduction-B block                 & 28               & 27             & conv2d            & 288              & 3              & 2                      & 1          \\
+	Reduction-B block                 & 29               & 24             & conv2d            & 256              & 1              & 1                      & 1          \\
+	Reduction-B block                 & 30               & 29             & conv2d            & 288              & 3              & 1                      & 1          \\
+	Reduction-B block                 & 31               & 30             & conv2d            & 230              & 3              & 2                      & 1          \\
+	Reduction-B block                 & 32               & 29             & maxpooling        & -                & 3              & 2                      & 1          \\
+	Reduction-B block                 & 33               & 26, 28, 31, 32 & concat            & -                & -              & -                      & -          \\\hdashline
+	\textit{Inception-ResNet-C block} & 34               & 33             & -                 & -                & -              & -                      & 10         \\\hdashline
+	Final convolution block           & 35               & 34             & conv2d            & 1536             & 1              & 1                      & 1          \\
+	Global Pooling                    & 36               & 35             & avgpooling        & -                & -              & -                      & -          \\
+	Dense                             & 37               & 36             & dense (30)        & -                & -              & - 				& -          \\ \hline
+	\end{tabular}%
+	}
+	\caption{Structure of the Inception-ResNetV2 model.
+Shape unless otherwise specified is square, and N is the number of times that layer or block is repeated.
+For information on the Inception-ResNet blocks see \cref{tab:incepresa,tab:incepresb,tab:incepresc}.
+- represents a non-relevant argument or in the case of input, the input to the model.}
+	\label{tab:incepresfull}
+	\end{table}
 
@@
 \label{sub:ResNeXt50} 
+The full details of the ResNeXt50 architecture and in what way it differs from and improves upon regular ResNets is available in the paper by Xie et al.\cite{Xie2016}.
+We reproduce here the table (\cref{tab:resnextarc})explaining the architecture and an image (\cref{fig:resnextblocks}) explaining the blocks from their paper for convenience.
+These models differ in their use of a \emph{cardinality} term, C.
+This is defined as the number of times a repeated series of convolutional filters in parallel with the exact same parameters are used within a block.
+All layers are followed by batch normalisation and ReLU activation.
 
+\begin{table}[]
+	\centering
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|c|c|c|c}
+	\textbf{Stage}         & \textbf{Output}                 & \textbf{Operation}                                                                   & \textbf{Filters} & \textbf{Shape} & \textbf{Stride} & \textbf{Cardinality (C)} & \textbf{N}         \\ \hline \hline
+	conv1                  & $112 \times 112$                & conv2d                                                                               & 64               & 7              & 2               & 1                        & 1                  \\ \hline
+	\multirow{4}{*}{conv2} & \multirow{4}{*}{$56 \times 56$} & max pooling                                                                          & -                & 3              & 2               & 1                        & 1                  \\ \cline{3-8}
+						   &                                 & \multirow{3}{*}{ResNext Block}                                                       & 128              & 1              & 1               & 1                        & \multirow{3}{*}{3} \\ \cline{4-7}
+						   &                                 &                                                                                      & 128              & 3              & 1               & 32                       &                    \\ \cline{4-7}
+						   &                                 &                                                                                      & 256              & 1              & 1               & 1                        &                    \\ \hline
+	\multirow{3}{*}{conv3} & \multirow{3}{*}{$28 \times 28$} & \multirow{3}{*}{ResNext Block}                                                       & 256              & 1              & 1               & 1                        & \multirow{3}{*}{4} \\ \cline{4-7}
+						   &                                 &                                                                                      & 256              & 3              & 1               & 32                       &                    \\ \cline{4-7}
+						   &                                 &                                                                                      & 512              & 1              & 1               & 1                        &                    \\ \hline
+	\multirow{3}{*}{conv4} & \multirow{3}{*}{$14 \times 14$} & \multirow{3}{*}{ResNext Block}                                                       & 512              & 1              & 1               & 1                        & \multirow{3}{*}{6} \\ \cline{4-7}
+						   &                                 &                                                                                      & 512              & 3              & 1               & 32                       &                    \\ \cline{4-7}
+						   &                                 &                                                                                      & 1024             & 1              & 1               & 1                        &                    \\ \hline
+	\multirow{3}{*}{conv5} & \multirow{3}{*}{$7 \times 7$}   & \multirow{3}{*}{ResNext Block}                                                       & 1024             & 1              & 1               & 1                        & \multirow{3}{*}{3} \\ \cline{4-7}
+						   &                                 &                                                                                      & 1024             & 3              & 1               & 32                       &                    \\ \cline{4-7}
+						   &                                 &                                                                                      & 2048             & 1              & 1               & 1                        &                    \\ \hline
+						   & $1 \times 1$                    & \begin{tabular}[c]{@{}c@{}}global average pooling\\ dense (30), softmax\end{tabular} & -                & -              & -               & 1                        & 1                  \\ \hline
+	\end{tabular}%
+	}
+	\caption{Structure of the ResNext-50 model. At the end of each block the input is added in (as with the original resnet models) - see example methods in \cref{fig:resnextblocks}.}
+	\label{tab:resnextarc}
+	\end{table}
+
+	\begin{figure}
+		\centering
+		\captionsetup{justification=centering}
+		\includegraphics[scale=0.5]{resnextblocks.PNG}
+		\caption{\cite{Xie2016} Equivalent methods for each ResNeXt block highlighting the inclusion of the input \emph{shortcut} path as part of the output. In our work the rightmost option was used.}
+		\label{fig:resnextblocks}
+	\end{figure}
 % subsection ResNeXt50 end 
@@
 \label{sec:Technology}
-List of software and packages/libraries including their versions.
+Due to the use of Google Colab for the vast majority of the work is it not possible to confirm which hardware was used over the course of the work, or even that the same was used over multiple sessions.
+Software and library versions:
+\begin{itemize}
+	\item Python: 3.6.9 (linux environment on Google Colab)
+	\begin{itemize}
+		\item argparse: 1.1 
+		\item classification\_model: 0.2.2 
+		\item IPython: 5.5.0 
+		\item joblib: 0.16.0 
+		\item matplotlib: 3.2.2 
+		\item skimage: 0.16.2 
+		\item sklearn: 0.22.2.post1 
+		\item tf\_keras\_vis: 0.5.2 
+		\item tqdm: 4.41.1 
+		\item cv2: 4.1.2 
+		\item numpy: 1.18.5 
+		\item pandas: 1.0.5 
+		\item re: 2.2.1 
+		\item tensorflow: 2.3.0
+	\end{itemize}
+	\item R: 4.0.0 (64-bit windows)
+	\begin{itemize}
+		\item ggforce: 0.3.2
+		\item tidyr: 1.1.0
+		\item ggplot2: 3.3.0
+		\item dplyr: 0.8.5
+		\item Rcpp: 1.0.4.6
+		\item rstudioapi: 0.11
+		\item magrittr: 1.5
+		\item MASS: 7.3-51.5
+		\item tidyselect: 1.1.0
+		\item munsell: 0.5.0
+		\item viridisLite: 0.3.0
+		\item colorspace: 1.4-1
+		\item R6: 2.4.1
+		\item rlang: 0.4.6
+		\item fansi: 0.4.1
+		\item tools: 4.0.0
+		\item grid: 4.0.0
+		\item gtable: 0.3.0
+		\item cli: 2.0.2
+		\item clipr: 0.7.0
+		\item withr: 2.2.0
+		\item ellipsis: 0.3.0
+		\item assertthat: 0.2.1
+		\item tibble: 3.0.1
+		\item lifecycle: 0.2.0
+		\item crayon: 1.3.4
+		\item gridExtra: 2.3
+		\item farver: 2.0.3
+		\item tweenr: 1.0.1
+		\item purrr: 0.3.4
+		\item viridis: 0.5.1
+		\item vctrs: 0.3.0
+		\item glue: 1.4.0
+		\item polyclip: 1.10-0
+		\item stringi: 1.4.6
+		\item compiler: 4.0.0
+		\item pillar: 1.4.4
+		\item scales: 1.1.1
+		\item pkgconfig: 2.0.3
+	\end{itemize}
+\end{itemize}
 % section Technology end 

### 2020-08-17 Tex/MSc Data Science Project.tex
Add abstract, intro, related work, bib fixes and small changes

@@
 \begin{abstract}
-	Coming soon...
+	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in and attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image.
+We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work.
+Using an Inception-ResNet-v2 we achieve Top-1 accuracy of 24.7\% and a Top-3 accuracy of 45.4\% which is an improvement over previous results.
 \end{abstract}
@@
 \label{sec:intro}
+Convolutional neural networks (CNNs) have been used for image classification for the last 30 years\cite{Cun1989}, with ever more accurate, and more complex, model architectures being produced and tuned on benchmark datasets.
+CNNs have been used to identify clothing types, objects, music, and even identify artists of paintings, but as far as we are aware only once has someone before asked the question of the ability of a CNN to judge a book by its cover i.e.
+predict the genre of the book based on the cover image.
+We will discuss the work of this person in \cref{sub:Related Works}.
 
-Introduce the project, describe the goal and the idea behind why this is important. Explain the structure of this work,
+The question of predicting a genre based only on a cover image is more difficult than it may first appear; many books at least slightly span genres, and designs are at their heart art which means some personal interpretation will always play a part.
+Being able to successfully build a model to solve this problem would allow designers to use such a tool to ensure that a design is clear to their target audience, it could allow some data-validation of book related data and could help uncover some underlying design principles in genre classification that is not evident to us currently.
 
-In this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.
+In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image.
+The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Technology} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs.
+In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images.
+\Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising out work in \cref{sec:Conclusion}.
+Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.
+
 \subsection{Related Works} 
 \label{sub:Related Works} 
-Discuss the work of Iwana as the main source, reference their methods and results but keep much for later where relevant for comparison with my work.
+So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work completed by Iwana et al.\cite{KenjiIwana}.
+In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}.
+These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}.
+In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.
+
+They go on in their work to discuss the results of their models with analysis and call out few cover design principles including the use of colour, objects, and text.
+We further investigate the colour angle in \cref{sub:Further_Analysis}.
+
+Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly epochs than the original paper.
 % subsection Related Works end 
@@
 	\item Bold items will refer to full vectors of matrices, with it clear which one from the context.
-	\item $\vec{X}$ will always refer to the vector input patterns i.e. images of a dataset. Each element contains more dimensions so this is technically a 4D-tensor.
-	\item $\vec{Y}$ will always refer to the vector of target classes for each of those patterns. Each element is a single value.
-	\item Bracketed roman superscripts i.e. $w^{(i)}$ will be used to refer to, in this case the i\textsuperscript{th}, elements of these vectors. These are 3D-tensors.
-
-	\item Greek subscripts will be used to refer to elements along the dimensions of each pattern i.e.
-$x^{(i)}_{\alpha, \beta, \gamma}$ refers to the $alpha$ element of down on the tensor, the $\beta$ element across on the tensor, and the $\gamma$ channel of the tensor.
-	\item $\bigodot$ will refer to element wise multiplication of matrices and tensors across all dimensions where indices are not explicitly summed over in an equation.
-	\item $\bigoplus$ will refer to element wise addition of tensors across all dimensions where indices are not expli
 \end{itemize}
@@
 \begin{equation}
-	f_{ReLU6}(x) = min(max(0, x), 6) \label{eq:act_relu6}
+	f_{ReLU6}(x) = \min(\max(0, x), 6) \label{eq:act_relu6}
 \end{equation}
@@
 \subsubsection{Residual Networks}
-Residual networks (ResNets) were introduced by He et al in 2015\cite{He2015a} to allow for far deeper networks than previously used.
+Residual networks (ResNets) were introduced by He et al.
+in 2015\cite{He2015a} to allow for far deeper networks than previously used.
 The basic premise is that the layer input forms part of the output directly via a scaled addition to the convolved, normalised, and activated output.
@@
 \label{sub:Activation_Maximisation} 
-Describe the theory and/or maths of activation maximisation and the methods used within the package we use the regularisations applied. 
+TBC: Describe the theory and/or maths of activation maximisation and the methods used within the package we use the regularisations applied. 
 % subsection Activation_Maximisation end 
@@
 
-The models themselves, MobileNetV2, Inception-ResnetV2, and ResNeXt50 were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}.
-MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResnetV2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures.
+The models themselves, MobileNetV2, Inception-ResNet-v2, and ResNeXt50 were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}.
+MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResNet-v2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures.
 These models, combined with the original work done by Iwana et al, span a range of complexity and training speed allowing for a good comparison between performance.
@@
 
-Iwana et al trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs.
+Iwana et al.
+trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs.
 It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.
 
-MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNeXt50 continued to see improvement until just before the 50th epoch/ ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
+MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNet-v2 continued to see improvement until just before the 50th epoch, ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
 % subsection Training_Performance end
@@
 As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset did not contain the set of records; but for a comparison it is suitable.
-Inception-ResnetV2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account.
+Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account.
 MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres.
@@
 
-At a per-genre level Inception-ResnetV2 shows the best results on 16 out of the 30 genres for top-1 predictions, followed by 10 for AlexNet.
-If we further explore the seemingly best performing classes in ResNeXt50 we can see by looking at the precision as presented in \cref{tab:test_prec} that in the genres where this model seems to perform well the precision is low (when compared to for example Inception-ResnetV2) which provides further evidence that for some classes it combines multiple genres and just predicts one of them i.e.
+At a per-genre level Inception-ResNet-v2 shows the best results on 16 out of the 30 genres for top-1 predictions, followed by 10 for AlexNet.
+If we further explore the seemingly best performing classes in ResNeXt50 we can see by looking at the precision as presented in \cref{tab:test_prec} that in the genres where this model seems to perform well the precision is low (when compared to for example Inception-ResNet-v2) which provides further evidence that for some classes it combines multiple genres and just predicts one of them i.e.
 whilst the model is able to correctly predict these \emph{chosen}	classes when it sees them, it is highly likely to predict other genres as one of these as well.
 
-Overall the results from Inception-ResnetV2 are promising given the low training time and adjusted dataset; the potential improvements are discussed further in \cref{sub:Discussion}.
+Overall the results from Inception-ResNet-v2 are promising given the low training time and adjusted dataset; the potential improvements are discussed further in \cref{sub:Discussion}.
 
@@
 	\begin{tabular}{lrrrrrrrrrrrrrr}
-								   & \multicolumn{2}{c}{\textbf{LeNet*}}                                     & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{AlexNet*}}                                   & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{MobileNetV2}}                                & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{Inception-ResnetV2}}                         & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{ResNeXt50}}                                  \\
+								   & \multicolumn{2}{c}{\textbf{LeNet*}}                                     & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{AlexNet*}}                                   & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{MobileNetV2}}                                & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{Inception-ResNet-v2}}                         & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{ResNeXt50}}                                  \\
 	\textbf{Genre}                 & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} \\ \hline
@@
 	\begin{tabular}{lrrr}
-	\textbf{Predicted Genre}      & \multicolumn{1}{c}{\textbf{MobileNetV2}} & \multicolumn{1}{c}{\textbf{InceptionResnet-V2}} & \multicolumn{1}{c}{\textbf{ResNeXt50}} \\ \hline
+	\textbf{Predicted Genre}      & \multicolumn{1}{c}{\textbf{MobileNetV2}} & \multicolumn{1}{c}{\textbf{Inception-ResNet-v2}} & \multicolumn{1}{c}{\textbf{ResNeXt50}} \\ \hline
 	Arts \& Photography           & 17.8                                     & 20.8                                            & 25.9                                   \\
@@
 As we can see in \cref{fig:class_overlap}, where only class overlaps of greater than 0.5 are displayed, the ResNeXt50 model sees large amounts of class overlap between many classes, which agrees with what we saw before that certain classes were being \emph{confused} for each other by the model.
-Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResnetV2 sees none.
+Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResNet-v2 sees none.
 When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; Self-Help has overlap with Health, Fitness and Dieting, Science and Maths has overlap with Medical Books, and Religion and Spirituality has overlap with Christian Books and Bibles.
@@
 
-We ran the activation maximisation on the Inception-ResnetV2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images.
+We ran the activation maximisation on the Inception-ResNet-v2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images.
 The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine.
@@
 	\includegraphics[scale=0.6]{visualize-dense-layer.png}
-	\caption{Activation maximisation outputs for the 3 best predicted classes of the Inception-ResnetV2 model}
+	\caption{Activation maximisation outputs for the 3 best predicted classes of the Inception-ResNet-v2 model}
 	\label{fig:act_max_output}
@@
 \label{sub:Discussion} 
-Overall the Inception-ResnetV2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results.
-The counterpoint to this is that the AlexNet model as trained by Iwana et al was trained for 450,000 epochs where out model was trained for less than 50 epochs total.
+Overall the Inception-ResNet-v2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results.
+The counterpoint to this is that the AlexNet model as trained by Iwana et al.
+was trained for 450,000 epochs where out model was trained for less than 50 epochs total.
 We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greatly difference between the results.
@@
 \label{sec:Conclusion} 
-In this work we have attempted to improve on the work by Iwana et al and better answer the question can a CNN judge a book by its cover.
+In this work we have attempted to improve on the work by Iwana et al.
+and better answer the question can a CNN judge a book by its cover.
 We reprocessed the dataset initially collected by Iwana and applied multiple pre-processing techniques in an attempt to identify the best method to deal with the impact of different shaped images.
@@
 
-We have seen that using a more modern CNN architecture, namely Inception-ResnetV2, we are able to improve on the results published by Iwana slightly whilst training for substantially less epochs.
+We have seen that using a more modern CNN architecture, namely Inception-ResNet-v2, we are able to improve on the results published by Iwana slightly whilst training for substantially less epochs.
 We have also provided some further supporting insight into the components that factor into book cover design i.e.
@@
 % subsection MobileNetV2 end  
-\subsection{Inception-ResnetV2} 
+\subsection{Inception-ResNet-v2} 
 \label{sub:Inception-ResnetV2} 
@@
 	\includegraphics[scale=0.25]{inception_resnetv2.png}
-	\caption{High level view of the inception-resnetv2 architecture. The early classifier (bottom of image) is not used in our work.}
+	\caption{High level view of the Inception-ResNet-v2 architecture. The early classifier (bottom of image) is not used in our work.}
 	\label{fig:incepresnetv2}
@@
 	}
-	\caption{Structure of the Inception-ResNetV2 model.
+	\caption{Structure of the Inception-ResNet-v2 model.
 Shape unless otherwise specified is square, and N is the number of times that layer or block is repeated.
@@
 
-% subsection Inception-ResnetV2 end 
+% subsection Inception-ResNet-v2 end 
 \subsection{ResNeXt50} 
@@
 	}
-	\caption{Structure of the ResNext-50 model. At the end of each block the input is added in (as with the original resnet models) - see example methods in \cref{fig:resnextblocks}.}
+	\caption{Structure of the ResNeXt50 model. At the end of each block the input is added in (as with the original resnet models) - see example methods in \cref{fig:resnextblocks}.}
 	\label{tab:resnextarc}

### 2020-08-25 Tex/MSc Data Science Project.tex
Update graph caption and report updates

@@
 \usepackage{mdframed}
-\usepackage[margin=3cm]{geometry}
+\usepackage[margin=2cm]{geometry}
 \usepackage[superscript,nomove]{cite}
@@
 \begin{abstract}
-	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in and attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image.
+	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image.
 We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work.
@@
 Convolutional neural networks (CNNs) have been used for image classification for the last 30 years\cite{Cun1989}, with ever more accurate, and more complex, model architectures being produced and tuned on benchmark datasets.
-CNNs have been used to identify clothing types, objects, music, and even identify artists of paintings, but as far as we are aware only once has someone before asked the question of the ability of a CNN to judge a book by its cover i.e.
+CNNs have been used to identify clothing types, objects, music, and even identify artists of paintings; but as far as we are aware only once has someone tested asked the question of the ability of a CNN to judge a book by its cover i.e.
 predict the genre of the book based on the cover image.
@@
 In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image.
-The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Technology} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs.
+The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs.
 In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images.
@@
 These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}.
-In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.
+In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% Top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.
 
@@
 \label{sec:Theory} 
-Whilst in this work we do not detail the full design and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks.
+Whilst in this work we do not detail the full design (beyond that provided in \cref{sec:CNN_Architectures}) and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks.
 In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}.
 We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks.
-For this section and the remainder of this work, unless otherwise specified, we observe the following notational conventions:
-\begin{itemize}
-	\item Unless otherwise stated all tensors (including vectors and matrices) are indexed from 1.
-	\item Roman indices will be used to reference specific components of a tensor within an equation, Greek indices will be used for dimensions to be operated over only.
-	\item Dimensions of each tensor when referring to an image or feature map will be referred to in the order of height, width, (colour) channel (interchangeable with depth), including in the indices.
-	\item Bold items will refer to full vectors of matrices, with it clear which one from the context.
-\end{itemize}
 \subsection{Components of Convolutional Neural Networks} 
@@
 \subsubsection{Convolutional layers}
-A convolutional layer, the key part of a neural network, is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN.
+A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN.
 A convolutional layer is defined by 3 key parameters, and the additional choice of padding.
@@
 This value is the the output value of the top left cell of the output tensor.
-The filter is then moved left by the number of elements defined by the stride; when the end of the row has been reached the filter is move down by the same stride.
+The filter is then moved left by the number of elements defined by the stride; when the end of the row has been reached the filter is moved down by the same stride.
 This process is repeated until the filter has been applied to the entire input.
@@
 Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area.
-These operations tend to either be maximum or average and is determined either by the user of the CNN architecture.
+These operations tend to either be maximum or average and are determined either by the user or the CNN architecture.
 These are applied per-channel and no aggregation across channels is completed afterwards.
@@
 all but the fully connected layers at the end, and just train the classification top of the model.
-In our work we took the second approach as given the size of the models we are using training the full model would have been computationally expensive, and given the relatively small amount of training data would also have when compared to the number of parameters in the model it would have been unlikely to have improved accuracy.
+In our work we took the second approach as, given the size of the models we are using, training the full model would have been computationally expensive, and given the relatively small amount of training data would also have, when compared to the number of parameters in the model, been unlikely to have improved accuracy.
 % subsection Transfer_Learning end
@@
 \label{sub:Activation_Maximisation} 
-TBC: Describe the theory and/or maths of activation maximisation and the methods used within the package we use the regularisations applied. 
+Activation maximisation is a technique that can be applied to neural networks, in particular CNNs, to help understand what an idealised input for a given output the network would expect.
+Is is often combined with saliency or attention maps to identify the area(s) of a specific image that the CNN is drawing on for the classification.
+Activation maximisation is used instead to produce an input for the network that maximises the activation of the target class.
+
+Given the potentially complex networks that can be reviewed using this technique the theory itself remains quite simple in nature.
+Given a trained model, as defined by the architecture and parameters, that ends with a fully connected dense layer the first step is to freeze all the parameters for the model i.e.
+the weights and biases.
+Next, the activation function in the final layer (most likely a softmax function) should be replaced with a linear function instead.
+The reason for this replacement is subtle but important; due to the softmax function taking as an input the entire vector of values it is possible to maximise a single value by minimising the other values.
+If we were to continue to use the softmax it could be the case that in trying to generate an image that most represents class A, we instead generate an image the least represents all other classes; by using a linear activation instead we avoid this issue entirely.
+
+Finally, a random input, $\bm{x}$ is chosen at the start and is evolved using backpropagation using the learning rule defined as
+
+\begin{equation}
+	\label{eq:act_max}
+	\Delta x_i = -\eta\frac{\partial \hat{y}_t}{\partial x_i} + \mathcal{R}
+\end{equation}
+where $\eta$ is the learning rate which in this work was set at 1, $\mathcal{R}$ is any regularisation terms, and $\hat{y}_t$ is the activation value of the target class.
+Simply put, change the values for the input in a way that increases the activation for that class within some constraints.
+One regularisation term used in this work is L2norm.
+It is defined as the root mean squared value of the vector and in this case helps ensure to generate a smooth image, penalising edges.
 % subsection Activation_Maximisation end 
@@
 the genre classification.
+Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.
 
-Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.
 At this point we attempted to identify if any records used in the bookcover30 train or test datasets were actually boxsets, and as such would likely have an unrepresentative cover image, an example of such an image can be seen in \cref{fig:bad_cover}.
@@
 264 records within the bookcover30 dataset matched at least one of these terms and were extracted into a new folder for manual review.
-The process was somewhat subjective but as a rule of thumb we rejected any image that was not a single front facing cover, that contained pictures of multiple books and was not an arrangement of covers; single covers that seemed to be specifically designed for a boxset were considered acceptable.
+The process was somewhat subjective but as a rule of thumb we rejected any image that was not a single front facing cover or that contained pictures of multiple books and was not an arrangement of covers; single covers that seemed to be specifically designed for a boxset were considered acceptable.
 In total 94 records needed to be removed due to these rules.
@@
 We chose to split the training data out into training and validation data using a 90/10 random split per class to ensure class balance in all of the training, validation, and test sets.
-In total the training set compromised of 46,170 records, the validation set 5130 records, and the test set 5700 records; with 1,539, 171, and 190 records per class respectively.
+In total the training set compromised of 46,170 records, the validation set 5,130 records, and the test set 5,700 records; with 1,539, 171, and 190 records per class respectively.
 
@@
 We next used 3 different image preparation methods on the datasets to attempt to identify the method that would lead to the best accuracy within our models.
-At this stage all methods output an image 299x299 pixels as this was the largest size we would need for any future models.
+At this stage all methods output an image of 299x299 pixels as this was the largest size we would need for any future models.
 By reducing the image size twice, first now and then again when using a specific model, we were potentially losing some information that would lead to a less accurate model; however we chose to do this to speed up training and testing by having to load much smaller data in compared to the original full sized images.
-Before pre-processing was done we collected the initial shape and average RGB values of each image for further analysis later.
+Before pre-processing was done we collected the initial shape and average RGB values of each image for further analysis later in \cref{sub:Further_Analysis}.
 
@@
 \label{sub:Configuration} 
-All model architectures consisted of their respective base CNN structure as detailed in \cref{sec:CNN_Architectures} followed by a global average pooling 2D layer to reduce the dimensionality of our tensor and to ensure a flat 1D vector which is fully connected to a 30 neuron output layer with a softmax activation.
+All model architectures consisted of their respective base CNN structure as detailed in \cref{sec:CNN_Architectures} followed by a global average pooling 2D layer to reduce the dimensionality of our tensor and to ensure a flat 1D vector, this is then fully connected to a 30 neuron output layer with a softmax activation.
 All models were also trained using the same setup in terms of hyperparameters, optimisers and loss functions; the loss function was sparse categorical cross entropy (sparse due to the usage of a non one-hot encoded target vector), the optimiser was the Adam optimiser \cite{Kingma2015}.
@@
 
-The models themselves, MobileNetV2, Inception-ResNet-v2, and ResNeXt50 were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}.
+The models themselves, MobileNetV2 \cite{Sandler2018}, Inception-ResNet-v2 \cite{Szegedy2016}, and ResNeXt50 \cite{Xie2016} were chosen to represent a wide spread of model capabilities and complexities as highlighted in Figure 1 of the Benchmark Analysis of CNNs completed by Bianco et al\cite{Bianco2018}.
 MobileNetV2 was chosen as it is a lightweight model designed to be capable of running on mobile devices and was meant to be a low complexity comparison in line with AlexNet (due to no pre-trained version of AlexNet being available within TensorFlow), Inception-ResNet-v2 for its high accuracy in other use cases, and ResNeXt50 as a smaller version of one of the newer CNN architectures.
@@
 The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most.
+The training was completed over multiple sessions due to timeout constraints of the Google Colab platform and the speed of training even using a GPU instance; during each session whenever the validation accuracy of an epoch outperformed the previous best this model was saved with the epoch identifier to Google Drive.
+When a new session was started the weights and biases of this snapshot were loaded directly in and training we restarted from the same epoch.
+This does mean that there may be slightly differences compared to had the model just been trained in a single session, but these effects should be minimal.
 
-The performance of the models throughout training can be seen in \cref{fig:train_perf} where whilst training accuracy continues to increase over time, the validation accuracy doesn't see as much consistent movement which suggests the model is starting to overfit to the training data as opposed to improving the identification of intrinsic underlying features within the images themselves.
+The performance of the models throughout training can be seen in \cref{fig:train_perf} where, whilst training accuracy continues to increase over time, the validation accuracy doesn't see as much consistent movement which suggests the model is starting to overfit to the training data as opposed to improving the identification of intrinsic underlying features within the images themselves.
 Iwana does not report their training accuracy so we are not able to compare the performance of our training with theirs.
@@
 Iwana et al.
-trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs.
+trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, order of magnitudes more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs.
 It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.
 
-MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNet-v2 continued to see improvement until just before the 50th epoch, ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
+MobileNetV2 achieved the best validation accuracy early on in its training, where Inception-ResNet-v2 continued to see improvement until just before the 50th epoch, ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance; the accuracy of ResNeXt50 at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
 % subsection Training_Performance end
@@
 A visual representation of this for our models only is shown in \cref{fig:test_perf}.
-As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset did not contain the set of records; but for a comparison it is suitable.
+As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the fill set of records; but for a comparison it is suitable.
 Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account.
@@
 	\centering
-	\resizebox{\textwidth}{!}{%
+	%\resizebox{\textwidth}{!}{%
 	\begin{tabular}{lrrr}
@@
 	\end{tabular}%
-	}
-	\caption{Top-1 Per class and total precision for each model}
+	%}
+	\caption{Top-1 Per class, and total, precision for each model \\ \tiny{Missing values represent no predictions made in that class over the test set}}
 	\label{tab:test_prec}
@@
 First, as displayed in \cref{fig:col_distributions}, we can see that whilst there is a wide range of average RGB values and width/aspect ratio (height is almost entirely consistent across genres), there are no standout values in colour alone that allow for easy genre classification.
-Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some of the more wider books than many other categories.
+Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some wider books when compared to other categories.
 Whilst none of this alone would be enough to provide good classification, it does suggest that explicit inclusion of dimensional information in a model could improve the performance when combined with the existing output.
@@
 The average overall colour per genre is presented in \cref{fig:avg_col} which does suggest some slight differences between the genres; similar to the callouts made by Iwana, Sci-Fi and Fantasy seem to be darker when compared to something like Cookbooks, Food and Wine.
-We can compare the average colour of the true classes vs those predicted by each model and visualise how these correlate with the recall per class as shown in \cref{fig:col_distance}.
+We can compare the average colour of the true classes against those predicted by each model and visualise how these correlate with the accuracy per class as shown in \cref{fig:col_distance}.
 Here the colour distance is defined as
@@
 \end{equation}
-where R/G/B are the average red, green, and blue values for each predicted and class respectively.
+where R/G/B are the average red, green, and blue values for each predicted and true class respectively.
 We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class.
@@
 Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResNet-v2 sees none.
-When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; Self-Help has overlap with Health, Fitness and Dieting, Science and Maths has overlap with Medical Books, and Religion and Spirituality has overlap with Christian Books and Bibles.
+When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; \emph{Self-Help} has overlap with \emph{Health, Fitness and Dieting,} \emph{Science and Maths} has overlap with \emph{Medical Books}, and \emph{Religion and Spirituality} has overlap with \emph{Christian Books and Bibles}.
 All of these are logical overlaps and were a human tasked with classifying books based on their cover into these categories it is highly probable that different people would classify them differently.
@@
 The goal of using feature visualisation was to attempt to understand what, if any, underlying structure or information the models were using to aid their classification.
-To do this we use activation maximisation as implemented into the python package \emph{keras-vis}\cite{keras-vis} to do all the work.
+To do this we use activation maximisation as implemented into the python package \emph{tf-keras-vis}\cite{keras-vis} to do the computational work.
 To produce the images the final layer activation function was changed from softmax to linear to ensure that class is being visualised as opposed to minimising other classes, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}.
@@
 The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine.
-During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure imagenet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).
+During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure ImageNet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).
+Even when all regularisation terms were removed, allowing the process to fully maximise the activation without constraints we still did not see an improvement in the loss values.
 
@@
 The counterpoint to this is that the AlexNet model as trained by Iwana et al.
-was trained for 450,000 epochs where out model was trained for less than 50 epochs total.
-We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greatly difference between the results.
+was trained for 450,000 epochs where our model was trained for less than 50 epochs total.
+We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greater difference between the results.
 Despite training on slightly less data (due to the validation set), we can see that more modern and powerful CNNs are able to predict book genres with slightly improved accuracy compared to older methods.
@@
 \label{sec:Conclusion} 
-In this work we have attempted to improve on the work by Iwana et al.
+In this work we have attempted to improve on the work of Iwana et al.
 and better answer the question can a CNN judge a book by its cover.
@@
 
-We have seen that using a more modern CNN architecture, namely Inception-ResNet-v2, we are able to improve on the results published by Iwana slightly whilst training for substantially less epochs.
+We have seen that by using a more modern CNN architecture, namely Inception-ResNet-v2, we are able to improve on the results published by Iwana slightly, whilst training for substantially less epochs.
 We have also provided some further supporting insight into the components that factor into book cover design i.e.

### 2020-08-25 Files and scripts/combined_file.R

@@ -329,13 +329,12 @@ comb_data_dist %>%
 
 results_clean_res %>%
   ggplot(aes(x = Result, y = reorder(classes, desc(classes)), group = Model, col = Model)) + 
-  geom_point(shape = 15, alpha = 0.7, size = 2.5) + 
+  geom_point(aes(shape = Model), alpha = 0.7, size = 2.5) + 
   scale_color_brewer(palette="Set1", name = 'Model') +
   facet_wrap(~N, scales = 'free_x') +
   theme_bw() +
   labs(x = 'Test Accuracy',
-       y = 'True Class',
-       caption = 'Top 5 classes by Average Accuracy across all models called out specifically') +
+       y = 'True Class') +
   theme(legend.position = 'bottom',
         plot.caption = element_text(face = 'italic'),
         panel.grid.major.x = element_blank(),

### 2020-09-18 Tex/MSc Data Science Project.tex
Add examples

@@
 \end{equation}
-where $\bm{m}_k$ is the feature map for output channel $k$, C is the number of channels in the input, $x$ is the input, $f$ is the $k^{th}$ kernel of size $l\times l$, $s$ is the stride.
-In total for a square input and kernel this leads to an output of the dimensions described in
+where $m_{i,j,k}$ is the component of the feature map $\bm{m}_k$ for output channel $k$ at position $(i, j)$, C is the number of channels in the input, $x$ is the 3 dimensional input tensor (height, width, and channels) input, $f$ is the $k^{th}$ kernel of size $l\times l$, $s$ is the stride.
+The complexity of the subscripts on $x$ arises from the need to describe the offset provided by the size and stride of the kernel.
+We apply this equation for the examples given in \cref{fig:conv_process} in the following, noticing the simplifications that $C = s = 1$ so that $k \subseteq \{1\}$, and in this case $l = 3$.
+Then for the examples given in the image we have
+\begin{align}
+	m_{1, 1, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (1-1)),1}	\\
+	m_{1, 1, 1} &= (1\times1 )+ (0 \times 2) + (1 \times 1) + (0 \times 2) + (1 \times 0) + (0 \times 0) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 5 \\
+	m_{1, 2, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (2-1)),1}	\\
+	m_{1, 2, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 1) + (0 \times 1) + (1 \times 0) + (0 \times 1) + (1 \times 0) + (0 \times 2) + (1 \times 1) = 3 \\
+	\vdots \nonumber	\\
+	m_{3, 3, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(3-1)), (\beta + (3-1)),1}	\\
+	m_{3, 3, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 0) + (0 \times 0) + (1 \times 2) + (0 \times 1) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 7
+\end{align}
+Note that in the image the kernel is displayed second visually but due to the order we wrote the equation is in the kernel elements are on the left of the multiplication.
+
+In total for a square input and kernel this leads to an output of the dimensions described by
 
@@
 When followed by a 1x1 convolution (also known as a pointwise convolution) this has the effect of merging the layers via addition and combined these two processes actually have a reduced computational cost when compared to a traditional convolution \cite{Sandler2018}.
+The results from depth-wise convolutions are identical to to traditional convolutions but due to this improvement in computational cost they are used within many complex models to improve performance.
 
@@
 Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned.
+For example global average pooling on the output feature map of \cref{fig:conv_process} would be a 1x1 result of 7, global average pooling would give 4.333\dots and max pooling of size 2x2 with a stride of 1 would lead to
+$$
+\begin{bmatrix}
+	6 & 6 \\
+	6 & 7 \\
+\end{bmatrix}
+$$
+as where average pooling of the same size and stride would result in
+$$
+\begin{bmatrix}
+	5.25 & 4.25 \\
+	5.25 & 4.5 \\
+\end{bmatrix}
+$$.
 

### 2020-09-25 Tex/MSc Data Science Project.tex
Final commit

@@
 \usepackage{bm}
+\usepackage{float}
 \usepackage{amsmath}
@@
 \begin{abstract}
-	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image.
+	In this work we revisit and build upon the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image.
 We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work.
@@
 \end{abstract}
+\renewcommand{\abstractname}{Acknowledgements}
+\begin{abstract}
+Thank you to Pete Stewart and Openreach Ltd for sponsoring this degree, and thank you David, Scott, and Craig who made working while studying a little easier and less stressful than it would have been otherwise.
+Thank you to my friends who have put up with me over the last 2 years.
+Finally most of all I want to thank my parents, who have always supported me and my learning, even though it has been a long time since they understood any of it.
+\end{abstract}
+
 \begin{figure}[!b]
@@
 The question of predicting a genre based only on a cover image is more difficult than it may first appear; many books at least slightly span genres, and designs are at their heart art which means some personal interpretation will always play a part.
-Being able to successfully build a model to solve this problem would allow designers to use such a tool to ensure that a design is clear to their target audience, it could allow some data-validation of book related data and could help uncover some underlying design principles in genre classification that is not evident to us currently.
+Being able to successfully build a model to solve this problem would allow designers to use such a tool to ensure that a design is clear to their target audience, it could allow some data-validation of book related data, and could help uncover some underlying design principles in genre classification that is not evident to us currently.
 
 In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image.
-The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs.
+The remainder of this work is structured as follows; in the rest of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs.
 In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images.
-\Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising out work in \cref{sec:Conclusion}.
+\Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising our work in \cref{sec:Conclusion}.
 Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.
@@
 \label{sub:Related Works} 
-So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work completed by Iwana et al.\cite{KenjiIwana}.
+So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work published by Iwana et al.\cite{KenjiIwana}.
 In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}.
@@
 
-Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly epochs than the original paper.
+Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly fewer epochs than the original paper.
 % subsection Related Works end 
@@
 In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}.
-We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks.
+We assume the reader already has familiarity with introductory machine learning concepts such as backpropagation and fully connected neural networks.
 \subsection{Components of Convolutional Neural Networks} 
@@
 \subsubsection{Activation functions}
-Most layers of a neural network have some kind of activation function; these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range).
+Most layers of a neural network have some kind of activation function (a function that transforms the output of the neuron before passing it to the next layer); these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range).
 Some normalisation is also usually applied which is discussed later.
@@
 \subsubsection{Convolutional layers}
-A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN.
+A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classification, forming the fundamental building block of a CNN.
 A convolutional layer is defined by 3 key parameters, and the additional choice of padding.
@@
 \end{equation}
-where $m_{i,j,k}$ is the component of the feature map $\bm{m}_k$ for output channel $k$ at position $(i, j)$, C is the number of channels in the input, $x$ is the 3 dimensional input tensor (height, width, and channels) input, $f$ is the $k^{th}$ kernel of size $l\times l$, $s$ is the stride.
+where $m_{i,j,k}$ is the component of the feature map $\bm{m}_k$ for output channel $k$ at position $(i, j)$, C is the number of channels in the input, $x$ is the 3 dimensional input tensor (height, width, and channels) input, $f$ is the $k^{th}$ filter of size $l\times l$, and $s$ is the stride.
 The complexity of the subscripts on $x$ arises from the need to describe the offset provided by the size and stride of the kernel.
@@
 When followed by a 1x1 convolution (also known as a pointwise convolution) this has the effect of merging the layers via addition and combined these two processes actually have a reduced computational cost when compared to a traditional convolution \cite{Sandler2018}.
-The results from depth-wise convolutions are identical to to traditional convolutions but due to this improvement in computational cost they are used within many complex models to improve performance.
+The results from depth-wise convolutions are identical to traditional convolutions but due to this improvement in computational cost they are used within many complex models to improve performance.
 
@@
 \end{bmatrix}
-$$.
+$$
 
@@
 transforming the output of that layer for that batch to have a mean of 0 and a standard deviation of 1.
-This has the effect of greatly reducing training time and increases accuracy by way of reducing the change in the data distribution between layers and reducing the change of this over time via backpropagation.
+This has the effect of greatly reducing training time and increases accuracy by way of reducing the change in the data distribution between different layers and more importantly reducing the change of this over time via backpropagation.
 
@@
 
-In transfer learning it is possible to take one of two approaches; the first is to use the existing model weight and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e.
+In transfer learning it is possible to take one of two approaches; the first is to use the existing model weights and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e.
 all but the fully connected layers at the end, and just train the classification top of the model.
@@
 \label{sub:Image Download and Manual Review} 
-The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}.
-The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway.
+The 207,572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}.
+The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates.
+The overhead of reading data between Google Drive and Google Colab meant that even checking before downloading an individual file if it already existed was not much faster than downloading and writing the file anyway.
 Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e.
@@
 
-The models were training once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point.
-No cross-validation was done due to resource limitations.
+The models were trained once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point.
+No k-fold cross-validation was done due to resource limitations.
 % subsection Configuration end 
@@
 A visual representation of this for our models only is shown in \cref{fig:test_perf}.
-As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the fill set of records; but for a comparison it is suitable.
+As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the full set of records; but for a comparison it is suitable.
 Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account.
@@
 \begin{equation}
-	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
+	d_{colour} = \sqrt{(R_2 - R_1)^2 + (G_2 - G_1)^2 +(B_2 - B_1)^2 }
 \end{equation}
@@
 We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class.
-This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements.
+This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however it highlights that even by just guessing based on closeness to average colour you can make some small improvements.
 This supports the claims made by Iwana in their work.
@@
 Finally, we can review the \emph{class overlap} for each model.
-We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where A has a higher probability) i.e.
+We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where B has a higher probability) i.e.
 \begin{equation}
@@
 To do this we use activation maximisation as implemented into the python package \emph{tf-keras-vis}\cite{keras-vis} to do the computational work.
-To produce the images the final layer activation function was changed from softmax to linear to ensure that class is being visualised as opposed to minimising other classes, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}.
+To produce the images the final layer activation function was changed from softmax to linear to ensure that target class is being visualised as opposed to just minimising minimising other class activations, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}.
 L2 normalisation was used as the default to attempt to avoid pure static images being produced.
@@
 The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine.
-During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure ImageNet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).
+During these steps there was some improvement to the loss function however did not appear to converge.
+When compared to the progress seen on a pure ImageNet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).
 Even when all regularisation terms were removed, allowing the process to fully maximise the activation without constraints we still did not see an improvement in the loss values.
@@
 Genre prediction from cover images remains a complex topic and given the subjectivity of the classifications is unlikely to ever be fully solved, however the work produced here gets a little closer than we were before.
-So it seems for now that, if you are a CNN, you should follow the original advice and not try to judge a book by its cover.
+So it seems for now that, if you are a CNN, you should follow the old adage and don't judge a book by its cover.
 % section Conclusion end
@@
 
-\begin{table}[]
+\begin{table}[H]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 
-	\begin{figure}
+	\begin{figure}[H]
 		\centering
